### 2013-05-29 docs/irmwriteup.tex
writeup of plan

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[citestyle=numeric]{biblatex}
\addbibresource{connectomics motif.bib}

\begin{document}

\title{Nonparametric stochastic blockmodels with spatial structure}
\author{Eric Jonas}

\maketitle

\section{overview}

Classic stochastic blockmodels posit the existence of a latent class
which sets the probability of a relation holding between any
collection of entities in a domain. That is, if $e^1_i$ is an entity
in domain $T1$ and in latent class $m$ and $e^2_j$ is an entity in
domain $T2$ in latent class $n$, then 
\begin{equation}
\operatorname{Relation}(e^1_i, e^2_j) \sim F(\cdot | \eta_{mn})
\end{equation}

Kemp \parencite{Kemp2006a} extends stochastic blockmodels to the nonparametric setting.

For our application we're interested in the directed graph of synapse
connectivity between all pairs of cells. Thus our relation is
``synapses on'' and is defined on $T1\times T1$. We want to determine
the underlying latent class of each cell (``cell type''), as well as
the cell-type-specific connectivity. For this sort of binary
(connected-or-not) data, we'd traditionally use a conjugate
beta-Bernoulli likelihood -- $\eta_{mn}$ is the probability of a
connection between cells of class $m$ and $n$.

In neural systems, we know that we have cell-type-specific
connectivity, but we also know that geometry matters tremendously. In
the retina, certain cell-types tessellate the entire retina, and only
synapse with a subset of neighboring cells. This spatially-selective
connectivity runs very counter to the blockmodel assumptions of latent class
being all that matters.

Based on an off-handed comment about a ``discriminative blockmodel''
that Kevin Murphy, we can extend the connectivity with

\begin{equation}
P(e_i, e_j) \sim \operatorname{Bernoulli}(w_{ab}\cdot f(e_i, e_j)
\end{equation}

where the goal is to have $w_{ab}$ be an arbitrary latent-class weight
matrix that is applied to an arbitrary discriminative function between
additional parameters of entities $e_i$ and $e_j$. That is, $f$ can be
defined on additional per-entity attributes, like spatial location.

But of course, we don't have to be so limiting -- we can in fact
learn an arbitrary class-type-specific function $f_{ab}(e_i, e_j)$. Right
now I'm planning Euclidian distance between $e_i$ and $e_j$ as well as learning
the $\lambda_{ab}$. That is, the probability of a connection between
cells $e_i$ and $e_j$ is now

\begin{equation}
p(e_i, e_j) \sim \operatorname{Bernoulli}\Big(p_{ab}\cdot \operatorname{Exp}(d(e_i, e_j) | \lambda_{a, b})\Big)
\end{equation}

\begin{itemize}
\item The choice of an exponential distribution is entirely arbitrary
  -- it restricts us to assuming that cells always prefer closer
  neighbors than far away ones. Thus we'll probbably eventually want
  to use something like a two-parameter gamma to enable ``cell class A
  only synapses on cell class B when it is more than 40 $\mu m$ away
  and less than $200 \mu m$ away.''
\item We could even go crazy and stick a DP in there and learn
  whole-hog the functional form of $f_{ab}$. But that sounds
  like overkill right now. 
\end{itemize}

\section{Inference}

Inference isn't that bad -- this really just looks like a nonconjugate
IRM! With the original beta-Bernoulli IRM, we integrate out the
$\eta_{ab}$, whereas here we are explicitly tracking the per-block
parameters. We need to stick a prior on them, and then do
non-conjugate inference, but thankfully other
people \parencite{Neal2000} have already done a passable job
making MCMC in nonconjugate models a bit easier.

As we get ``real'' large data, we'll want to certainly explore
more efficient inference approaches, but we can fight that fight
when we come to it.

Other things: we're estimating the latent class number, should we be
using a mixture of finite mixtures? The DP is inconsistent
\parencite{Miller2012} but minor tweaks allow us to use nearly-identical
inference and more robustly learn the number of latent categories.

\printbibliography

\end{document}

### 2013-05-30 irm/fastmodel.h
rough skeleton builds

@@ -76,117 +76,6 @@ struct BetaBernoulli {
     }
 }; 
 
-template<typename CM>
-class  ComponentContainer{ 
-    
-    struct sswrapper_t { 
-        size_t count; 
-        typename CM::suffstats_t ss; 
-    }; 
-
-public:
-    ComponentContainer(int ndim, char * data, 
-                       size_t data_shape[] ) : 
-        NDIM_(ndim), 
-        pdata_(static_cast<typename CM::value_t *>(data))
-    {
-        for(int i = 0; i < NDIM_; i++) { 
-            data_shape_[i] = data_shape[i]; 
-        }
-
-        
-    }
-    
-    void create_component(size_t group_coords[]) {
-        group_hash_t gp = hash_coords(group_coords); 
-        sswrapper_t * ssw = new sswrapper_t; 
-        CM::ss_init(&(ssw->ss), &hps_); 
-        ssw->count = 0; 
-        components_.insert(std::make_pair(gp, ssw)); 
-    }
-
-    void delete_component(size_t group_coords[]) {
-        group_hash_t gp = hash_coords(group_coords); 
-        typename components_t::iterator i = components_.find(gp); 
-        delete i->second; 
-        components_.erase(i); 
-    }
-
-    float total_score() {
-        typename components_t::iterator i = components_.begin(); 
-        float score = 0.0; 
-        for(; i != components_.end(); ++i) { 
-            if(i->second->count > 0) { 
-                score += CM::score(&(i->second->ss), &hps_); 
-            }
-        }
-        return score; 
-           
-    }
-    
-
-    float post_pred(size_t group_coords[], size_t dp_pos)
-    {
-        group_hash_t gp = hash_coords(group_coords); 
-        typename CM::value_t val = pdata_[dp_pos]; 
-        sswrapper_t * ssw = components_.find(gp)->second; 
-        return CM::post_pred(&(ssw->ss), &hps_, val); 
-    }
-    
-    void add_dp(size_t group_coords[], size_t dp_pos) {
-        group_hash_t gp = hash_coords(group_coords); 
-        typename CM::value_t val = pdata_[dp_pos]; 
-        sswrapper_t * ssw = components_.find(gp)->second; 
-        CM::ss_add(&(ssw->ss), &hps_, val); 
-        
-    }
-
-
-    void rem_dp(size_t group_coords[], size_t dp_pos) {
[…]

### 2013-05-30 irm/helloworld.cc

@@ -1,0 +1,7 @@
+#include <iostream>
+
+int main()
+{
+    std::cout << "Hello World" << std::endl; 
+
+}

### 2013-05-30 irm/relation.cc

@@ -1,0 +1,165 @@
+#include "relation.h"
+
+namespace irm { 
+
+Relation::Relation(axesdef_t axes_def, domainsizes_t domainsizes, 
+                   IComponentContainer * cm) :
+    axes_(axes_def), 
+    pCC_(cm), 
+    DIMS_(axes_def.size()), 
+    DOMAINN_(domainsizes.size()), 
+    group_id_(0), 
+    datapoint_groups_(pCC_->dpcount()), 
+    datapoint_entity_index_(pCC_->dpcount())
+{
+    // for (int d = 0; d < DOMAINN_; ++d) { 
+    //     std::vector<int> axispos; 
+    //     for (int i = 0; i < axes_.size(); ++i) { 
+    //         if(axes_[i] == d) { 
+    //             axispos.push_back(d); 
+    //         }
+            
+    //     }
+    //     domain_to_axispos_.push_back(axispos); 
+    //     domain_entity_assignment_[d] = boost::assign::repeat(domainsizes_[d], NOT_ASSIGNED); 
+        
+    //     domain_groups[d].push_back(group_set_t()); 
+    // }
+       
+    
+    // for(int dpi = 0; dpi < pCC_->dpcount(); ++dpi) { 
+    //     datapoint_groups[dpi] = boost::assign::repeat(DIMS_, NOT_ASSIGNED)
+    // }
+
+    // FIXME: add in the cartesian product coordinate stuff
+
+}
+
+
+void Relation::assert_unassigned() { 
+
+
+}
+
+
+void Relation::assert_assigned() {
+    
+
+}
+
+size_t Relation::assigned_dp_count()
+{
+    // FIXME: implement
+
+}
+
+std::vector<int> get_axispos_for_domain(domainpos_t domain) 
+{
+
+
+}
+
+std::vector<int> datapoints_for_entity(domainpos_t domain, entitypos_t entitypos)
+{
+
+
+}
+
+std::vector<int> get_dp_entity_coords(dppos_t dp)
+{
+
+}
+
+
+groupid_t Relation::create_group(domainpos_t domain)
+{
+
+}
+
+void Relation::delete_group(domainpos_t dom, groupid_t gid)
+{
+
+
+}
+
+std::vector<groupid_t> Relation::get_all_groups(domainpos_t)
+{
+
+
+}
+
+// FIXME come up with some way of caching the mutated components
+float Relation::add_entity_to_group(domainpos_t domain, groupid_t group_id, 
[…]

### 2013-05-30 irm/relation.h

@@ -1,0 +1,59 @@
+#ifndef __IRM_RELATION_H__
+#define __IRM_RELATION_H__
+
+#include "util.h"
+#include "componentcontainer.h"
+
+namespace irm { 
+
+class Relation
+{
+    /*
+      Domains are numbered, and axes are a list of domain integers. So for a T1xT1 
+      we would specify 
+      
+      
+    */
+public:
+    Relation(axesdef_t axes_def, domainsizes_t domainsizes, 
+             IComponentContainer * cm); 
+    
+    void assert_unassigned(); 
+    void assert_assigned();
+    size_t assigned_dp_count();
+
+    groupid_t create_group(domainpos_t domain); 
+    
+    void delete_group(domainpos_t dom, groupid_t gid); 
+
+    std::vector<groupid_t> get_all_groups(domainpos_t); 
+
+    float add_entity_to_group(domainpos_t, groupid_t, entitypos_t); 
+    
+    float remove_entity_from_group(domainpos_t, groupid_t, entitypos_t); 
+    float post_pred(domainpos_t, groupid_t, entitypos_t); 
+
+    float total_score(); 
+private:
+    const axesdef_t axes_; 
+    IComponentContainer * pCC_; 
+    const int DIMS_; 
+    const int DOMAINN_; 
+    std::vector<std::vector<int> > domain_to_axispos_; 
+
+    groupid_t group_id_; 
+
+    // maps from a datapoint to that datapoint's current group coordinates
+    std::vector<group_coords_t> datapoint_groups_; 
+    // maps from dp index to entity coordinates
+    std::vector<entity_coords_t> datapoint_entity_index_; 
+
+    // for each domain, maps from its entities to the domain
+    std::vector<std::vector<std::vector<int> > > entity_to_dp; 
+    
+}; 
+
+}
+
+
+#endif 

### 2013-05-30 irm/relation.py

@@ -205,7 +205,7 @@ class FastRelation(object):
             self._set_dp_group_coords(dp, ngc)
         self._set_entity_group(domain_name, entity_pos, NOT_ASSIGNED)
 
-
+        
     def post_pred(self, domain_name, group_id, entity_pos):
         """
         Well crap. This can't be a mutation-free operation. Damn. 

### 2013-05-30 irm/testrun.py

@@ -45,7 +45,7 @@ print "COMPUTING TRUE SCORE", "-"*40
 #print "ALL ONE SCORE", irm_model.total_score()
 
 
-SAMPLES = 100
+SAMPLES = 10
 ITERS_PER_SAMPLE = 1
 for s in range(SAMPLES):
     print "sample", s

### 2013-05-30 irm/util.h

@@ -1,0 +1,31 @@
+#ifndef __IRM_UTIL_H__
+#define __IRM_UTIL_H__
+
+#include <list>
+#include <set>
+#include <vector>
+#include <stdlib.h>
+
+namespace irm { 
+
+const int MAX_AXES = 10; 
+
+typedef std::vector<int> axesdef_t; 
+typedef std::vector<size_t> domainsizes_t; 
+typedef int domainpos_t; 
+typedef int groupid_t; 
+typedef size_t dppos_t;
+typedef size_t entitypos_t; 
+
+typedef std::set<groupid_t> group_set_t; 
+
+typedef std::vector<groupid_t> group_coords_t; 
+typedef std::vector<entitypos_t> entity_coords_t; 
+
+
+#define NOT_ASSIGNED -1
+
+}
+
+
+#endif

### 2013-10-20 docs/paper/paper.tex
updates to runner long-running drosophila results

\documentclass{article}
\usepackage[english,american]{babel}
\usepackage[disable]{todonotes}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}

\title{Neural Circuit Discovery}
\author{Eric Jonas \\ Konrad Kording}

\begin{document}
\maketitle

\listoftodos

\begin{abstract}
  The brain is made out of multiple types of neurons, and a neuron’s
  type affects both its behavior and which other neurons it connects
  to. Emerging techniques (connectomics) allow measuring the
  connectivity matrix between many neurons. Estimating types from
  connectomics is difficult as distance is usually more important than
  type. Here we describe a nonparametric Bayesian technique that
  overcomes this problem and discovers neuron types based on
  connections only. We show that the approach recovers known neuron
  types in the retina, reveals interesting structure in the nervous
  system of c. elegans, and automatically discovers the structure of
  microprocessors. Extracting meaningful structure from connectivity
  data promises to enable new experiments and to constrain theories of
  brain function.

\end{abstract}

\section{Introduction}
``type'' is a recurreing theme in biology
Cell type in bio literature, stem cells, something something

It is only through understanding the types of the system that we can understand
its function.

\section{Methods}
Stochastic blockmodels long known

We take an infinite extension of the SBM (the IRM ) \todo{cite both}
and modify it to learn the structure of the underlying code.

Spatial component by performing a GLM-style link function.

Describe the model

We develop a series of Markov-chain monte carlo scheems to solve for
the posterior, presenting the MAP estimate. To avoid local minima we
slowly anneal, over 1000 iterations \todo{Fix numbers}.

Hyperparameter inference.

We show that we recover ground truth.

\section{Results}

\subsubsection{Mouse Retina}
The recent release of connectome-scale data for the mouse retina.

Look we recover known classes!

\subsection{Drosophila}

\subsection{C. elegans}

\section{Discussion}

\end{document}

### 2013-10-20 docs/techreport.tex
updates to runner long-running drosophila results

\documentclass{article}
\usepackage[english,american]{babel}
\usepackage{todonotes}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{fullpage}

\usepackage{csquotes}% Recommended
\usepackage{amssymb}
\usepackage{biblatex}

\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}

\title{Spatial Relational Models for Neural Circuit Discovery}
\author{Eric Jonas \\ Konrad Kording}

\begin{document}
\maketitle

\listoftodos

\begin{abstract}
  The role of ``type''
  Clustering in molecular biology
\end{abstract}

\section{Introduction}

Things I believe

\begin{itemize}
\item The rise of complex data demands better computational tools to ``see'' into it
\item Clustering has worked for molecular biology and genetics
\item ``Type'' is a useful concept for understanding how neurobiological systems work
\item Many things people do with graphs are boring, not biologically relevant
\item Neuroanatomy needs to be a computational field
\end{itemize}

\section{Methods}
Stochastic blockmodels have been around forever, assume a latent ``type'' dictates
the probability of a latent class. We use recently-developed infinite
stochastic block models that

\subsection{infinite Spatial Relational Models}

We adopt the terminology from \autocite{Kemp}. A domain $T_n$ is a
collection of entities $\{e^t_i | i \in 0 \dots N^t -1 \} $ that share
some common property. A relation is a collection of observation
defined on a collection of domains. For example, if one domain is
``students'' and another is ``courses'', the relation
``$\operatorname{has-taken}(e^s_i, e^c_j)$ is true if student $e^s_i$ has
taken course $e^c_j$.

Our model assumes that the probabiltiy of a relation taking on a
particular observed value is only a function of some hidden ``type''
(or class, or cluster) of an entity. Each entity can belong to a
single latent class.

That is, if $e^1_i$ is an entity
in domain $T^1$ and in latent class $m$ and $e^2_j$ is an entity in
domain $T^2$ in latent class $n$, then 
\begin{equation}
\operatorname{Relation}(e^1_i, e^2_j) \sim F(\cdot | \eta_{mn})
\end{equation}

where $\eta_{mn}$ is an unobserved (latent) parameter.
\todo{Better framing of this}

Note that this framework is equivalent to the classic stochastic block
model, defined on a graph -- a relation $\operatorname{has-edge}(e_i,
e_j)$ is defined on the domain $T1$ of nodes.

\subsubsection{Arbitrary discriminative funtions}
We can extend models of this type

\begin{equation}
P(e_i, e_j) \sim F(g_{mn}(e_i, e_j))
\end{equation}

where $g_{mn}$ is an arbitrary link function $g$ defined per-latent
class -- if $e_i$ is of latent class $m$ and $e_J$ is of latent class
$n$, then the likelihood function $F$ is evaluated at $g_{mn}(e_i,
e_j)$ instead of a fixed parameter $\eta_{mn}$ in the classical
model. This allows an arbitrary discriminative function between
additional parameters of entities $e_i$ and $e_j$. That is, $g$ can be
defined on additional per-entity attributes, like spatial location.

We're principally interested in spatial models, so we associate with each
entity $e_i$ a location in space, and parameterize a per-latent-class
link function 
\todo{later}

\subsubsection{Infinite prior}
exchangability 
chinese restaurant process prior

\subsection{Likelihoods}
We evaluated a collection of link functions. 
Many of these are needed to map from $\mathbb{R}^+$ to some closed interval. 
TODO where do we talk about how this is like GLMs?

Let $x_i$ and $x_j$ be the spatial locations of entities $e_i$ and $e_j$. $D(e_i, e_j$ is the euclidian distance between these two points.

\subsubsection{Logistic Distance}
We use a binary observation for the relation (Bernoulli), and the $p$ is determined
on a per-class basis

prior on mu, lambda

\subsubsection{Normal Distance Fixed Width}
This likeilihood model uses a logistic transform from distance to the
probability $p$ of a connection between two entities $e_i$ and
$e_j$. The latent class component parameters are the center of the
logistic $\mu_{mn}$ and the scale $\lambda_{mn}$. If the distance
between $e_i$ and $e_j$ is less than $\mu_{mn}$ then they are more
likely to be connected. To smooth out the likelihood, we scale the logistic
function by $p_{\textrm{min}}$ and $p_{\textrm{max}}$

So the generative model is
\[\mu_{ab} \sim \exp(\mu^{hp}) \]
\[\lambda{ab} \sim \exp(\lambda^{hp}) \]
\[p^* = \frac{1.0}{1.0 + \exp{\frac{(d(e_i, e_j) - \mu_{ab})}{\lambda_{mn}}}}\]
and then 
\[ r \sim \operatorname{Bernoulli}(p\cdot (p_{max} - p_{min}) + p_{min}) \]

\todo{example figure of parameterization, and notes}
We don't do hyperparameter inference on $p_{min}$ and $p_{max}$.

\subsubsection{Exponential Distance Bernoulli}
\todo{Implement and test this model}

\subsubsection{Exponential Distance Poisson}
Some of our observations are count-valued.

\subsubsection{Exponential Distance Normal}

\subsubsection{Non-spatial, Conjugate likelihoods for comparison}
Beta Bernoulli
Gamma-Poisson

\subsection{Inference}
\subsubsection{Structural Inference}

\subsubsection{Per-component parameters in nonconjugate case}
Slice sample on latent parameters

\subsubsection{Hyperparameter Inferece}
Gridded gibbs sampling for hyperparameter inference

\section{Results}

\subsection{Synthetic Data}
Show that traditional stochastic block model fails.

The goal of the synthetic results is to show that we recover ground truth. 
1. where is the data from? 
2. how did we generate it?

\subsection{Mouse Retina}

Dense serial electron microscopy of a $VOLUMNE$ in the mouse
retina by \autocite{MouseReinta} yielded a listing of places where
neurons come into contact. There were $number$ cells originally, whcih
were refined through various classification schemes. We selected the
$NUMBER$ for which the location of the soma could be reconstruted from
the provided cell plots (soma locations were not provided by the
study's authors in machien-readble form).

The authors were unable to disambiguate actual ``synapses'' in the
full data set, but were able to measure the points of contact between
cells. Thus we don't have ``real'' synapses. We also don't have directionality.

\subsection{C. elegans}
Connectome across several animals traced out by sidney brenner back in the day.
Small number of cells
position is linear along body axis
cell divisions are really precise

only look at hermaphrdite, ignore phyranx.

Have directionality, chemical vs electrical, number

\subsection{Drosophila medulla}
\cite{DrosophilaConnectome}
No cell bodies -- what to use for distance? 
Human annotation
what the hell is a cell

\subsection{MOS 6502}

Hobbiest groups seeking to preserve historical microprocessors have
measured the ``connectomes'' of silicon integrated circuits. This
process involves obtaining classic chips, removing the casing
chemically, and then taking multi-gigapixel visible-light micrographs
of the silicon layers contained within. Then, a partially automated
process is used to read out the silicon polygons and reconstruct the
processor circuit contained within. The results are quite impressive,
with groups having successfully measured the original MOS6502 (Found
in the Apple I), and then used this ``connectome'' to perform
cycle-accruate simulation.

We extracted this data at the per-transistor level from the online
github project, resulting in $NUMBER$ transistors. Each transistor has
three terminals: $g$, $c_1$, and $c2$. Electronic circuits are fundamentally
undirected at the physical leve.

Remove CLK, VCC/VSS.

This would work better up one level of abstraction.

Challenges: not directed. Three terminals. Dense connections.

THIS SHOULD WORK

\subsection{Zipcode data}

This data is from career builder. It contains who applied to what jobs. This is our first non-graph example of a T1xT2 relation.

\section{Discussion}
Do we have enough data to make this work?

Why does it fail sometimes?

Tendency of stochastic block models to have block-diagonal structure -- that's not really how neuro works

Is type a static thing or a continuum?

\section{Next Steps / Future Work}
How in gods name do we evaluate this?

What abotu mixed membership models / better structural priors

What about incorporating additional metadata (``attribtue'' data)

Can we use something like a GP to learn the link function? 
Parametric link functions are nice, but you'd like to learn it from the data

How in gods name do we make this scale to 100k nodes? 
VB? spectral methods? Shitty heuristics?

How in gods name do we make this hierarchical? 
We expect there to be hierarchy

\end{document}

### 2013-10-20 docs/fig1.py

@@ -1,0 +1,88 @@
+import numpy as np
+from matplotlib import pylab
+import irm
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+
+SIDE_N = 5
+
+# class_conn = {(0, 1) : ('d', 1.0, 0.7), 
+#               (1, 2) : ('d', 2.0, 0.8), 
+#               (3, 2) : ('p', 0.1), 
+#               (3, 0) : ('d', 1.7, 0.9), 
+#               (1, 3) : ('p', 0.1)}
+
+
+# nodes_with_class, connectivity = irm.data.generate.c_mixed_dist_block(SIDE_N, 
+#                                                                       class_conn, JITTER=0.1, rand_conn_prob=0.01)
+
+
+# class_conn = {(0, 1) : (1.0, 0.8, 0.1), 
+#               (1, 2) : (2.0, 0.4, 0.3), 
+#               (3, 2) : (0.7, 0.9, 0.2), 
+#               (3, 0) : (1.7, 0.7, 0.5), 
+#               (1, 3) : (0.5, 0.5, 0.2)}
+
+
+# nodes_with_class, connectivity = irm.data.generate.c_bump_dist_block(SIDE_N, 
+#                                                                       class_conn, JITTER=0.3, rand_conn_prob=0.01)
+
+
+
+class_conn = {(0, 1) : (1.0, 0.8), 
+              (1, 2) : (2.0, 0.6), 
+              (3, 2) : (0.7, 0.9), 
+              (2, 1) : (0.7, 0.8), 
+              (0, 3) : (1.7, 0.7)}
+
+
+nodes_with_class, connectivity = irm.data.generate.c_class_neighbors(SIDE_N, 
+                                                                     class_conn, JITTER=0.3, rand_conn_prob=0.01)
+
+
+
+CELL_N = len(connectivity)
+
+conn_and_dist = np.zeros((CELL_N, CELL_N), 
+                         dtype = [('link', np.uint8), 
+                                  ('distance', np.float32)])
+
+for ni, (ci, posi) in enumerate(nodes_with_class):
+    for nj, (cj, posj) in enumerate(nodes_with_class):
+        conn_and_dist[ni, nj]['link'] = connectivity[ni, nj]
[…]

### 2013-10-20 experiments/drosophila/diagnostic.py

@@ -7,38 +7,20 @@ from preprocess import *
 # plot spatial distribution of each cell type
 # plot area vs cell body distance
 
-@files("countmatrix.pickle", 
-       "adj_comp.pdf")
-def adj_mat_plot(infile, adj_plot):
+@files("data.all.pickle", 
+       "pre_post_dist_hist.pdf")
+def pre_post_dist_hist(infile, adj_plot):
 
-    d = pickle.load(open(infile, 'r'))
-    cell_ids = np.array(d['cell_ids'])
-    cell_id_pos = np.argsort(cell_ids)
-    conn = d['conn']
-    conn = conn[cell_id_pos]
-    conn = conn[:, cell_id_pos]
-
-    f = pylab.figure()
-
-    ax_conn = f.add_subplot(1, 1, 1)
+    df = pickle.load(open(infile, 'r'))
     
-    pos = []
-    sizes = []
-    for xi, x in enumerate(conn):
-        for yi, count in enumerate(x):
-            print count.shape
-            if count > 0:
-                pos.append((xi, yi))
-                sizes.append(count)
+    f = pylab.figure(figsize=(6, 12))
+    for i, coord in enumerate(['x', 'y', 'z']):
 
-    pos = np.array(pos)
-    ax_conn.scatter(pos[:, 0], pos[:, 1], 
-                    s = np.array(sizes)/1.5, edgecolor='none', alpha=0.5)
-    ax_conn.set_xlim(0, len(conn))
-    ax_conn.set_ylim(0, len(conn))
-    ax_conn.set_xlabel("presynaptic")
-    ax_conn.set_ylabel("postsynaptic")
-    ax_conn.set_title("Drosophila optic medulla")
+        ax_conn = f.add_subplot(3,1, i+1)
+        ax_conn.hist(df['pre.%s' % coord] - df['post.%s' % coord], bins=np.linspace(-2, 2, 100))
+        ax_conn.grid()
+        ax_conn.set_title("distance between pre and post %s" % coord)
+
     f.savefig(adj_plot)
 
 @files("celldata.pickle", 
@@ -69,5 +51,38 @@ def celldata_plot(infile, cell_locations):
             
     f.savefig(cell_locations)
 
+@files("countmatrix.pickle", 
+       "adj_comp.pdf")
+def adj_mat_plot(infile, adj_plot):
+
+    d = pickle.load(open(infile, 'r'))
[…]

### 2013-10-20 experiments/drosophila/pipeline.py

@@ -30,9 +30,15 @@ EXPERIMENTS = [
     ('drosophila.gp', 'fixed_10_100', 'nc_10'), 
     ('drosophila.bb', 'fixed_10_100', 'nc_10'), 
     ('drosophila.ld', 'fixed_10_100', 'nc_10'), 
-    #('drosophila.gp', 'fixed_100_200', 'anneal_slow_400'), 
-    #('drosophila.bb', 'fixed_100_200', 'anneal_slow_400'), 
-    #('drosophila.ld', 'fixed_100_200', 'anneal_slow_400'), 
+    ('drosophila.edp', 'fixed_10_100', 'nc_10'), 
+    ('drosophila.gp', 'fixed_100_200', 'anneal_slow_400'), 
+    ('drosophila.bb', 'fixed_100_200', 'anneal_slow_400'), 
+    ('drosophila.ld', 'fixed_100_200', 'anneal_slow_400'), 
+    ('drosophila.edp', 'fixed_100_200', 'anneal_slow_400'), 
+    ('drosophila.gp', 'fixed_100_200', 'anneal_glacial_1000'), 
+    ('drosophila.bb', 'fixed_100_200', 'anneal_glacial_1000'), 
+    ('drosophila.ld', 'fixed_100_200', 'anneal_glacial_1000'), 
+    ('drosophila.edp', 'fixed_100_200', 'anneal_glacial_1000'), 
 ]
 
 INIT_CONFIGS = {'fixed_10_100' : {'N' : 10, 
@@ -44,11 +50,15 @@ INIT_CONFIGS = {'fixed_10_100' : {'N' : 10,
                 
 
 slow_anneal = irm.runner.default_kernel_anneal()
+glacial_anneal = irm.runner.default_kernel_anneal(64.0, 800)
+
 default_nonconj = irm.runner.default_kernel_nonconj_config()
 
 KERNEL_CONFIGS = {
                   'anneal_slow_400' : {'ITERS' : 400, 
                                        'kernels' : slow_anneal},
+                  'anneal_glacial_1000' : {'ITERS' : 1000, 
+                                           'kernels' : glacial_anneal},
                   'nc_10' : {'ITERS' : 10, 
                              'kernels' : default_nonconj},
 
@@ -78,20 +88,37 @@ def create_count_matrix(infile, outfile):
                  'conn' : conn}, 
                 open(outfile, 'w'))
 
[…]

### 2013-10-20 experiments/drosophila/preprocess.py

@@ -38,17 +38,27 @@ def extract_useful_cells(infile, outfile):
 
     id_vc = df['pre.id'].value_counts()
     idx = np.array(np.array(map(type, id_vc.index.values)) == unicode)
-    cells_of_interest = id_vc[idx].index.values
-    print len(cells_of_interest)
+    pre_cells_of_interest = set(id_vc[idx].index.values)
 
-    s_post = set(df['post.id'])
-    s = s_post.intersection(cells_of_interest)
+    id_vc = df['post.id'].value_counts()
+    idx = np.array(np.array(map(type, id_vc.index.values)) == unicode)
+    post_cells_of_interest = set(id_vc[idx].index.values)
+
+    # average the synapse locations, because diagnostics suggests it doesn't matter
+    # much
+    for c in ['x', 'y', 'z']:
+        df[c] = (df['pre.' + c] + df['post.' + c]) / 2.0
+        del df['pre.' + c]
+        del df['post.' + c]
+
+    # now we only want synapses containing these cells
+    
+
+    s = pre_cells_of_interest.union(post_cells_of_interest)
     print "There are", len(s), "overlapping cells with unicode names" 
 
-    sub_df = df[df['pre.id'].isin(s)]
-    sub_df = sub_df[sub_df['post.id'].isin(s)]
-
-
+    sub_df = df[df['pre.id'].isin(s) & df['post.id'].isin(s)]
+    
     pickle.dump({'synapses' : sub_df, 
                  'cell_ids' : list(s)}, 
                 open(outfile, 'w'))
@@ -61,8 +71,8 @@ def per_cell_data(indata, outfile):
     post_mean = synapses.groupby('post.id').mean()
     pre_mean =  synapses.groupby('pre.id').mean()
 
-    df = pandas.concat([pre_mean[['pre.x', 'pre.y', 'pre.z']], post_mean[['post.x', 'post.y', 'post.z']]], axis=1)
-
+    df = pandas.concat([pre_mean[['x', 'y', 'z']], post_mean[['x', 'y', 'z']]], axis=1)
+    df.columns = ['pre.x', 'pre.y', 'pre.z', 'post.x', 'post.y', 'post.z']
     celltype_re = re.compile("(.+) (\d+)")
     df['type'] = [celltype_re.match(n).groups()[0] for n in df.index.values]
 

### 2013-10-20 experiments/mixing/pipeline.py

@@ -26,6 +26,8 @@ EXPERIMENTS = [('trivial', 'fixed_4_10', 'default200'),
                ('trivial_bump', 'fixed_4_10', 'default200'), 
                ('con_sparse', 'fixed_10_40', 'default200'), 
                ('con_sparse', 'fixed_10_40', 'default200seq'), 
+               ('con_sparse', 'fixed_10_40', 'default200seq'), 
+               ('con_sparse', 'fixed_10_40', 'default_anneal'), 
                # ('connmat0', 'fixed_10_40', 'default20'), 
                #('connmat0', 'fixed_10_40', 'default200'), 
                #('connmat0', 'fixed_10_40', 'nc_contmh_200'), 

### 2013-10-20 irm/runner.py

@@ -41,14 +41,14 @@ def add_relation_hp_grid_kernel(kernel_list, grids=None):
     return kl
 
 
-def default_kernel_anneal():
+def default_kernel_anneal(start_temp = 32.0, iterations=100):
     dk = default_kernel_nonconj_config()
     dk = add_domain_hp_grid_kernel(dk)
     dk = add_relation_hp_grid_kernel(dk)
     
-    return [('anneal', {'anneal_sched': {'start_temp' : 32.0, 
+    return [('anneal', {'anneal_sched': {'start_temp' : start_temp, 
                                          'stop_temp' : 1.0, 
-                                         'iterations' : 100}, 
+                                         'iterations' : iterations}, 
                          'subkernels': dk})]
 
               

### 2013-10-20 docs/techreport.tex
updates to figures

@@
 \subsection{Inference}
+
+We use Markov-chain Monte Carlo techniques to sample from the posterior 
+distribution of our model conditioned on the data. Inference is
+performed over cluster assignments, the per-component latent parameters, 
+and the per-relation, datatype-specific parameters.
+
 \subsubsection{Structural Inference}
+Algorithm 8 from radford neal \cite{nealalgo8}
 
@@
 Gridded gibbs sampling for hyperparameter inference
+
+\subsubsection{Annealing}
+For many examples of the data here we perform annealing.
 

### 2013-10-21 docs/techreport.tex
added figs

@@
 \usepackage{multicol}
+
+\DeclareGraphicsRule{.ai}{pdf}{.ai}{}
 
@@
 stochastic block models that
+
+\begin{figure}
+  \centering 
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{f1.raw.pdf}
+    \caption{Raw data}
+    \label{fig:gull}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.55\textwidth}
+    \includegraphics[width=\textwidth]{f1.sorted.pdf}
+    \caption{Raw data}
+    \label{fig:gull}
+  \end{subfigure}
+\end{figure}
 
@@
 
-\subsubsection{Normal Distance Fixed Width}
+\begin{figure}
+  \centering 
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=0.3\textwidth]{logisticdistance.ai}
+    \caption{Raw data}
+    \label{fig:gull}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=0.3\textwidth]{exponentialdistancepoisson.ai}
+    \caption{Raw data}
+    \label{fig:gull}
+  \end{subfigure}
+
+\end{figure}
+
 This likeilihood model uses a logistic transform from distance to the
@@
 We don't do hyperparameter inference on $p_{min}$ and $p_{max}$.
+
+\subsubsection{Normal Distance Fixed witdh}
 

### 2013-10-21 docs/techreport.tex
additional diagnostic plots

@@
 \usepackage{amssymb}
-\usepackage{biblatex}
+\usepackage[citestyle=numeric-comp,url=false,hyperref=auto,isbn=false]{biblatex}
+\addbibresource{connectomics motif.bib}
+\addbibresource{connectomics motif-relational models.bib}
 
@@
 
-We adopt the terminology from \autocite{Kemp}. A domain $T_n$ is a
+We adopt the terminology from \autocite{Kemp2006a}. A domain $T_n$ is a
 collection of entities $\{e^t_i | i \in 0 \dots N^t -1 \} $ that share
@@
   \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=0.3\textwidth]{logisticdistance.ai}
-    \caption{Raw data}
+    \includegraphics[height=1.5in]{logisticdistance.ai}
+    \caption{Logistic-Distance Bernoulli Likelihood}
     \label{fig:gull}
   \end{subfigure}
-
   \begin{subfigure}[b]{0.4\textwidth}
-    \includegraphics[width=0.3\textwidth]{exponentialdistancepoisson.ai}
-    \caption{Raw data}
+    \includegraphics[height=1.5in]{exponentialdistancepoisson.ai}
+    \caption{Exponential-Distance Poisson Likelihood}
     \label{fig:gull}
@@
 \subsubsection{Exponential Distance Poisson}
-Some of our observations are count-valued.
+For distance-based models with count-based observations, we use an exponential
 
@@
 \subsubsection{Structural Inference}
-Algorithm 8 from radford neal \cite{nealalgo8}
+Algorithm 8 from radford neal \autocite{Neal2000}
 
 \subsubsection{Per-component parameters in nonconjugate case}
-Slice sample on latent parameters
+Slice sample on latent parameters \autocite{Neal2003}
 
@@
 Dense serial electron microscopy of a $VOLUMNE$ in the mouse
-retina by \autocite{MouseReinta} yielded a listing of places where
+retina by \autocite{Helmstaedter2013} yielded a listing of places where
 neurons come into contact. There were $number$ cells originally, whcih
 were refined through various classification schemes. We selected the
-$NUMBER$ for which the location of the soma could be reconstruted from
+$950$ for which the location of the soma could be reconstruted from
 the provided cell plots (soma locations were not provided by the
@@
 \subsection{Drosophila medulla}
-\cite{DrosophilaConnectome}
+\autocite{Takemura2013}
 No cell bodies -- what to use for distance? 
@@
 
+\printbibliography
+
 \end{document}

### 2013-10-21 experiments/mouseretina/diagnostic_plots.py

@@ -211,6 +211,78 @@ def plot_example_cells((pos_file, synapse_file), output_files):
 
         f.savefig(output_file)
 
+@files(["type_metadata.pickle", "soma.positions.pickle", "xlsxdata.pickle"], 
+       ['adjmat.byclass.png'])
+def plot_adjmat_byclass((type_file, pos_file, xlsxdata_file), (adj_mat_plot,)):
+    soma_pos = pickle.load(open(pos_file, 'r'))
+    type_metadata = pickle.load(open(type_file, 'r'))['type_metadata']
+    xlsdata = pickle.load(open(xlsxdata_file, 'r'))
+    types = xlsdata['types']
+    area_mat = xlsdata['area_mat']
+    pos_vec = soma_pos['pos_vec']
+    CELL_N = len(pos_vec)
+    print "CELL_N=", CELL_N
+
+
+    f = pylab.figure(figsize=(12, 12))
+    ax = f.add_subplot(1, 1, 1)
+    
+    x_pos = []
+    y_pos = []
+    s = []
+    for i in range(CELL_N):
+        for j in range(CELL_N):
+            a = area_mat[i, j]
+            if a > 1e-6:
+                x_pos.append(i)
+                y_pos.append(j)
+                s.append(a)
+    ax.scatter(x_pos, y_pos, s=s, edgecolor='none', alpha=0.5, c='k')
+    ax.set_xlim(0, CELL_N)
+    ax.set_ylim(CELL_N, 0)
+    ax.set_title("mouse retina connectivity matrix")
+    f.tight_layout()
+    f.savefig(adj_mat_plot, dpi=200)
+
+
+@files(["type_metadata.pickle", "soma.positions.pickle", "xlsxdata.pickle"], 
+       ['adjmat.byz.png'])
+def plot_adjmat_byz((type_file, pos_file, xlsxdata_file), (adj_mat_plot,)):
+    soma_pos = pickle.load(open(pos_file, 'r'))
+    type_metadata = pickle.load(open(type_file, 'r'))['type_metadata']
+    xlsdata = pickle.load(open(xlsxdata_file, 'r'))
+    types = xlsdata['types']
+    area_mat = xlsdata['area_mat']
+    pos_vec = soma_pos['pos_vec']
+    CELL_N = len(pos_vec)
+    print "CELL_N=", CELL_N
+    ai =  np.argsort(pos_vec[:, 2]).flatten()
+    area_mat = area_mat[ai]
[…]

### 2013-10-21 docs/techreport.tex
added output plots cleaned up plots for generation

@@
 
+\begin{figure}
+  \centering 
+  \includegraphics[width=0.5\textwidth]{mouseretina/adjmat.byz.png}
+  \caption{Mouse connectome for the 950 cells with known soma positions, ordered by the position of the soma along the $z$ axis.
+The size of each point is proportional to the total area of contact between the cells.}
+  \label{fig:mouse_adj}
+\end{figure}
+
 \subsection{C. elegans}
@@
 
+279 neurons (non-phyrnx)
+
+\begin{figure}
+  \centering 
+  \includegraphics[width=0.5\textwidth]{celegans/data.pos.pdf}
+  \caption{Connectome for the 279 non-pharynx cells in \textit{c.
+elegans}.
+Cells are ordered by their position along the body axis.
+Blue dots are chemical synapses, red dots are electrical synapses, and the area of each dot is equivalent to the number of synaptic contacts between the cells.
+Note that the chemical synapses (blue) are directional, thus the red dots form a symmetric matrix}
+  \label{fig:celegans_adj}
+\end{figure}
+
 \subsection{Drosophila medulla}
@@
 what the hell is a cell
+357 cells by the time we get through it
+
+\begin{figure}
+  \centering 
+  \includegraphics[width=0.5\textwidth]{drosophila/adj_comp.pdf}
+  \caption{Connectome for the 357 cells in \textit{drosophila} medulla.
+Cells are ordered by the order they appear in the data file.
+Dot area is proportional to the number of synapses; data are directional.}
+  \label{fig:drosophila_adj}
+\end{figure}
 

### 2013-10-21 experiments/celegans/diagnostic.py

@@ -79,8 +79,9 @@ def plot_adjmat(infile, plot_pos_filename):
         ax.scatter(elec_points[:, 0], elec_points[:, 1], c='r', 
                    edgecolor='none', s= elec_points[:, 2]*3, alpha=0.5)
         ax.set_xlim(0, NEURON_N)
-        ax.set_ylim(0, NEURON_N)
+        ax.set_ylim(NEURON_N, 0)
         ax.set_title("c. elegans. herm. somatic connectome adj. matrix %s" % sortby)
+        f.tight_layout()
         f.savefig(outfile)
 
 @files("data.processed.pickle", ['data.all.class.pdf', 'data.big.class.pdf'])

### 2013-10-21 experiments/drosophila/diagnostic.py

@@ -59,10 +59,11 @@ def adj_mat_plot(infile, adj_plot):
     cell_ids = np.array(d['cell_ids'])
     cell_id_pos = np.argsort(cell_ids)
     conn = d['conn']
+    print len(conn)
     conn = conn[cell_id_pos]
     conn = conn[:, cell_id_pos]
 
-    f = pylab.figure()
+    f = pylab.figure(figsize=(6,6))
 
     ax_conn = f.add_subplot(1, 1, 1)
     
@@ -75,13 +76,14 @@ def adj_mat_plot(infile, adj_plot):
                 sizes.append(count)
 
     pos = np.array(pos)
-    ax_conn.scatter(pos[:, 0], pos[:, 1], 
-                    s = np.array(sizes)/1.5, edgecolor='none', alpha=0.5)
+    ax_conn.scatter(pos[:, 1], pos[:, 0], 
+                    s = np.array(sizes)*3, edgecolor='none', alpha=0.3, c='k')
     ax_conn.set_xlim(0, len(conn))
-    ax_conn.set_ylim(0, len(conn))
-    ax_conn.set_xlabel("presynaptic")
-    ax_conn.set_ylabel("postsynaptic")
+    ax_conn.set_ylim(len(conn), 0)
+    ax_conn.set_ylabel("presynaptic")
+    ax_conn.set_xlabel("postsynaptic")
     ax_conn.set_title("Drosophila optic medulla")
+    f.tight_layout()
     f.savefig(adj_plot)
 
 if __name__ == "__main__":

### 2013-10-21 docs/techreport.tex
updates to the indicated plots tech report progressing

@@
 \usepackage[english,american]{babel}
+\usepackage{graphicx}
 \usepackage{todonotes}
@@
 \usepackage{subcaption}
-\usepackage{multicol}
+
+\usepackage[firstpage]{draftwatermark}
+\SetWatermarkScale{4}
 
@@
 \subsubsection{Structural Inference}
-Algorithm 8 from radford neal \autocite{Neal2000}
+Algorithm 8 from radford neal \autocite{Neal2000} for dirichlet process
+inference.
 
@@
 Gridded gibbs sampling for hyperparameter inference
+We set the grid by hand (empirical bayes)
 
 \subsubsection{Annealing}
-For many examples of the data here we perform annealing.
+For many examples of the data here we perform annealing on the global likelihood during the burn-in period.
 
@@
 the provided cell plots (soma locations were not provided by the
-study's authors in machien-readble form).
+study's authors in machine-readble form).
 
@@
 The size of each point is proportional to the total area of contact between the cells.}
-  \label{fig:mouse_adj}
+  \label{fig:mouseretina:adj}
+\end{figure}
+
+\subsubsection{Results}
+
+\begin{figure}
+  \centering 
+  \includegraphics[width=\textwidth]{mouseretina/retina.1.0.ld.0.0.data-fixed_20_100-anneal_slow_400.0.clusters.pdf}
+  \caption{Identified clusters from the mouse retina connectome using a Logistic-Distance-Bernoulli likelihood.
+Each column of plots is a cluster, wider plots for clusters with more cells.
+\textbf{first row} of plots compares the identified cluster to the experimenter-provided latent cell types, of which there were 74.
+Each bar is the fraction (between 0 and 1) of the cells of that type that are in this cluster.
+Note that the experimentors ordered their latent cell types, so cell type 60 is closer to 61 than 65 (see main text).
+Colors reflect the macro cell types (red = retinal ganglion cells, blue are amacrine, green are bipolar, and black are other.
+\textbf{Second row} is the spatial distribution of soma (blue) and synapses (red) for this cluster along a direction approximately normal to the retina surface.
+\textbf{Third row} is the spatial distribution of the soma (blue dots) and synapses (red dots) in this cluster, across the retina surface.}
+  \label{fig:mouseretina:clusters}
+\end{figure}
+
+\begin{figure}
+  \centering 
+  \begin{subfigure}[b]{0.46\textwidth}
+    \includegraphics[width=\textwidth, page=1]{mouseretina/retina.1.0.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}
+    \caption{Clustered latent connectivity matrix.}
+    \label{fig:gull}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.46\textwidth}
+    \includegraphics[width=\textwidth, page=2]{mouseretina/retina.1.0.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}
+    \caption{Latent cluster parameters. }
+    \label{fig:gull}
+  \end{subfigure}
+  \caption{Latent structure for mouse retina connectome clusters above.
+The latent cluster parameters in (b.) show (blue) the empirical distribution of connectivity for this cluster as a function of distance, and (red) the fit logistic curve.
+The vertical black line is the value of $\mu$.}
+  \label{fig:mouseretina:latent}
 \end{figure}
@@
 \subsection{C. elegans}
-Connectome across several animals traced out by sidney brenner back in the day.
-Small number of cells
-position is linear along body axis
-cell divisions are really precise
 
-only look at hermaphrdite, ignore phyranx.
-
-Have directionality, chemical vs electrical, number
-
-279 neurons (non-phyrnx)
+The connectome of the nematode Caenorhabditis elegans was traced by
+hand from serial electron micrographs from Sidney Brenner's laboratory
+in 1986 \autocite{White1986}. \textit{C. elegans} exhibits extremely
+stereotypic neural development, with each hermaphrodite animal having
+exactly 302 neurons, and the synaptic connectivity being 75\%
+identical \todo{where did this number come from} across animals. Here we
+examine 279 nonpharyngeal neurons with 6393 chemical synapes and 890 electrical junctions originally cleaned up in \autocite{Chen2006}. The adjacency matrix is shown in figure~\ref{fig:celegans:adj}
 
@@
   \includegraphics[width=0.5\textwidth]{celegans/data.pos.pdf}
-  \caption{Connectome for the 279 non-pharynx cells in \textit{c.
+  \caption{Connectome for the 279 nonpharyngeal cells in \textit{c.
 elegans}.
@@
 Note that the chemical synapses (blue) are directional, thus the red dots form a symmetric matrix}
-  \label{fig:celegans_adj}
+  \label{fig:celegans:adj}
 \end{figure}
+
+The structure of the \textit{c. elegans} nervous system differs considerably 
+from the primary sensory connectomes we examined:
+
+\begin{itemize}
+\item Lots of ganglia with global projections
+\item Many of Brenner's originally-identified cell types contain only two member cells, one on each side of the animal. 
+\item We have both electrical and chemical synapses, where the chemical synapses are directional. 
+\item Our position data only indicates the position along the body axis (on the interval $[0, 1]$. 
+\end{itemize}
+
+The neurons can be broadly classified into ``sensory'', ``motor'', and
+``interneurons''. In addition, there are several broad classes of both
+sensory and motor neurons containing up to twelve cells.
+Figure ~\ref{fig:celegans:classestruth} shows the locations the cells of the biggest classes.
+Note the anterior concentration of sensory and internurons in contrast to the broad spatial distribution of the motor neurons.
+
+\begin{figure}
+  \centering 
+  \includegraphics[width=0.8\textwidth]{celegans/class.positions.pdf}
+  \caption{Positions of \textit{C.
+elegans} neurons grouped by cell type along the anterior-posterior axis.
+The cell type is listed to the right.
+Blue cells are motor neurons, green are sensory, and grey are interneurons.
+}
+  \label{fig:celegans:classestruth}
+\end{figure}
+
+\subsubsection{results}
+They are crappy.
+
+\todo{Should we try to include the neuromuscular and sensory junctions? The data exists from wormatlas.org}
 
@@
   \centering 
-  \includegraphics[width=0.5\textwidth]{drosophila/adj_comp.pdf}
+  \includegraphics[width=0.7\textwidth]{drosophila/adj_comp.pdf}
   \caption{Connectome for the 357 cells in \textit{drosophila} medulla.
@@
 \end{figure}
+
+\subsubsection{results}
 
@@
 
+\subsubsection{results}
+
 \subsection{Zipcode data}
 
-This data is from career builder. It contains who applied to what jobs. This is our first non-graph example of a T1xT2 relation.
+This data is from career builder. It contains who applied to what
+jobs. This is our first non-graph example of a T1xT2 relation.
 

### 2013-10-21 experiments/mouseretina/pipeline.py

@@ -481,7 +481,7 @@ def plot_scores_z(exp_results, (plot_latent_filename, plot_truth_filename)):
                                  plot_truth_filename)
 
 @transform(get_results, suffix(".samples"), 
-           [(".%d.clusters.pdf" % d, ".%d.latent.pdf" % d )  for d in range(3)])
+           [(".%d.clusters.pdf" % d, ".%d.latent.pdf" % d )  for d in range(1)])
 def plot_best_cluster_latent(exp_results, 
                      out_filenames):
 

### 2013-10-21 experiments/mouseretina/util.py

@@ -21,7 +21,7 @@ def plot_cluster_properties(assignments, true_cell_types,
     """
 
 
-    f = pylab.figure(figsize=(16, 9))
+    f = pylab.figure(figsize=(18, 12))
     X_MIN = np.min(pos_vec[:, 0])
     X_MAX = np.max(pos_vec[:, 0])
 
@@ -54,14 +54,20 @@ def plot_cluster_properties(assignments, true_cell_types,
         ax_x.plot(syn_hist, bins[:-1], c='r', 
                   linewidth=1, alpha=0.5)
         ax_yz.scatter(syn_th['y'], syn_th['z'], c='r', s=1, edgecolor='none')
+        if (col_i > 0):
+            ax_yz.set_xticklabels([])
+            ax_yz.set_yticklabels([])
 
+            ax_x.set_xticklabels([])
+            ax_x.set_yticklabels([])
 
     irm.plot.plot_purity_hists_h(f, assignments, true_cell_types, extra_rows=2, 
                                  extra_row_func=extra_plot_func, 
                                  thold = cluster_thold, class_colors = class_colors)
     #f.tight_layout()
-    f.suptitle(output_filename)
-    f.savefig(output_filename)
+    #f.suptitle(output_filename)
+    
+    f.savefig(output_filename, bbox_inches='tight')
 
 def reorder_synapse_ids(synapse_df, cell_id_permutation):
     """

### 2013-10-21 irm/experiments.py

@@ -85,13 +85,18 @@ def plot_latent(latent, dist_matrix,
 
     a = np.array(latent['domains']['d1']['assignment'])
 
-    f = pylab.figure(figsize= (24, 26))
+    if ground_truth_assign == None:
+        f = pylab.figure(figsize= (24, 24))
+        gs = gridspec.GridSpec(1, 1)
+        ax = pylab.subplot(gs[0])
+    else:
+        f = pylab.figure(figsize= (24, 26))
     
-    gs = gridspec.GridSpec(2, 1, height_ratios=[1,12])
-    
-    
-    ax = pylab.subplot(gs[1])
-    ax_types = pylab.subplot(gs[0])
+        gs = gridspec.GridSpec(2, 1, height_ratios=[1,12])
+        ax_types = pylab.subplot(gs[0])        
+        ax = pylab.subplot(gs[1])
+
+
     print "plot_t1t1_latent"
     ai = irm.plot.plot_t1t1_latent(ax, dist_matrix['link'], a)
 
@@ -123,7 +128,8 @@ def plot_latent(latent, dist_matrix,
                               latent['relations']['R1']['hps'], 
                               MAX_DIST=PLOT_MAX_DIST, model=model, 
                               MAX_CLASSES=MAX_CLASSES   )
-    f.suptitle(latent_filename)
+
+    f2.tight_layout()
     f2.savefig(pp, format='pdf')
 
     pp.close()

### 2013-10-21 docs/techreport.tex
updates including the c. elegans results.

@@
 
-\subsubsection{results}
+\subsubsection{Results}
+We used two relations, one for chemical and one for electrical synapses. 
+We binarize the synapse counts at greater than zero.
+
+\begin{figure}
+  \centering 
+  \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{celegans/celegans.2r.ld.00.data-fixed_100_200-anneal_slow_400.0.clusters.pdf}
+    \caption{Logistic-Distance Bernoulli}
+    \label{fig:celegans:clusters:ld}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{celegans/celegans.2r.ndfw.00.data-fixed_100_200-anneal_slow_400.0.clusters.pdf}
+    \caption{Normal-Distance Fixed-Width}
+    \label{fig:celegans:clusters:ndfw}
+  \end{subfigure}
+  \caption{C. elegans clusters. }
+  \label{fig:celegans:clusters}
+\end{figure}
+
 They are crappy.
@@
 \todo{Should we try to include the neuromuscular and sensory junctions? The data exists from wormatlas.org}
+
+\todo{should we use the exponentail-distance-count likelihood?}
 

### 2013-10-21 experiments/celegans/pipeline.py

@@ -538,7 +538,7 @@ def plot_best_cluster(exp_results,
         #                             PLOT_MAX_DIST=1.2)
         a = irm.util.canonicalize_assignment(sample_latent['domains']['d1']['assignment'])
         fig = plot_clusters_pretty_figure(a, neurons, no)
-        fig.savefig(cluster_fname)
+        fig.savefig(cluster_fname, bbox_inches='tight')
 
 # @transform(get_results, suffix(".samples"), [".clusters.html"])
 # def cluster_interpretation(exp_results, (output_filename,)):

### 2013-10-21 docs/techreport.tex
updates to plots for diagnostics

@@
   \centering 
-  \includegraphics[width=0.8\textwidth]{celegans/class.positions.pdf}
+  \includegraphics[width=0.7\textwidth]{celegans/class.positions.pdf}
   \caption{Positions of \textit{C.
@@
 \subsubsection{Results}
-We used two relations, one for chemical and one for electrical synapses. 
-We binarize the synapse counts at greater than zero.
+We model \textit{C.
+elegans} using two $T1\times T1$ binary relations, one for the chemical synapses and one for the electrical synapses.
+Two cells are connected if they have at least one synapse between them, giving rise to a binary matrix for each synapse type.
+We see the results of two likelihood models in figure~\ref{fig:celegans:clusters}.
+The logistic-distance Bernoulli likelihood segements the sensory and interneurons from the spatially-distributed motor neurons, whereas the normal-distance fixed-width likelihood \todo{Write about this; adopt consistent naming} creates an odd collection of clusters.
 
@@
   \centering 
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.42\textwidth}
     \includegraphics[width=\textwidth]{celegans/celegans.2r.ld.00.data-fixed_100_200-anneal_slow_400.0.clusters.pdf}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{subfigure}[b]{0.47\textwidth}
     \includegraphics[width=\textwidth]{celegans/celegans.2r.ndfw.00.data-fixed_100_200-anneal_slow_400.0.clusters.pdf}
@@
   \end{subfigure}
-  \caption{C. elegans clusters. }
+  \caption{C.
+elegans clusters for two different likelihoods, showing the clustering of the largest classically-known cell types.
+Green cells are sensory, grey are interneurons, and blue are motor neurons.
+The top row shows the cells in the cluster and where they are spatially; the bottom shows cluster purity.
+}
   \label{fig:celegans:clusters}
@@
 
-They are crappy.
+\todo[inline]{Should we try to include the neuromuscular and sensory junctions? The data exists from wormatlas.org}
 
-\todo{Should we try to include the neuromuscular and sensory junctions? The data exists from wormatlas.org}
+\todo[inline]{should we use the exponentail-distance-count likelihood?}
 
-\todo{should we use the exponentail-distance-count likelihood?}
-
-\subsection{Drosophila medulla}
-\autocite{Takemura2013}
-No cell bodies -- what to use for distance? 
-Human annotation
-what the hell is a cell
-357 cells by the time we get through it
+\todo[inline]{Better understanding of how we model ganglia-like structures}
+\subsection{Drosophila optic medulla}
+Serial electron microscopy was used to reconstruct several columns from the Drosophila optic medulla \autocite{Takemura2013}.
+Experimenters were able to measure the physical location of the pre- and post-synaptic side of each synapse in the neuropil, and thus the directionality.
+The neural cell bodies in the cortex surround the neuropil and were not measured due to computational infesability \autocite{MeinertzhagenPersonalComm}, thus we don't know the spatial location of the soma.
 
@@
   \centering 
-  \includegraphics[width=0.7\textwidth]{drosophila/adj_comp.pdf}
-  \caption{Connectome for the 357 cells in \textit{drosophila} medulla.
+  \begin{subfigure}[b]{0.55\textwidth}
+    \includegraphics[width=\textwidth]{drosophila/adj_comp.pdf}
+    \caption{Drosophila optic medulla connectome}
+    \label{fig:drosophila:connectome}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{drosophila/pre_post_dist_hist.pdf}
+    \caption{Pre and post-synaptic distance}
+    \label{fig:drosophila:distprepost}
+  \end{subfigure}
+  \caption{a.) Connectome for the 357 cells in \textit{drosophila} medulla.
 Cells are ordered by the order they appear in the data file.
-Dot area is proportional to the number of synapses; data are directional.}
+Dot area is proportional to the number of synapses; data are directional.
+b.) Histogram of spatial distance between pre- and post-synaptic sites.}
   \label{fig:drosophila_adj}
@@
 
-\subsubsection{results}
+\subsubsection{Results}
+From the provided spreadsheet of synapses we extract all synapses with cells with non-numeric IDs (357). We reduced the location of every synapse to the average of its pre and post-synaptic sites.
+
+\todo[inline]{Is this the right thing to use for spatial location?}
 

### 2013-10-21 experiments/celegans/diagnostic.py

@@ -35,12 +35,20 @@ def plot_positions(infile, outfile):
                 c = "#AAAAAA"
             ax.plot([0, 1], [plotted, plotted], c='k', alpha=0.5, linewidth=1)
             ax.scatter(g['soma_pos'], np.ones(N) * plotted, 
-                       c=c)
-            ax.text(1.02, plotted, gi, fontsize=8)
+                       c=c, s=30)
+            ax.text(1.01, plotted - 0.15, gi, fontsize=10)
             plotted +=1
+
     ax.set_xlim(-0.05, 1.1)
+    ax.set_xlabel("position along anterior-posterior axis")
     ax.set_ylim(-1, plotted)
+    ax.set_yticks([])
     ax.set_title('c elegans cell class positions')
+    ax.spines['right'].set_visible(False)
+    ax.spines['top'].set_visible(False)
+    ax.spines['left'].set_visible(False)
+    ax.xaxis.set_ticks_position('bottom')
+
     f.savefig(outfile)
     
 @files('data.processed.pickle', 'data.pos.pdf')

### 2013-10-21 experiments/drosophila/diagnostic.py

@@ -13,14 +13,15 @@ def pre_post_dist_hist(infile, adj_plot):
 
     df = pickle.load(open(infile, 'r'))
     
-    f = pylab.figure(figsize=(6, 12))
+    f = pylab.figure(figsize=(6, 6))
     for i, coord in enumerate(['x', 'y', 'z']):
 
         ax_conn = f.add_subplot(3,1, i+1)
-        ax_conn.hist(df['pre.%s' % coord] - df['post.%s' % coord], bins=np.linspace(-2, 2, 100))
+        ax_conn.hist(df['pre.%s' % coord] - df['post.%s' % coord], bins=np.linspace(-2, 2, 100), normed=True)
         ax_conn.grid()
         ax_conn.set_title("distance between pre and post %s" % coord)
-
+        ax_conn.set_xlabel("um")
+    f.tight_layout()
     f.savefig(adj_plot)
 
 @files("celldata.pickle", 

### 2013-10-22 docs/techreport.tex
updates to plots, TR getting closer

@@
 
-\subsection{infinite Spatial Relational Models}
+\subsection{Infinite Spatial Relational Models}
 
@@
 
-\subsubsection{Arbitrary discriminative funtions}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{comparison.pdf}
+  \caption{Posterior distributions for the number of found latent classes.
+Synthetic datasets with spatially-localized connectivity and one (left) and t two (right) true latent classes were created.
+The regular beta-bernoulli infinite relational model (top) over-segements the data dramatically, whereas the spatial-relational model with the logistic-distance bernoulli likelihood recovers the true number of classes}
+\end{figure}
+
+\subsubsection{Arbitrary discriminative functions}
 We can extend models of this type
@@
 For distance-based models with count-based observations, we use an exponential
+
+\todo[inline]{Poisson is a pain because of linked mean/var}
 
@@
 Experimenters were able to measure the physical location of the pre- and post-synaptic side of each synapse in the neuropil, and thus the directionality.
-The neural cell bodies in the cortex surround the neuropil and were not measured due to computational infesability \autocite{MeinertzhagenPersonalComm}, thus we don't know the spatial location of the soma.
+The neural cell bodies in the cortex surround the neuropil and were not measured due to technique limitations \autocite{MeinertzhagenPersonalComm}, thus we don't know the spatial location of the soma.
 
@@
 \subsubsection{Results}
-From the provided spreadsheet of synapses we extract all synapses with cells with non-numeric IDs (357). We reduced the location of every synapse to the average of its pre and post-synaptic sites.
 
-\todo[inline]{Is this the right thing to use for spatial location?}
+From the provided spreadsheet of synapses we extract all synapses for cells with non-numeric IDs (357).
+We reduced the location of every synapse to the average of its pre and post-synaptic sites, as the average distance is small (figure~\ref{fig:drosophila:distprepost}).
+For a given postsynaptic cell, we average the spatial location of all of its postsynaptic sites, and assign this point as the cell's location.
+The resulting clusters from using the exponential-distance-Poisson model with synapse count observations is in figure~\ref{fig:drosophila:clusters_edp}.
+The clusterings are not obvious, and don't reflect any particular clean division of their existing labeled subtypes, although the block structure appears pretty solid.
+
+Next steps are:
+
+\begin{itemize}
+  \item Is the centroid of the presynaptic inputs the best point to use for the cell's ``distance'' calculations ? 
+  \item I've tried both EDP and LD models for count and binary data respectively, without any obvious benefits
+  \item How well do we really understand this system?
+My understanding both of the neuroanatomy and the data acquisition process for the mouse retina really guided my modelling efforts.
+I don't have that here, and need to read \autocite{Fischbach1989} to understand cell types.
+  \item Did the experimenters measure just a ``single'' column completely and then whatever happened to be in nearby columns that connected to this focus column?
+If so, we have complete observation data for only one microcircuit.
+    \item Are we sure this isn't just the result of some preprocessing bug? I need to run synthetic data through this pipeline.
+
+    \end{itemize}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{drosophila/drosophila.edp.data-fixed_100_200-anneal_slow_400.0.clusters.pdf}
+  \caption{}
+  \label{fig:drosophila:clusters_edp}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{drosophila/drosophila.edp.data-fixed_100_200-anneal_slow_400.0.latent.pdf}
+  \caption{}
+  \label{fig:drosophila:latent_edp}
+\end{figure}
 
@@
 
+\begin{figure}
+  \centering 
+  \begin{subfigure}[b]{0.43\textwidth}
+    \includegraphics[width=\textwidth]{mos6502/6502_shrunk.png}
+    \caption{MOS 6502 die}
+    \label{fig:mos6502:die}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.50\textwidth}
+    \includegraphics[width=\textwidth]{mos6502/transistors.pdf}
+    \caption{Identified transistor locations}
+    \label{fig:mos6502:transistors}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{mos6502/all.adj.png}
+    \caption{Adjacency matrix for entire IC}
+    \label{fig:mos6502:adj}
+  \end{subfigure}
+  \caption{}
+  \label{fig:mos6502:adjj}
+\end{figure}
+
+\subsubsection{Results}
+
 We extracted this data at the per-transistor level from the online
-github project, resulting in $NUMBER$ transistors. Each transistor has
-three terminals: $g$, $c_1$, and $c2$. Electronic circuits are fundamentally
-undirected at the physical leve.
+github project, resulting in $NUMBER$ transistors. Each MOS transistor
+has three terminals: $g$, $c_1$, and $c2$. The resulting data is then
+a bipartate graph between the terminals of the transistors and various
+voltage ``nodes'', the wires that connect them. At the physical level
+electronic circuits are fundamentally undirected, and all terminals
+connected to a node will be at the same voltage. Ideally we'd be
+operating at a higher level of abstraction with digital elements (logic gates,
+registers) with more clearly-defined inputs and outputs, but this
+was the only data set available.
 
-Remove CLK, VCC/VSS.
+For our encoding, the entities are transistors, and the relation is
+true if any terminal of the transistor connects to any terminal of any
+other transistor. This results in an incredibly densely-connected
+graph, as almost all transistors are connected to very common nodes
+like ``power'', ``ground'', and the global clock, which we remove
+prior to analysis. We look at two subsets of the total chip, the
+instruction decoding region and the region consisting of the ``X'',
+``Y'', and ``S'' registers.
 
-This would work better up one level of abstraction.
+\begin{figure}[h]
+\centering
+\begin{subfigure}[b]{0.43\textwidth}
+  \includegraphics[width=\textwidth]{mos6502/mos6502.all.decode.ld.data-fixed_20_200-anneal_slow_400.0.latent.pdf}
+  \caption{MOS 6502 Instruction Decode Latent}
+\end{subfigure}
 
-Challenges: not directed. Three terminals. Dense connections.
+\begin{subfigure}[b]{0.43\textwidth}
+  \includegraphics[width=\textwidth]{mos6502/mos6502.all.xysregs.ld.data-fixed_20_200-anneal_slow_400.0.latent.pdf}
+  \caption{MOS 6502 XYS Register Latent}
+\end{subfigure}
+\caption{}
+\label{}
+\end{figure}
 
-THIS SHOULD WORK
+\todo[inline]{Show where those regions on the chip are}
+\todo[inline]{Run on entire chip}
+\todo[inline]{Other ways of encoding (multiple relations?) }
+\todo[inline]{Run on entire chip with annealing}
+\todo[inline]{interpret the damn latent structure}
 
-\subsubsection{results}
-
-\subsection{Zipcode data}
+\subsection{Job application data}
 

### 2013-10-22 experiments/drosophila/pipeline.py

@@ -35,10 +35,10 @@ EXPERIMENTS = [
     ('drosophila.bb', 'fixed_100_200', 'anneal_slow_400'), 
     ('drosophila.ld', 'fixed_100_200', 'anneal_slow_400'), 
     ('drosophila.edp', 'fixed_100_200', 'anneal_slow_400'), 
-    ('drosophila.gp', 'fixed_100_200', 'anneal_glacial_1000'), 
-    ('drosophila.bb', 'fixed_100_200', 'anneal_glacial_1000'), 
-    ('drosophila.ld', 'fixed_100_200', 'anneal_glacial_1000'), 
-    ('drosophila.edp', 'fixed_100_200', 'anneal_glacial_1000'), 
+    #('drosophila.gp', 'fixed_100_200', 'anneal_glacial_1000'), 
+    #('drosophila.bb', 'fixed_100_200', 'anneal_glacial_1000'), 
+    #('drosophila.ld', 'fixed_100_200', 'anneal_glacial_1000'), 
+    #('drosophila.edp', 'fixed_100_200', 'anneal_glacial_1000'), 
 ]
 
 INIT_CONFIGS = {'fixed_10_100' : {'N' : 10, 
@@ -67,12 +67,19 @@ KERNEL_CONFIGS = {
 pickle.dump(default_nonconj, open("kernel.config", 'w'))
 
 
-@files('synapses.pickle', 'countmatrix.pickle')
-def create_count_matrix(infile, outfile):
+@files(['synapses.pickle', 'celldata.pickle'],  'countmatrix.pickle')
+def create_count_matrix((synapse_infile, celldata_infile), outfile):
     
-    data = pickle.load(open(infile, 'r'))
+    data = pickle.load(open(synapse_infile, 'r'))
     synapse_df = data['synapses']
-    cell_ids = data['cell_ids']
+
+    celldata_df = pickle.load(open(celldata_infile, 'r'))['celldata']
+
+    celldata_df = celldata_df[np.isfinite(celldata_df['post.x'])] # identify missing entities
+
+    cell_ids = celldata_df.index.values
+
+    synapse_df = synapse_df[synapse_df['post.id'].isin(cell_ids) & synapse_df['pre.id'].isin(cell_ids)]
     
     
     # create the matrix
@@ -83,6 +90,8 @@ def create_count_matrix(infile, outfile):
         pre_idx = name_to_pos[row['pre.id']]
[…]

### 2013-10-22 experiments/mos6502/preprocess.py

@@ -139,8 +139,15 @@ def plot_transistors(input_file, output_file):
 
     f = pylab.figure()
     ax = f.add_subplot(1, 1, 1)
-    ax.scatter(tfdf['x'], tfdf['y'], s=5, edgecolor='none', alpha=0.5)
-    f.savefig(output_file)
+    ax.scatter(tfdf['x'], tfdf['y'], s=5, edgecolor='none', c='k', alpha=0.5)
+    ax.set_xlim(0, 9000)
+    ax.set_ylim(0, 10000)
+    ax.set_xlabel("um")
+    ax.set_ylabel("um")
+
+    ax.set_aspect(1.0)
+    f.tight_layout()
+    f.savefig(output_file, bbox_inches='tight')
 
 @files(load_data, "graph.pickle")
 def create_raw_graph(input_file, output_file):
@@ -457,7 +464,8 @@ def plot_adj_matrix(infile, (adj_matrix_outfile, distance_outfile)):
     # ax_dist.imshow(adj_mat['distance'], 
     #                interpolation='nearest')
     ax_link.set_xlim(0, len(adj_mat))
-    ax_link.set_ylim(0, len(adj_mat))
+    ax_link.set_ylim(len(adj_mat), 0)
+    f.tight_layout()
     f.savefig(adj_matrix_outfile, dpi=300)
     print "next fig" 
     f = pylab.figure(figsize=(12, 12))

### 2013-10-22 irm/plot.py

@@ -273,6 +273,7 @@ def plot_purity_hists(fig, assign_vect, true_assign_vect,
 def plot_purity_hists_h(fig, assign_vect, true_assign_vect, 
                         thold=0.9, clust_labels = None, plot_zero=True, 
                         extra_rows = 0, extra_row_func = None, 
+                        true_class_labels = tuple(), 
                         class_colors = tuple()):
     """
 
@@ -348,7 +349,14 @@ def plot_purity_hists_h(fig, assign_vect, true_assign_vect,
         ax.set_xlim(0, 1.0)
         ax.set_xticks([0, 1.0])
         if plot_zero:
-            if class_i > 0:
+            if class_i == 0:
+                ax.set_yticks(xpos - BAR_SPACE/2 + BAR_WIDTH/2., minor=True)
+                if len(true_class_labels) > 0:
+                    ax.set_yticklabels(true_class_labels, minor=True, 
+                                       size='xx-small')
+                    ax.set_yticklabels([], major=True)
+
+            else:
                 ax.set_yticks(xpos - BAR_SPACE/2 + BAR_WIDTH/2., minor=True)
                 ax.set_yticklabels([])
                 

### 2013-10-23 experiments/jobs/pipeline.py
added zip merge preprocessing

@@ -443,7 +443,8 @@ def plot_scores_z(exp_results, (plot_latent_filename, plot_truth_filename)):
     f.savefig(plot_latent_filename)
 
 @transform(get_results, suffix(".samples"), 
-           [(".%d.latent.pdf" % d, ".%d.latent.pickle" % d)  for d in range(3)])
+           [(".%d.latent.pdf" % d, ".%d.latent.pickle" % d, 
+             ".%d.latent.maps.pdf")  for d in range(3)])
 def plot_best_latent(exp_results, 
                      out_filenames):
 
@@ -468,7 +469,7 @@ def plot_best_latent(exp_results,
 
     # get data
     
-    for chain_pos, (latent_fname, latent_pickle) in enumerate(out_filenames):
+    for chain_pos, (latent_fname, latent_pickle, maps_fname) in enumerate(out_filenames):
         best_chain_i = chains_sorted_order[chain_pos]
         best_chain = chains[best_chain_i]
         sample_latent = best_chain['state']
@@ -513,6 +514,8 @@ def plot_best_latent(exp_results,
 
         pickle.dump(sample_latent, open(latent_pickle, 'w'))
 
+
+        # 
 def plot_t1t2_params(fig, conn_and_dist, a1, a2, ss, hps, MAX_DIST=10, 
                      model="LogisticDistance", MAX_CLASSES = 20):
     """

### 2013-10-23 experiments/jobs/preprocess.py

@@ -37,16 +37,44 @@ def load_apps_df(infile, outfile):
                 open(outfile, 'w'))
 
 
-@files(os.path.join(RAW_DATA_DIR, "dl/splitjobs/jobs2.tsv"), 
-       "jobs.%d.pickle" % WINDOW_N)
-def load_jobs(infile, outfile):
-    #jobs1 has an encoding error because everything remotely associated with tsv is a cl
+@merge(os.path.join(RAW_DATA_DIR, "dl/splitjobs/jobs*.tsv"), 
+       "jobs.pickle")
+def load_jobs(infiles, outfile):
 
-    jobs_df = pandas.io.parsers.read_csv(infile, sep='\t', index_col=0, 
-                                          converters={'Zip5' : zip_conv })
+    jobs = []
+    USECOLS = [0, 1, 2,  4, 5, 6, 7, 8, 9, 10] # drop descr
+    for infile in infiles:
+        jobs.append(pandas.io.parsers.read_csv(infile, sep='\t', index_col=0, 
+                                               converters={'Zip5' : zip_conv }, 
+                                               usecols= USECOLS,
+                                               error_bad_lines=False))
+    jobs_df = pandas.concat(jobs)
     pickle.dump({'jobs' : jobs_df}, 
                 open(outfile, 'w'))
 
+@files([load_user_df, load_apps_df, load_jobs], 'appszips.pickle')
+def clean_merge_apps((users_filename, apps_filename, jobs_filename), 
+                     out_filename):
+    """
+    Clean up the apps, merge in zipcodes, 
+    """
+
+    apps_df = pickle.load(open(apps_filename, 'r'))['apps']
+    users_df = pickle.load(open(users_filename, 'r'))['users']
+    jobs_df = pickle.load(open(jobs_filename, 'r'))['jobs']
+    jobs_zips = jobs_df['Zip5']
+    users_zips = users_df['ZipCode']
+    apps_df = apps_df.join(users_zips, on='UserID')
+    apps_df = apps_df.join(jobs_zips, on='JobID')
+    
+    apps_df = apps_df.dropna(subset=['UserID', 'JobID'])
+
+    
+    pickle.dump({'apps' : apps_df}, 
+                open(out_filename, 'w'))
+
+                 
+    
 @files(os.path.join(RAW_DATA_DIR, "zip_codes_states.csv"), 
        "zipcodes.pickle")
 def load_zipcodes(infile, outfile):
[…]

### 2013-10-23 experiments/jobs/util.py

@@ -1,0 +1,21 @@
+import numpy as np
+from matplotlib import pylab
+import cPickle as pickle
+
+zipcodeshapes = pickle.load(open("jobs/zipcodeshapes.pickle"))
+
+
+def plot_zipcode_heatmap(ax, zip_vals, default_color='white', border=1):
+    for zip_i, zip_row in zipcodeshapes.iterrows():
+        z = int(zip_i)
+        points = np.array(zip_row['points'])
+        if z in zip_vals:
+            v = zip_vals[z]
+            if len(points) > 0:
+                print points
+                patch = pylab.Polygon(points , closed=True, fill=True, color=v)
+                ax.add_patch(patch)
+        else:
+            if len(points) > 0:
+                patch = pylab.Polygon(points , closed=True, fill=True, facecolor=default_color, linewidth=border, edgecolor='k')
+                ax.add_patch(patch)     

### 2013-11-03 docs/paper/paper.tex
updates to plot

@@
 \section{Introduction}
-``type'' is a recurreing theme in biology
-Cell type in bio literature, stem cells, something something
+Computing systems, biological or human-engineered, contain computing
+elements that can be classified into ``type'', which give
+insight into function.
 
-It is only through understanding the types of the system that we can understand
-its function.
+The rise of connectomics data allows discovery of cell types
+based on connectivity information.
+
+Here we describe a Bayesian non-parametric model that can discover
+both the cell types and their patterns of interconnection automatically
+from connectomics data.
+
+We apply it to the recently-released mouse retina connectome, the
+c. elegans connectome, and a ``connectome'' obtained by
+reverse-engineering a portion of a classical microprocessor.
 
 \section{Methods}
-Stochastic blockmodels long known
+Stochastic block models assume a hidden or ``latent'' type is associated
+wtih each node that ultimately influences its connectivity.
 
-We take an infinite extension of the SBM (the IRM ) \todo{cite both}
-and modify it to learn the structure of the underlying code.
+We extend this class of model to incorporate notions of spatial locality, 
+as the connectivity of many neural systems is substantially constrained.
 
-Spatial component by performing a GLM-style link function.
-
-Describe the model
-
-We develop a series of Markov-chain monte carlo scheems to solve for
-the posterior, presenting the MAP estimate. To avoid local minima we
-slowly anneal, over 1000 iterations \todo{Fix numbers}.
-
-Hyperparameter inference.
-
-We show that we recover ground truth.
+We perform posterior inference in this model using a series of Markov-chain
+Monte Carlo techniques.
 
@@
 
+\subsection{Synthetic Data}
+We show our model works by generating synthetic data with known spatial/connectivity patterns and recovering ground truth, even when the generating process makes assumptions very different from our model.
+
 \subsubsection{Mouse Retina}
-The recent release of connectome-scale data for the mouse retina.
-
-Look we recover known classes!
-
-\subsection{Drosophila}
+When applied to the mouse retina dataset, we recover spatially-homogeneous patterns of activity that reflect a coarsening of ground truth.
 
 \subsection{C. elegans}
+When applied to c. elegans, we segregate ``head and sensor'' interneurons from the spatially-distributed neurons along the body axis.
+
+\subsection{Microprocessor}
+
+We recover the logical structure of the three primary registers of the
+MOS6502 integrated circuit.
 

### 2013-11-05 docs/paper/paper.tex
Konrad's comments ont ext

@@
 
-The rise of connectomics data allows discovery of cell types
-based on connectivity information.
+The advent of connectomics data allows discovery of cell types based
+on connectivity information. The rise of clustering had a substantial
+impoact on molecular biology, and is now the only way that molecular
+biologists are able to deal with high-throughput sequencing
+technologies.
+
+[bAyesian nonparametric techniques have been proposed to 
+discover structure in data, which could be applied to connecomics data]
 
@@
 
-We apply it to the recently-released mouse retina connectome, the
-c. elegans connectome, and a ``connectome'' obtained by
-reverse-engineering a portion of a classical microprocessor.
+We apply it to three very-different computing systems:
+recently-released mouse retina connectome, the c. elegans connectome,
+and a ``connectome'' obtained by reverse-engineering a portion of a
+classical microprocessor.
+
+In all cases, we recover known types and suggest possible new types.
+
+\section{Results}
+[add paragraph that summarizes methods and the question]
+
+\subsection{Synthetic Data}
+We show our model works by generating synthetic data with known
+spatial/connectivity patterns and recovering ground truth, even when
+the generating process makes assumptions very different from our
+model.
+
+\subsubsection{Mouse Retina}
+intro paragraph about retina data
+
+When applied to the mouse retina dataset, we recover
+spatially-homogeneous patterns of activity by anlayzing the
+connectivity data.
+
+This recovers structure similar to that known by neuroanatomists [ground truth]
+
+\subsection{C. elegans}
+intro paragraph to c elegans data
+
+When applied to c. elegans, we segregate ``head and sensor'' interneurons from the spatially-distributed neurons along the body axis.
+
+\subsection{Microprocessor}
+intro paragraph to microprocessor data
+
+We recover the logical structure of the three primary registers of the
+MOS6502 integrated circuit.
+
+\section{Discussion}
+We present a machine learning technique that allows cell types to be
+disocvered using only connectivity data.
+
+[Two parts: First is how you suck -- cover the things that are missing
+such as adding additional features ]
+
+[second is : how this changes things]
+
+In the future, we plan to extend our model with the ability to handle
+other classes of information, including genetic and histochemical
+metadata.
 
@@
 
-\section{Results}
-
-\subsection{Synthetic Data}
-We show our model works by generating synthetic data with known spatial/connectivity patterns and recovering ground truth, even when the generating process makes assumptions very different from our model.
-
-\subsubsection{Mouse Retina}
-When applied to the mouse retina dataset, we recover spatially-homogeneous patterns of activity that reflect a coarsening of ground truth.
-
-\subsection{C. elegans}
-When applied to c. elegans, we segregate ``head and sensor'' interneurons from the spatially-distributed neurons along the body axis.
-
-\subsection{Microprocessor}
-
-We recover the logical structure of the three primary registers of the
-MOS6502 integrated circuit.
-
-\section{Discussion}
-
 \end{document}

### 2013-11-05 experiments/celegans/pipeline.py
fixed the cell types for c elegans

@@ -739,15 +739,19 @@ def plot_clusters_pretty_figure(purity, metadata_df, no, thold=0.9):
                            width_ratios=height_ratios, 
                            height_ratios=[0.6, 0.3])
     NEURON_CLASSES = ['AS', 'DA', 'DB', 'DD', 
-                      'VA', 'VB', 'VD', 'IL1', 
+                      'VA', 'VB', 'VC', 'VD', 'IL1', 
                       'IL2', 'OLQ', 'RMD', 'SAA', 
                       'SIB', 'SMB', 'SMD', 
-                      'URA', 'URB', 'URY']
+                      'CEP', 'SIA', 
+                      'URA', # 'URB' only has 2, 
+                      'URY']
     NEURON_CLASSES_ROLES = ['M', 'M', 'M', 'M', 
-                            'M', 'M', 'M', '',
+                            'M', 'M', 'M', 'M', '',
                             '', 'S', 'M', 'S', 
                             '', '', 'M', 
-                            'S', 'S', 'S']
+                            'S', '', 
+                            'S', 
+                            'S']
     # the RMD and SMD types are taken from the wormatlas.org per-cell-type LUT 
     # http://www.wormatlas.org/ver1/MoW_built0.92/cells/rmd.html
 

### 2013-11-05 experiments/mos6502/pipeline.py
additional runners for mos6502

@@ -41,16 +41,20 @@ EXPERIMENTS = [('mos6502.all.decode.bb', 'fixed_20_200', 'anneal_slow_400'),
 
                ('mos6502.typed.xysregs.bb', 'fixed_2_40', 'debug'), 
                ('mos6502.typed.xysregs.ld', 'fixed_2_40', 'debug'), 
+               ('mos6502.typed.xysregs.ldfl', 'fixed_2_40', 'debug'), 
 
                ('mos6502.typed.xysregs.bb', 'fixed_100_200', 'anneal_slow_400'), 
                ('mos6502.typed.xysregs.ld', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.xysregs.ldfl', 'fixed_100_200', 'anneal_slow_400'), 
 
-               #('mos6502.typed.decode.bb', 'fixed_100_200', 'anneal_slow_400'), 
-               #('mos6502.typed.decode.ld', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.decode.bb', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.decode.ld', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.decode.ldfl', 'fixed_100_200', 'anneal_slow_400'), 
 
-               #('mos6502.typed.lower.bb', 'fixed_100_200', 'anneal_slow_400'), 
-               #('mos6502.typed.lower.ld', 'fixed_100_200', 'anneal_slow_400'), 
-
+               ('mos6502.typed.lower.bb', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.lower.ld', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.lower.ldfl', 'fixed_100_200', 'anneal_slow_400'), 
+               
                #('mos6502.typed.all.bb', 'fixed_100_200', 'anneal_slow_400'), 
                #('mos6502.typed.all.ld', 'fixed_100_200', 'anneal_slow_400'), 
                # ('mos6502.all.bb', 'fixed_20_200', 'default_100'), 
@@ -92,6 +96,22 @@ def grid_logistic_distance_hypers():
 
 slow_anneal[0][1]['subkernels'][-1][1]['grids']['LogisticDistance'] = grid_logistic_distance_hypers()
 
[…]

### 2013-11-05 irm/experiments.py
hp plotting for LDFL

@@ -252,6 +252,10 @@ def plot_chains_hypers(f, chains, data):
             for p in ['mu_hp', 'lambda_hp', 'p_min', 'p_max']:
                 per_r_hp[r].append(p)
                 hp_n +=1 
+        elif m == 'LogisticDistanceFixedLambda':
+            for p in ['mu_hp', 'lambda', 'p_min', 'p_scale_alpha_hp', 'p_scale_beta_hp']:
+                per_r_hp[r].append(p)
+                hp_n +=1 
         elif m == 'NormalDistanceFixedWidth':
             for p in ['mu_hp', 'p_alpha', 'p_beta', 'p_min', 'width']:
                 per_r_hp[r].append(p)

### 2013-11-05 irm/plot.py
for LDFL

@@ -228,6 +228,16 @@ def plot_t1t1_params(fig, conn_and_dist, assign_vect, ss, hps, MAX_DIST=10,
                 ax.text(0, 0.2, r"mu: %3.2f" % c['mu'], fontsize=4)
                 ax.text(0, 0.6, r"lamb: %3.2f" % c['lambda'], fontsize=4)
                 ax.axvline(c['mu'], c='k')
+
+            elif model == "LogisticDistanceFixedLambda":
+                c = ss[(c1, c2)]
+                print "MAX_DISTANCE=", MAX_DIST, np.max(fine_bins), np.max(bins), c
+                y = util.logistic(fine_bins, c['mu'], hps['lambda']) 
+                y = y * (c['p_scale'] - hps['p_min']) + hps['p_min']
+                ax.plot(fine_bins, y, c='r') 
+                ax.text(0, 0.2, r"mu: %3.2f" % c['mu'], fontsize=4)
+                ax.text(0, 0.6, r"lamb: %3.2f" % hps['lambda'], fontsize=4)
+                ax.axvline(c['mu'], c='k')
             elif model == "ExponentialDistancePoisson":
                 c = ss[(c1, c2)]
                 print "MAX_DISTANCE=", MAX_DIST, np.max(fine_bins), np.max(bins), c

### 2013-11-05 experiments/celegans/pipeline.py
I believe we have been computing the dramatically wrong distance information the entire time. I find that... amazing. What the hell have we actually been running then?

@@ -210,9 +210,8 @@ def create_latents_2r_paramed(infile,
     neurons = data['neurons']
     canonical_neuron_ordering = data['canonical_neuron_ordering']
     NEURON_N = len(canonical_neuron_ordering)
-    dist_matrix = np.zeros((NEURON_N, NEURON_N), 
-                           dtype=[('link', np.uint8), 
-                                  ('distance', np.float32)])
+    dist_matrix = np.zeros((NEURON_N, NEURON_N), dtype=np.float32)
+
 
     for n1_i, n1 in enumerate(canonical_neuron_ordering):
         for n2_i, n2 in enumerate(canonical_neuron_ordering):

### 2013-11-11 experiments/celegans/pipeline.py
well i guess it's time to make the plotting not suck

@@ -36,6 +36,9 @@ EXPERIMENTS = [
     #('celegans.2r.sdb.00', 'fixed_100_200', 'anneal_slow_400'),  
     #('celegans.2r.ndfw.00', 'fixed_100_200', 'anneal_slow_400'),  
     ('celegans.2r.edp.00', 'fixed_100_200', 'anneal_slow_400'),  
+    ('celegans.2r.ldp.00', 'fixed_100_200', 'anneal_slow_400'),  
+    #('celegans.2r.ldp.00', 'fixed_10_100', 'anneal_slow_20'),  
+    #('celegans.2r.edp.00', 'fixed_10_100', 'anneal_slow_20'),  
 
            ]
 
@@ -61,6 +64,7 @@ LDFL_HPS = [(0.2, 0.4, 0.01, 1.0, 1.0)] # , (0.1, 0.1, 0.5, 0.001),
 NDFW_HPS = [(0.1, 0.001, 0.1, 1.0, 1.0)]
 SDB_HPS = [(1.0, 1.0, 0.1, 0.001, 0.5, 4.0)]
 EDP_HPS = [(0.1, 1.0), (1.0, 1.0), (0.1, 2.0)]
+LDP_HPS = [(1.0, 1.0, 0.01, 2.0)]
 default_nonconj = irm.runner.default_kernel_nonconj_config()
 default_conj = irm.runner.default_kernel_config()
 default_anneal = irm.runner.default_kernel_anneal()
@@ -142,14 +146,21 @@ slow_anneal[0][1]['subkernels'][-1][1]['grids']['NormalDistanceFixedWidth'] = ge
 
 slow_anneal[0][1]['subkernels'][-1][1]['grids']['BetaBernoulli'] =  [{'alpha' : a, 'beta' : b} for a, b in irm.util.cart_prod([irm.util.logspace(0.001, 1.0, 20), irm.util.logspace(0.1, 10, 20)])]
 
-slow_anneal[0][1]['subkernels'][-1][1]['grids']['ExponentialDistancePoisson'] = irm.gridgibbshps.default_grid_exponential_distance_poisson(dist_scale = 50.0, rate_scale_scale = 50.0, GRIDN = 50)
[…]

### 2013-11-11 experiments/mos6502/sixrelation.py

@@ -4,98 +4,125 @@ from matplotlib import pylab
 from matplotlib.backends.backend_pdf import PdfPages
 import irm    
 
-raw_data = pickle.load(open("data.pickle", 'r'))
-wiredf = raw_data['wiredf']
-tfdf = raw_data['tfdf']
+for dataset, model in [('xysregs', 'ldfl'), 
+                       ('xysregs', 'ld'), 
+                       ('xysregs', 'bb'), 
+                       ('decode', 'ldfl'), 
+                       ('decode', 'ld'), 
+                       ('lower', 'ldfl'), 
+                       ('lower', 'ld'), 
 
-x = pickle.load(open("typed.xysregs.region.pickle"))
-adj_mat = x['adj_mat']
-#typed_adjmat = pickle.load(open("typed.adjmat.pickle"))
-#pin_pairs = typed_adjmat['pin_pairs'] 
+]:
 
-# Should really load these from above file
+    raw_data = pickle.load(open("data.pickle", 'r'))
+    wiredf = raw_data['wiredf']
+    tfdf = raw_data['tfdf']
 
-pin_pairs = [('gate', 'gate'), ('gate', 'c1'), ('gate', 'c2'), 
-             ('c1', 'c1'), ('c1', 'c2'), ('c2', 'c2')]
+    x = pickle.load(open("typed.%s.region.pickle" % dataset))
+    adj_mat = x['adj_mat']
+    #typed_adjmat = pickle.load(open("typed.adjmat.pickle"))
+    #pin_pairs = typed_adjmat['pin_pairs'] 
 
+    # Should really load these from above file
 
-RELATION_LIST = ["%s_%s" % a for a in pin_pairs]
-
-sample = pickle.load(open("data/mos6502.typed.xysregs.ld.data-fixed_100_200-anneal_slow_400.0.latent.pickle", 'r'))
-print "main data loaded"
-
-df = x['subdf']
-print x['region']
-df['cluster'] = sample['domains']['d1']['assignment']
-for pin in ['gate', 'c1', 'c2']:
-    df = df.join(wiredf['name'], on=[pin], rsuffix='.%s'% pin)
-print df.head()
-
-fid = open("output.html", 'w')
-
-for c_n, c in df.groupby('cluster'):
-    fid.write(c.sort('name.gate').to_html())
+    pin_pairs = [('gate', 'gate'), ('gate', 'c1'), ('gate', 'c2'), 
[…]

### 2013-11-11 experiments/mouseretina/pipeline.py

@@ -588,7 +588,7 @@ def plot_best_cluster_latent(exp_results,
 
         util.plot_latent(sample_latent, dist_matrix, latent_fname, 
                          model = data['relations']['R1']['model'],
-                         PLOT_MAX_DIST=120.0, MAX_CLASSES=20)
+                         PLOT_MAX_DIST=150.0, MAX_CLASSES=20)
         
 @transform(get_results, suffix(".samples"), [".hypers.pdf"])
 def plot_hypers(exp_results, (plot_hypers_filename,)):
@@ -650,6 +650,69 @@ def plot_params(exp_results, (plot_params_filename,)):
     
     f.savefig(plot_params_filename)
 
+CIRCOS_DIST_THRESHOLDS = [20, 50, 100]
+
+@transform(get_results, suffix(".samples"), 
+           [(".circos.%02d.png" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
+def plot_circos_latent(exp_results, 
+                       out_filenames):
+
+    sample_d = pickle.load(open(exp_results))
+    chains = sample_d['chains']
+    
+    exp = sample_d['exp']
+    data_filename = exp['data_filename']
+    data = pickle.load(open(data_filename))
+    data_basename, _ = os.path.splitext(data_filename)
+    meta = pickle.load(open(data_basename + ".meta"))
+
+    meta_infile = meta['infile']
+    print "meta_infile=", meta_infile
+
+    d = pickle.load(open(meta_infile, 'r'))
+    conn = d['dist_matrix']['link']
+    cell_id_permutation = d['cell_id_permutation']
+
+    dist_matrix = d['dist_matrix']
+    orig_data = pickle.load(open(d['infile']))
+    cell_types = d['types'][:len(conn)]
+
+    type_metadata_df = pickle.load(open("type_metadata.pickle", 'r'))['type_metadata']
+
+    chains = [c for c in chains if type(c['scores']) != int]
+    CHAINN = len(chains)
+
+    chains_sorted_order = np.argsort([d['scores'][-1] for d in chains])[::-1]
+    chain_pos = 0
+
+    best_chain_i = chains_sorted_order[chain_pos]
+    best_chain = chains[best_chain_i]
+    sample_latent = best_chain['state']
+    cell_assignment = np.array(sample_latent['domains']['d1']['assignment'])
+
+    model_name = data['relations']['R1']['model']
+    for fi, circos_filename in enumerate(out_filenames):
[…]

### 2013-11-11 experiments/mouseretina/util.py

@@ -13,7 +13,7 @@ from irm.experiments import plot_latent
 
 def plot_cluster_properties(assignments, true_cell_types, 
                             pos_vec, synapses, output_filename,
-                            cluster_thold= 0.98, 
+                            cluster_thold= 1.01, 
                             class_colors=tuple()):
 
     """

### 2013-11-11 irm/experiments.py

@@ -268,6 +268,10 @@ def plot_chains_hypers(f, chains, data):
             for p in ['rate_scale_hp', 'mu_hp']:
                 per_r_hp[r].append(p)
                 hp_n +=1 
+        elif m == 'LogisticDistancePoisson':
+            for p in ['rate_scale_hp', 'mu_hp', 'lambda']:
+                per_r_hp[r].append(p)
+                hp_n +=1 
         else:
             raise RuntimeError("Unknown model'%s'" % m)
     pos = 1

### 2013-11-11 irm/gridgibbshps.py

@@ -122,7 +122,7 @@ def default_grid_logistic_distance_poisson(dist_scale = 1.0, rate_scale_scale = 
         for r in rate_scale:
             hps.append({'mu_hp' : m, 
                         'lambda' : m, 
-                        'rate_scale_hp' : rate_scale, 
+                        'rate_scale_hp' : r, 
                         'rate_min' : 0.01})
     return hps
 

### 2013-11-11 irm/plot.py

@@ -253,6 +253,23 @@ def plot_t1t1_params(fig, conn_and_dist, assign_vect, ss, hps, MAX_DIST=10,
                 ax.set_ylim(-1, 20.0)
                 ax.axvline(c['mu'], c='k')
 
+            elif model == "LogisticDistancePoisson":
+                if len(flatten_dist_val) > 0:
+                    x_jitter = np.random.normal(0, 0.01, len(flatten_dist_val))
+                    y_jitter = np.random.normal(0, 0.05, len(flatten_dist_val))
+                    ax.scatter(flatten_dist_val[:, 0] + x_jitter, 
+                               flatten_dist_val[:, 1] + y_jitter, 
+                               edgecolor='none', 
+                               s=2)
+                c = ss[(c1, c2)]
+                y = util.logistic(fine_bins, c['mu'], hps['lambda']) 
+                y = y * (c['rate_scale'] - hps['rate_min']) + hps['rate_min']
+                ax.plot(fine_bins, y, c='r') 
+                ax.text(0, 1, r"mu: %3.2f" % c['mu'], fontsize=4)
+                ax.text(0, 3, r"rate_scale: %3.2f" % c['rate_scale'], fontsize=4)
+                ax.set_ylim(-1, 20.0)
+                ax.axvline(c['mu'], c='k')
+
             elif model == "LinearDistance":
                 print "MAX_DISTANCE=", MAX_DIST, np.max(fine_bins), np.max(bins)
                 c = ss[(c1, c2)]

### 2013-11-11 irm/plots/circos.py
added multi-ribbon-plotting

@@ -47,7 +47,7 @@ class CircosPlot(object):
         self.labels = None
         self.labels_config = {'label_size' : '40p'}
         self.links = None
-        self.class_ribbons = None
+        self.class_ribbons = []
 
     def set_entity_labels(self, labels, **kargs):
         assert len(labels) == len(self.init_assign_vect)
@@ -63,10 +63,14 @@ class CircosPlot(object):
         self.links = links
 
     def set_class_ribbons(self, ribbons):
+        self.add_class_ribbons(ribbons)
+
+    def add_class_ribbons(self, ribbons, color='black_a5'):
         """
         list of (source group, dest group, width)
         """
-        self.class_ribbons = ribbons
+        self.class_ribbons.append((ribbons, color))
+
     
 def write(config, outfilename, tempdir=None):
     """
@@ -124,11 +128,11 @@ def write(config, outfilename, tempdir=None):
             fid.close()
 
         # write the ribbons
-        if config.class_ribbons != None:
+        for ribbon_i, (class_ribbons, ribbon_color) in enumerate(config.class_ribbons):
             ribbon_template = Template(read_template("circos_ribbons.template"))
             ribbons = []
             # write wribbons
-            for src_id, dest_id, width in config.class_ribbons:
+            for src_id, dest_id, width in class_ribbons:
                 ribbons.append({'src_c' : "c%d" % src_id, 
                                 'src_len' : len(config.chromosomes[src_id]), 
                                 'dest_c' : "c%d" % dest_id, 
@@ -136,7 +140,7 @@ def write(config, outfilename, tempdir=None):
                                 'width' : width})
 
             ribbons_str = ribbon_template.render(ribbons= ribbons)
-            fid = open("ribbons.txt", 'w')
+            fid = open("ribbons.%d.txt" % ribbon_i, 'w')
             fid.write(ribbons_str)
             fid.close()
         
@@ -146,7 +150,7 @@ def write(config, outfilename, tempdir=None):
         conf_template = Template(read_template("circos_conf.template"))
         conf_str = conf_template.render(has_labels = (config.labels != None), 
                                         has_links = (config.links != None), 
-                                        has_ribbons = (config.class_ribbons != None), 
+                                        ribbons = config.class_ribbons, 
                                         labels_config = config.labels_config)
         
         fid = open("circos.conf", 'w')

### 2013-11-11 tests/test_plots_circos.py

@@ -51,9 +51,20 @@ def test_ribbons():
     cp = circos.CircosPlot(av)
     cp.set_entity_labels(["conn%d" % x for x in  np.arange(N)])
     
-    cp.set_class_ribbons([(0, 4), (7, 3)])
+    cp.set_class_ribbons([(0, 4, 2), (7, 3, 2)])
                  
     circos.write(cp, 'ribbons.100.png', tempdir='plots')
+
+def test_multi_ribbons():
+    N = 100
+    av = np.arange(N, dtype=np.int32) % 10
+    cp = circos.CircosPlot(av)
+    cp.set_entity_labels(["conn%d" % x for x in  np.arange(N)])
+    
+    cp.add_class_ribbons([(0, 4, 2), (7, 3, 2)], 'red_a5')
+    cp.add_class_ribbons([(2, 3, 6), (1, 4, 4)], 'blue_a5')
+                 
+    circos.write(cp, 'ribbons.multi.100.png', tempdir='plots')
 
     
 def test_both():
@@ -63,6 +74,6 @@ def test_both():
     cp.set_entity_labels(["conn%d" % x for x in  np.arange(N)])
     cp.set_entity_links(zip(np.random.permutation(N), np.random.permutation(N)))
     
-    cp.set_class_ribbons([(0, 4), (7, 3)])
+    cp.set_class_ribbons([(0, 4, 2), (7, 3, 4)])
                  
     circos.write(cp, 'both.100.png', tempdir='plots')

### 2013-11-14 docs/paper/paper.tex
well i guess it's time to make the plotting not suck
added a bunch of figures

@@
 
+\begin{figure}
+  \centering 
+  \caption{This is the pedagogy figure where we explain what is is that the model does }
+\end{figure}
+
 \section{Results}
@@
 
+\begin{figure}
+  \centering 
+  \caption{This is the synthetic data figure where we show that our model works with synthetic data across a variety of topologies and geometries}
+\end{figure}
+
 \subsubsection{Mouse Retina}
-intro paragraph about retina data
+
+\begin{figure}
+  \centering 
+  \caption{Mouse retina data}
+\end{figure}
+
+\autocite{retina} used serial-section electron microscopy to acquire the connectome
+of a 100um x 70 um region of the mouse retina, measuring the number of putative
+synaptic touch-points between nearly 1000 cells.
 
@@
 
+\begin{figure}
+  \centering 
+  \caption{C elegans data}
+\end{figure}
+
 \subsection{Microprocessor}
@@
 MOS6502 integrated circuit.
+
+\begin{figure}
+  \centering 
+  \caption{Microprocessor data}
+\end{figure}
 

### 2013-11-14 docs/fig1.py

@@ -7,7 +7,7 @@ def dist(a, b):
 
 np.random.seed(0)
 
-SIDE_N = 5
+SIDE_N = 6
 
 # class_conn = {(0, 1) : ('d', 1.0, 0.7), 
 #               (1, 2) : ('d', 2.0, 0.8), 
@@ -32,11 +32,13 @@ SIDE_N = 5
 
 
 
-class_conn = {(0, 1) : (1.0, 0.8), 
+class_conn = {(0, 1) : (0.7, 0.8), 
               (1, 2) : (2.0, 0.6), 
-              (3, 2) : (0.7, 0.9), 
+              #(3, 2) : (0.7, 0.9), 
               (2, 1) : (0.7, 0.8), 
-              (0, 3) : (1.7, 0.7)}
+              (0, 3) : (1.7, 0.7), 
+              (3, 4) : (2.4, 0.8), 
+              (4, 0) : (1.0, 0.3)}
 
 
 nodes_with_class, connectivity = irm.data.generate.c_class_neighbors(SIDE_N, 
@@ -65,7 +67,12 @@ c_r = c_r[:, ai]
 
 ax.imshow(c_r, interpolation='nearest', cmap=pylab.cm.Greys)
 
+ax.set_xticks([])
+ax.set_yticks([])
+
 c_class = nodes_with_class['class']
+CLASS_N = len(np.unique(c_class))
+
 ai = np.argsort(c_class).flatten()
 c_sorted = connectivity[ai, :]
 c_sorted = c_sorted[:, ai]
@@ -80,6 +87,8 @@ ax.imshow(c_sorted, interpolation='nearest', cmap=pylab.cm.Greys)
 for d in di:
     ax.axhline(d+0.5)
     ax.axvline(d+0.5)
+ax.set_xticks([])
+ax.set_yticks([])
 f2.savefig("source.f1.sorted.pdf")
 
 # hilariously construct suffstats and hps by hand
@@ -88,12 +97,12 @@ hps = {'mu_hp' : 1.0,
        'p_max' : 0.9, 
        'p_min' : 0.0001}
 ss = {}
-for c1 in range(4):
-    for c2 in range(4):
+for c1 in range(CLASS_N):
+    for c2 in range(CLASS_N):
         c = (c1, c2)
         if c in class_conn:
             mu = class_conn[c][0]
-            lamb = class_conn[c][0]/8
+            lamb = class_conn[c][0]/4
         else:
             mu = 0.0001
             lamb = 0.0001
[…]

### 2013-11-14 experiments/mos6502/sixrelation_circos.py

@@ -42,7 +42,6 @@ for dataset, model in [('xysregs', 'ldfl'),
     df['cluster'] = sample['domains']['d1']['assignment']
     for pin in ['gate', 'c1', 'c2']:
         df = df.join(wiredf['name'], on=[pin], rsuffix='.%s'% pin)
-    print df.head()
 
     fid = open("sixrelation.%s.%s.output.html" % (dataset, model), 'w')
 
@@ -95,6 +94,89 @@ for dataset, model in [('xysregs', 'ldfl'),
 
             circos_p.add_class_ribbons(ribbons, color)
 
+            def conv(s):
+                out = []
+                for n in s:
+                    a = 'A'
+                    if n == 'cclk':
+                        a = 'C'
+                    elif n == 'vss':
+                        a = 'O'
+                    # elif n in ['x%d' % d for d in range(8)]:
+                    #     a = 'I'
+                    # elif n in ['y%d' % d for d in range(8)]:
+                    #     a = 'I'
+                    # elif n in ['s%d' % d for d in range(8)]:
+                    #     a = 'I'
+                    elif type(n) == str and (('not' in n) or ('~' in n)):
+                        a = 'I'
+                    out.append(a)
 
-            irm.plots.circos.write(circos_p, "sixrelation.%s.%s.circos.%d.%s.svg" % (dataset, model, fi, relation))
+
+
+                return out
+
+            def clean(x):
+                out = []
+                for y in x:
+                    if type(y) == float:
+                        out.append('_')
+                    elif '#' in y:
+                        out.append("n%s" % y[1:] )
+                    else:
+                        out.append(y)
+                return out
+
+            # # add pin glyphs for 'vss' and 'clk'
+            # circos_p.add_plot('text', {'r0' : '1.05r', 
+            #                            'r1' : '1.10r', 
+            #                            'label_size' : '20p', 
+            #                            'label_font' : 'glyph', 
+            #                            'label_rotate' : 'yes', 
+            #                        }, 
+            #                   conv(df['name.c2']))
+
+            # circos_p.add_plot('text', {'r0' : '1.10r', 
+            #                            'r1' : '1.15r', 
+            #                            'label_size' : '20p', 
+            #                            'label_font' : 'glyph', 
+            #                            'label_rotate' : 'yes', 
+            #                        }, 
+            #                   conv(df['name.c1']))
+
+            # circos_p.add_plot('text', {'r0' : '1.15r', 
+            #                            'r1' : '1.20r', 
[…]

### 2013-11-14 experiments/mouseretina/pipeline.py

@@ -650,10 +650,11 @@ def plot_params(exp_results, (plot_params_filename,)):
     
     f.savefig(plot_params_filename)
 
-CIRCOS_DIST_THRESHOLDS = [10, 20, 50, 90]
+CIRCOS_DIST_THRESHOLDS = [10, 20, 40, 60, 80]
 
 @transform(get_results, suffix(".samples"), 
-           [(".circos.%02d.png" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
+           [(".circos.%02d.svg" % d, 
+             ".circos.%02d.small.svg" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
 def plot_circos_latent(exp_results, 
                        out_filenames):
 
@@ -694,7 +695,7 @@ def plot_circos_latent(exp_results,
     pos_vec = soma_positions['pos_vec'][cell_id_permutation]
     print "Pos_vec=", pos_vec
     model_name = data['relations']['R1']['model']
-    for fi, circos_filename in enumerate(out_filenames):
+    for fi, (circos_filename_main, circos_filename_small) in enumerate(out_filenames):
         circos_p = irm.plots.circos.CircosPlot(cell_assignment)
 
         if model_name == "LogisticDistance":
@@ -736,9 +737,25 @@ def plot_circos_latent(exp_results,
                                             'max' : 72}, 
                               cell_types)
                                             
-        irm.plots.circos.write(circos_p, circos_filename)
+        irm.plots.circos.write(circos_p, circos_filename_main)
         
+        circos_p = irm.plots.circos.CircosPlot(cell_assignment, ideogram_radius="0.5r")
         
+        if model_name == "LogisticDistance":
+            v = irm.irmio.latent_distance_eval(CIRCOS_DIST_THRESHOLDS[fi], 
+                                               sample_latent['relations']['R1']['ss'], 
+                                               sample_latent['relations']['R1']['hps'], 
+                                               model_name)
+            thold = 0.50 
+            ribbons = []
+            links = []
+            for (src, dest), p in v.iteritems():
+                print src, dest, p 
+                if p > thold:
+                    ribbons.append((src, dest, int(40*p)))
+            circos_p.set_class_ribbons(ribbons)
+                                            
+        irm.plots.circos.write(circos_p, circos_filename_small)
 
 
 pipeline_run([data_retina_adj_bin, 

### 2013-11-14 irm/plots/circos.py

@@ -30,9 +30,11 @@ class CircosPlot(object):
     2. group id: controls which order it is plotted in 
     """
 
-    def __init__(self, init_assign_vect, ideogram_radius="0.8r"):
+    def __init__(self, init_assign_vect, ideogram_radius="0.8r", 
+                 ideogram_thickness="40p", 
+                 karyotype_colors=None):
         self.init_assign_vect = init_assign_vect
-
+        self.karyotype_colors = karyotype_colors
 
         # compute chromosomes
         self.chromosome_ids_ordered = sorted(np.unique(init_assign_vect))
@@ -51,6 +53,7 @@ class CircosPlot(object):
         self.plots = []
 
         self.ideogram_radius = ideogram_radius
+        self.ideogram_thickness = ideogram_thickness
 
     def set_entity_labels(self, labels, **kargs):
         assert len(labels) == len(self.init_assign_vect)
@@ -146,7 +149,8 @@ def write(config, outfilename, tempdir=None):
             chl.append({'name' : 'c%d' % chromosome_id, 
                        'label' : 'c%d' % chromosome_id, 
                        'entities' : c})
-        karyotype_str = karyotype_template.render(chromosomes = chl)
+        karyotype_str = karyotype_template.render(chromosomes = chl, 
+                                                  colors = config.karyotype_colors)
         fid = open('karyotype.txt', 'w')
         fid.write(karyotype_str)
         fid.close()
@@ -214,7 +218,8 @@ def write(config, outfilename, tempdir=None):
                                         ribbons = config.class_ribbons, 
                                         labels_config = config.labels_config, 
                                         plots = config.plots, 
-                                        ideogram_radius=config.ideogram_radius)
+                                        ideogram_radius=config.ideogram_radius, 
+                                        ideogram_thickness = config.ideogram_thickness)
         
         fid = open("circos.conf", 'w')
         fid.write(conf_str)

### 2013-11-14 experiments/mos6502/sixrelation_circos.py
added custom color output

@@ -180,3 +180,10 @@ for dataset, model in [('xysregs', 'ldfl'),
 
             irm.plots.circos.write(circos_p, "sixrelation.%s.%s.circos.%d.%s.png" % (dataset, model, fi, relation))
 
+
+            # now the tiny plots
+        
+            circos_p = irm.plots.circos.CircosPlot(assign, ideogram_radius="0.4r", 
+                                                   ideogram_thickness="60p")
+            circos_p.add_class_ribbons(ribbons, color)
+            irm.plots.circos.write(circos_p, "sixrelation.%s.%s.circos.%d.%s.tiny.png" % (dataset, model, fi, relation))

### 2013-11-14 experiments/mouseretina/pipeline.py

@@ -5,6 +5,7 @@ import copy
 import os, glob
 import time
 from matplotlib import pylab
+import matplotlib
 
 import matplotlib.gridspec as gridspec
 
@@ -653,8 +654,8 @@ def plot_params(exp_results, (plot_params_filename,)):
 CIRCOS_DIST_THRESHOLDS = [10, 20, 40, 60, 80]
 
 @transform(get_results, suffix(".samples"), 
-           [(".circos.%02d.svg" % d, 
-             ".circos.%02d.small.svg" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
+           [(".circos.%02d.png" % d, 
+             ".circos.%02d.small.png" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
 def plot_circos_latent(exp_results, 
                        out_filenames):
 
@@ -696,7 +697,23 @@ def plot_circos_latent(exp_results,
     print "Pos_vec=", pos_vec
     model_name = data['relations']['R1']['model']
     for fi, (circos_filename_main, circos_filename_small) in enumerate(out_filenames):
-        circos_p = irm.plots.circos.CircosPlot(cell_assignment)
+        CLASS_N = len(np.unique(cell_assignment))
+        
+
+        class_ids = sorted(np.unique(cell_assignment))
+
+        custom_color_map = {}
+        for c_i, c_v in enumerate(class_ids):
+            c = np.array(pylab.cm.Set1(float(c_i) / CLASS_N)[:3])*255
+            custom_color_map['ccolor%d' % c_v]  = c.astype(int)
+
+        colors = np.linspace(0, 360, CLASS_N)
+        color_str = ['ccolor%d' % int(d) for d in class_ids]
+        print "custom_color_map=", custom_color_map
+
+        circos_p = irm.plots.circos.CircosPlot(cell_assignment, 
+                                               karyotype_colors = color_str, 
+                                               custom_color_map = custom_color_map)
 
         if model_name == "LogisticDistance":
             v = irm.irmio.latent_distance_eval(CIRCOS_DIST_THRESHOLDS[fi], 
@@ -713,8 +730,8 @@ def plot_circos_latent(exp_results,
             circos_p.set_class_ribbons(ribbons)
             pos_min = 40
             pos_max = 120
-            pos_r_min = 1.05
-            pos_r_max = 1.25
+            pos_r_min = 1.10
+            pos_r_max = 1.30
             circos_p.add_plot('scatter', {'r0' : '%fr' % pos_r_min, 
[…]

### 2013-11-14 irm/plots/circos.py

@@ -32,7 +32,8 @@ class CircosPlot(object):
 
     def __init__(self, init_assign_vect, ideogram_radius="0.8r", 
                  ideogram_thickness="40p", 
-                 karyotype_colors=None):
+                 karyotype_colors=None, 
+                 custom_color_map = None):
         self.init_assign_vect = init_assign_vect
         self.karyotype_colors = karyotype_colors
 
@@ -54,6 +55,7 @@ class CircosPlot(object):
 
         self.ideogram_radius = ideogram_radius
         self.ideogram_thickness = ideogram_thickness
+        self.custom_color_map = custom_color_map
 
     def set_entity_labels(self, labels, **kargs):
         assert len(labels) == len(self.init_assign_vect)
@@ -219,7 +221,8 @@ def write(config, outfilename, tempdir=None):
                                         labels_config = config.labels_config, 
                                         plots = config.plots, 
                                         ideogram_radius=config.ideogram_radius, 
-                                        ideogram_thickness = config.ideogram_thickness)
+                                        ideogram_thickness = config.ideogram_thickness, 
+                                        custom_color_map = config.custom_color_map)
         
         fid = open("circos.conf", 'w')
         fid.write(conf_str)

### 2013-11-14 tests/test_plots_circos.py

@@ -131,3 +131,15 @@ def test_glyph_plot():
                 data_points)
     
     circos.write(cp, 'plots.glyphs.png', tempdir='plots')
+
+def test_custom_colors():
+    # all one
+    av = np.arange(100, dtype=np.int32) % 10
+    
+    custom_color_map = {'mycolor%d' % d : (255, d * 20 , 0) for d in range(10)}
+
+    cp = circos.CircosPlot(av, 
+                           karyotype_colors = ['mycolor%d' % d for d in range(10)], 
+                           custom_color_map = custom_color_map)
+
+    circos.write(cp, 'custom.colors.png', tempdir='plots')

### 2013-11-15 experiments/mouseretina/pipeline.py
well some version of these retina plots is making progress

@@ -712,6 +712,7 @@ def plot_circos_latent(exp_results,
         print "custom_color_map=", custom_color_map
 
         circos_p = irm.plots.circos.CircosPlot(cell_assignment, 
+                                               ideogram_radius="0.7r", 
                                                karyotype_colors = color_str, 
                                                custom_color_map = custom_color_map)
 
@@ -730,8 +731,9 @@ def plot_circos_latent(exp_results,
             circos_p.set_class_ribbons(ribbons)
             pos_min = 40
             pos_max = 120
-            pos_r_min = 1.10
-            pos_r_max = 1.30
+            pos_r_min = 1.00
+            pos_r_max = pos_r_min + 0.25
+
             circos_p.add_plot('scatter', {'r0' : '%fr' % pos_r_min, 
                                           'r1' : '%fr' % pos_r_max, 
                                           'min' : pos_min, 
@@ -748,15 +750,48 @@ def plot_circos_latent(exp_results,
                                                   'thickness' : 1, 
                                                   'spacing' : '0.05r'})]})
             
-            circos_p.add_plot('heatmap', {'r0' : '1.05r', 
-                                            'r1' : '1.10r', 
-                                            'min' : 0, 
-                                            'max' : 72}, 
-                              cell_types)
+            # circos_p.add_plot('heatmap', {'r0' : '1.05r', 
+            #                                 'r1' : '1.10r', 
+            #                                 'min' : 0, 
+            #                                 'max' : 72}, 
+            #                   cell_types)
+            type_color_map = {'gc' : 0, 
+                              'ac' : 1, 
+                              'bc' : 2, 
+                              'other' : 3}
+
+            TYPE_N = np.max(cell_types) + 1
+
+            type_lut = []
+            for i in range(TYPE_N):
+                if (i < 70):
+                    d = type_metadata_df.loc[i+1]['desig']
+                else:
+                    d = "  "
+                type_lut.append(type_color_map.get(d[:2], 3))
+
+            circos_p.add_plot('heatmap', {'r0' : '1.25r', 
+                                          'r1' : '1.28r', 
+                                          'min' : 0, 
+                                          'max' : 3, 
+                                          'color': "red,blue,green,white"}, 
+                              [type_lut[i] for i in cell_types])
+            
+            # circos_p.add_plot('scatter', {'r0' : '1.01r', 
+            #                               'r1' : '1.10r', 
[…]

### 2013-11-15 tests/test_plots_circos.py

@@ -143,3 +143,4 @@ def test_custom_colors():
                            custom_color_map = custom_color_map)
 
     circos.write(cp, 'custom.colors.png', tempdir='plots')
+

### 2013-11-30 experiments/mos6502/pipeline.py
updates to figs paper making progress

@@ -42,10 +42,12 @@ EXPERIMENTS = [('mos6502.all.decode.bb', 'fixed_20_200', 'anneal_slow_400'),
                ('mos6502.typed.xysregs.bb', 'fixed_2_40', 'debug'), 
                ('mos6502.typed.xysregs.ld', 'fixed_2_40', 'debug'), 
                ('mos6502.typed.xysregs.ldfl', 'fixed_2_40', 'debug'), 
+               ('mos6502.typed.xysregs.ndfw', 'fixed_2_40', 'debug'), 
 
                ('mos6502.typed.xysregs.bb', 'fixed_100_200', 'anneal_slow_400'), 
                ('mos6502.typed.xysregs.ld', 'fixed_100_200', 'anneal_slow_400'), 
                ('mos6502.typed.xysregs.ldfl', 'fixed_100_200', 'anneal_slow_400'), 
+               ('mos6502.typed.xysregs.ndfw', 'fixed_100_200', 'anneal_slow_400'), 
 
                ('mos6502.typed.decode.bb', 'fixed_100_200', 'anneal_slow_400'), 
                ('mos6502.typed.decode.ld', 'fixed_100_200', 'anneal_slow_400'), 
@@ -111,6 +113,26 @@ def grid_logistic_distance_fixed_lambda_hypers():
     return res
 
 slow_anneal[0][1]['subkernels'][-1][1]['grids']['LogisticDistanceFixedLambda'] = grid_logistic_distance_fixed_lambda_hypers()
+
+def grid_normal_distance_fixed_width_hypers():
+    p_mins = np.array([0.001])
+    mus = irm.util.logspace(10, 100, 10)
+    widths = irm.util.logspace(10, 100, 10)
+    alphabetas = [0.1, 1.0, 2.0]
+    hps = []
+    for p_min in p_mins:
+        for mu in mus:
+            for width in widths:
+                for a in alphabetas:
+                    for b in alphabetas:
+                        hps.append({'mu_hp' : mu, 
+                                    'p_min' : p_min, 
+                                    'width' : width, 
+                                    'p_alpha' : a,
+                                    'p_beta' : b})
+    return hps
+
+slow_anneal[0][1]['subkernels'][-1][1]['grids']['LogisticDistanceFixedLambda'] = grid_normal_distance_fixed_width_hypers()
 
 default_nonconj = irm.runner.default_kernel_nonconj_config()
 
[…]

### 2013-11-30 experiments/mos6502/sixrelation.py

@@ -7,6 +7,7 @@ import irm
 for dataset, model in [('xysregs', 'ldfl'), 
                        ('xysregs', 'ld'), 
                        ('xysregs', 'bb'), 
+                       ('xysregs', 'ndfw'), 
                        ('decode', 'ldfl'), 
                        ('decode', 'ld'), 
                        ('lower', 'ldfl'), 

### 2013-11-30 experiments/mos6502/sixrelation_circos.py

@@ -8,6 +8,7 @@ CIRCOS_DIST_THRESHOLDS = [20, 50, 100, 200, 500]
 
 for dataset, model in [('xysregs', 'ldfl'), 
                        ('xysregs', 'ld'), 
+                       ('xysregs', 'ndfw'), 
                        #('xysregs', 'bb'), 
                        #('decode', 'ldfl'), 
                        #('decode', 'ld'), 

### 2013-11-30 experiments/mouseretina/pipeline.py

@@ -7,6 +7,7 @@ import time
 from matplotlib import pylab
 import matplotlib
 import pandas
+import colorbrewer
 
 import matplotlib.gridspec as gridspec
 
@@ -699,7 +700,48 @@ def plot_circos_latent(exp_results,
     print "Pos_vec=", pos_vec
     model_name = data['relations']['R1']['model']
 
-    TGT_CMAP = pylab.cm.Spectral
+    # this is potentially fun: get the ranges for each type
+    TYPE_N = np.max(cell_types) + 1
+
+    df2 = pandas.DataFrame(index=np.arange(1, TYPE_N))
+    df2['des'] = type_metadata_df['coarse']
+    df2 = df2.fillna('other')
+    df2['id'] = df2.index.values.astype(int)
+    gc_mean_i = df2.groupby('des').mean().astype(int)
+    gc_min_i = df2.groupby('des').min().astype(int)
+    gc_max_i = df2.groupby('des').max().astype(int)
+
+
+
+    # create the color map
+    # P = np.array([0, gc_max_i.ix['gc'], 
+    #               gc_max_i.ix['ac'], 
+    #               gc_max_i.ix['bc'], 
+    #               TYPE_N], dtype=float)/TYPE_N
+
+    # cdict = {'green': ((P[0], 0.0, 0.0),
+    #                 (P[1], 1.0, 1.0),
+    #                 (P[2], 1.0, 1.0), 
+    #                 (P[3], 1.0, 1.0), 
+    #                 (P[4], 1.0, 1.0)),
+    #        'blue': ((P[0], 0.0, 0.0),
+    #                 (P[1], 0.0, 0.0),
+    #                 (P[2], 1.0, 1.0), 
+    #                 (P[3], 1.0, 1.0), 
+    #                 (P[4], 1.0, 1.0)),
+    #         'red': ((P[0], 0.0, 0.0),
+    #                 (P[1], 0.0, 0.0),
+    #                 (P[2], 0.0, 0.0), 
+    #                 (P[3], 1.0, 1.0), 
+    #                 (P[4], 1.0, 1.0))}
+    # my_cmap = matplotlib.colors.LinearSegmentedColormap('my_colormap',cdict,256)
+
+    TGT_CMAP = pylab.cm.gist_heat
+    coarse_colors = {'other' : [210, 210, 210]}
+    for n_i, n in enumerate(['gc', 'nac', 'mwac', 'bc']):
+        coarse_colors[n] = colorbrewer.Set1[4][n_i]
+
+    
[…]

### 2013-11-30 experiments/mouseretina/preprocess.py

@@ -252,6 +252,15 @@ def type_metadata(xlsx_file, output_file):
                     'certainty' : certainty})
     df = pandas.DataFrame(res)
     df = df.set_index(df['id'])
+
+    # the following preprocessing information was acquired from
+    # page 1 of the s1
+    df['coarse'] = df['desig'].apply(lambda x: "other")
+    df['coarse'][(df['id'] <=12)] = "gc"
+    df['coarse'][(df['id'] > 12) & (df['id'] <= 24)] = "nac"
+    df['coarse'][(df['id'] > 24) & (df['id'] <= 57)] = "mwac"
+    df['coarse'][(df['id'] > 58) & (df['id'] <= 71)] = "bc"
+    
     del df['id']
                     
     pickle.dump({'type_metadata' : df}, 

### 2013-11-30 irm/models.py

@@ -595,6 +595,14 @@ class NormalDistanceFixedWidth(object):
         
         """
 
+    def param_eval(self, d, ss, hps):
+        """
+        At distance dist, evaluate the prob of connection
+        """
+        prob =  np.exp(-0.5 * (d - ss['mu'])**2/hps['width']**2) * ss['p'] + hps['p_min']
+
+        return prob
+
 class SquareDistanceBump(object):
     """
     Just a placeholder

### 2013-11-30 docs/paper.tex
actually commit the paper

\documentclass{article}
\usepackage[english,american]{babel}
\usepackage[disable]{todonotes}
\usepackage{subfiles}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage[citestyle=authoryear,url=false,hyperref=auto,isbn=false]{biblatex}

\title{Discovering cell type motifs in connectomics data}
\author{Eric Jonas \\ Konrad Kording}

\begin{document}
\maketitle

\listoftodos

\begin{abstract}
New techniques produce massive data about neural connectivity, necessitating new analysis methods to discover underlying structure. There is strong evidence for microcircuitry, where the probability of connections depends on the type of pre- and post-synaptic neuron and their distance. Here we thus developed a nonparametric Bayesian technique that identifies type and circuitry motifs from the connection data. We show that the approach recovers known neuron types in the retina, reveals interesting structure in the nervous system of c. elegans, and automatically discovers the structure of microprocessors. Our approach is a first step towards condensing connectomics data into meaningful, human readable structure.
\end{abstract}

\section{Introduction}
Emerging techniques (cite Lichtman, Zador, Denk etc) promise to quantify the location of each neuron within a volume of interest, along with its connections to all other neurons. Far exceeding the capacity of neuroanatomists to trace small circuits, this leads to enormous datasets that quantify aspects of the nervous system. The rise of high-throughput sequencing techniques necessitated the development of novel computational methods to understand genomic structure, ushering in an era of bioinformatics as an independent discipline (cite).

The brain consists of multiple kinds of neurons, each of which is hypothesized to have a specific role in the overall computation. Neuron types differ in many ways, e.g. chemical or morphological, but they also differ in the way they connect to one another. In fact, the idea of well defined local connectivity motifs has been prominent in many areas, ranging from sensory (e.g. retina, cite) via processing (e.g. neocortex) to movement (e.g. spinal cord) (cites). Indeed, such motifs also exists in human-made computing circuits. It remains an important challenge to develop algorithms  to use anatomical data, e.g. connectomics, to back out underlying connectivity motifs.

The discovery of structure is a crucial aspect of network science. Early approaches focused on global graph properties, such as the types of scaling present in the network (cite Strogatz scale-free stuff).  While this approach leads to an understanding of global network, more recent work aims at identifying repeat patterns, or “motifs” in networks. A motif is a set of nodes with conserved relations that appears multiple times in a network, for example a set of three nodes that are each connected to one another. When it comes to neuroscience, we might expect that a motif consists of the kinds of cells that jointly make up a microcircuit along with their distance and cell-type dependent connectivity. The same motif should then appear in multiple locations in the nervous system. Combining motif-discovery ideas with specific knowledge about the kinds of motifs in the nervous system promises better structure discovery.

The discovery of structure in probabilistic graphs is a crucial aspect in machine learning. Commonly used algorithms include spectral methods (cite), community-based-detection methods (cite), and stochastic block models (cite). While these approaches powerfully incorporate the probabilistic nature of neural connections (cite) they do not generally consider the known fact that the probability of connections between neurons depends on distance. Combining such probabilistic models with a spatial component to model distance promises to improve the detection of microcircuitry from neural data.

Here we describe a Bayesian non-parametric model that can discover circuit structure automatically from connectomics data: both the cell types and their spatial patterns of interconnection. We apply it to three computing systems: the mouse retina connectome (cite), the c. elegans connectome (cite), and a ``connectome'' of a classical microprocessor. In all cases, we discover cell-type to cell-type connectivity that depends on spatial distance and these connectivity functions depend on the two involved cell types. Comparing the cell types discovered by the algorithms with those obtained by anatomists based on orthogonal information reveals a high degree of agreement. We present a scalable approach to infer microcircuitry from connectivity data.

\begin{figure}
  \centering 
  \caption{This is the pedagogy figure where we explain what is is that the model does }
\end{figure}

\section{Results}
We use a Bayesian structure discovery approach that is based on assumptions about the way positions and cell types affect the probability of connections between cells,  a so-called generative model. The probability of connections between pairs of neurons decreases with distance, but how high the baseline probability of connection is and how quickly it decays with distance depends on the two cell-types. We combine annealing with Markov Chain Monte Carlo to perform simultaneously joint posterior inference over the structure, parameters, and hyperparameters of the model. This procedure yields a division of all neurons into cell-types along with the distance and type dependent connectivity probabilities. These results can then be compared with the results of previous anatomical studies that have used other aspects of neurons, such as chemical properties and cell-morphology to divide neurons into types.

Before we can use a probabilistic model to analyze real data we first need to assess if the model works properly. We thus started out by simulating data for which we know the correct structure and comparing the estimated structure based on the algorithm (see methods) with the one we used for simulation. We find that the model does a good job of recovering the structure we used to generate the data (Fig 1A). Indeed, it does far better than a simpler and well established model (stochastic block) that assumes that there is no distance dependence to the connectivity. The model converges relatively quickly, which is enabled by xxx cool tricks we used to speed it up. It appears that the model behaves the way it should and allows application to large datasets.

\begin{figure}
  \centering 
  \caption{This is the synthetic data figure where we show that our model works with synthetic data across a variety of topologies and geometries}
\end{figure}

\subsubsection{Mouse Retina}
The retina of the mouse (cite) is an example where we should expect to have connectivity patterns that are well approximated by our generative model. It is known that there are multiple classes of cells that can be grouped into ganglion cells that transmit information to the rest of the brain, bipolar cells that connect between different cells, and amacrine cells that feed into the ganglion cells (cite some review). Recent research (cite) has produced a large dataset containing both the types of cells from orthogonal approaches, and also the connectivity matrix between all reconstructed cells. We want to know to which level we can classify neurons into types based on this connectivity only.

The algorithm took 8 hours to converge to a locally optimal solution, dividing neurons into a set of cell types (Fig 2a). For each pair of neurons there is a specific distance dependent connection probability (Fig 2b), which is well approximated by the model fit. Moreover, each type of cell is rather isotropically distributed across space (Fig 2c) as should be expected for true cell types. The algorithm is thus able to extract meaningful structure from the data.

Comparing the results of the algorithm to other information sources allows evaluating the quality of the type determination. We find that the types tend to reflect the known laminar distribution in the retina (Fig 2a, outer ring), which is exciting as the algorithm did not actually use this type of information. Moreover, we can compare the results from our algorithm with cell-type determinations by professional anatomists, who distinguish between 71 types of neurons. The algorithm yields a separation of neurons into a smaller number of types that are, however, highly correlated with those of the anatomists.

Our types closely reflect the (human-determined) segmentation of cells into retinal ganglion, narrow amacrine, medium/wide amacrine, and bipolar cells (Figure 2a, outermost ring). [TODO: Do we actually assign any of the unknown cells correctly? are types 72-79 “differet” or simply unknown?]

\begin{figure}
  \centering 
  \caption{Mouse retina data}
\end{figure}

\subsection{C. elegans}

The roundworm Caenorhabditis elegans is a model system in developmental neuroscience, with the location and connectivity of each of 302 neurons deterministically known. This incredible cross-organism consistency led to early mapping of the c. elegans chemical and electrical (gap junction) connectome. Unlike the retina, only the motor neurons in c. elegans exhibit regular distribution in space, as they follow the body axis. Most interneurons are concentrated in various ganglia that project throughout the entire animal, and the sensory neurons are primarily located in the animal head.

Using both the chemical and electrical connectivity (see methods), we determined the underlying clusters explained by connectivity and distance (fig ba). A superficial inspection of the results shows clustering into groups consisting roughly homogeneously of motor neurons, sensory neurons, and interneurons. Closer examination reveals agreement with the classifications originally outlined by Brenner in 1986. 
[create structure that mirrors the mouse retina paragraphs] 
[start paragraph with summary sentence] Motorneuron types AS, DA, and VA, all exclusively postsynaptic, are clustered together, as are motorneuron types VD and DD. VC, DB, and VB also mostly share a cluster. Various head motor neurons, including classes SMD, RMD, and RIM are clustered together. Interneurons with known anatomically-distinct connectivity patterns, such as RIA (2 cells) or AVA (2 cells) are clustered into pure groups. The algorithm even correctly places the single-cell class DVB in a distinct group. Note our clustering does not perfectly recover known groupings -- several combinations of head and sensory neurons are combined, and a difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are grouped together. [algorithm recovers surprising structure based merely on connectivity even though only 278 cells and not spatially homogeneous]

\begin{figure}
  \centering 
  \caption{C elegans data}
\end{figure}

\subsection{Microprocessor}
To show the applicability of our method to other connectome-style datasets, we obtained the spatial location and interconnectivity of a classic microprocessor, the MOS Technology 6502 used in the original Apple II, Atari, and Commodore 64 (cite datasource). We identified a region of the processor with complex but known structure containing the primary general-purpose registers X, Y, and S (figure).

Our algorithm identifies areas of spatial homogeneity 
[mirror structure]

\begin{figure}
  \centering 
  \caption{Microprocessor data}
\end{figure}

\section{Discussion}
We have presented a machine learning technique that allows cell types and connectivity structure to be discovered using only spatial and connectivity data. We have shown its applicability to a newly published connectomics dataset from the mouse retina, the classical connectivity of the worm c elegans, and the connectivity of a historical man-made microprocessor. We have found that spatial connectivity alone is sufficient to discover interesting motifs that were known to exist in the systems based on decades of previous research which had utilized extensive additional information. “

For probabilistic models like ours, no known solution exists to exactly find the most probable parsing of the neurons into cell-types and connectivity patterns [homogenize these two wordings across the paper]. We employ a collection of Markov-chain Monte carlo techniques [reference methods] but while different initializations converge to similar ultimate values, we can never realistically obtain the global optimum. There are a broad range of techniques that may offer better approximations to the global optimum (cite) and future work could adapt them to the problem we addressed here.

Our inference becomes slower as the amount of data increases. Our algorithm required 10xxx hours for 1000 neurons and would therefore take months for 10,000 neurons, which would not be acceptable. Various new techniques promise to allow similar calculations to take less time and scale to larger datasets (cites).

*Fixed and hand selected likelihood link functions We picked the sigmoid. Should be estimated, Our small collection of parametric forms for spatial functions are obviously not exhaustive

*We do not use other sources of information that are totally there in the datasets. Cell morphology could be used. Laminar organization.

*Larger datasets will generally allow algorithms to distinguish more distinct types (cite supplemental data) and more data should algorithms to get closer to the results from anatomy. Moreover, in general, for such problems precision increases with the size of the dataset and the 9xxx cells that we have are not sufficient to statistically distinguish all the cell types known in anatomy (71xxx). Still, using only connectivity it is possible to meaningfully divide neurons into classes.

[ how this changes things]
There exist a range of previous approaches to the discovery of microcircuitry [coherence of language] (cites). These generally involve a great deal of manual labor and ad-hoc determination of what constitutes a “type” of cell -- to this day there are disagreements in the literature as to the “true” types in the mammalian retina. Much as phylogenomics has changed our understanding of animal ontologies, modern large scale data will allow the efficient unbiased discovery of structure. The sheer amount of available data demands the introduction of algorithmic approaches.

*A new way of phenotyping, combining spatial motif ideas with clustering ideas

*Novel data analysis techniques are required to lead us into the big data age. Our approach is orthogonal to all others (because in high d everything is orthogonal).

Ultimately, we will need to combine information across different sources into a joint understanding of neural function. Distinct cell types differ in morphology, connectivity, transcriptomics, relation to behavior or stimuli and many other ways. Algorithms based on each type information can then be combined to synthesize all the available information from one experiment or even across experiments into a joint model of brain function.

\section{Methods}
\section{Methods Summary}

% this should clearly describe the inputs

% the outputs

% the model

We assume as input a collection of connectivity matrices between
cells, as well as a distance function $d(e_i, e_j)$. We assume there
exist an unknown number of latent (unobserved) cell types, $k$, $k \in
\{1, 2, 3, \dots\}$, and that each cell $e_i$ belongs to a single cell
type, $c_i = k$. The observed connectivity between two cells $R(e_i,
e_j)$ then only depends on their latent type and their distance
through a link function $f(c_i, c_j, d(e_i, e_j))$ If $c_i=m$ and
$c_j=n$ then there exists a latent-class-specific set of parameters
$\eta_{mn}$ which parametrizes $f$, as well as a set of global hyper parameters $\theta$.

We then jointly compute the maximum a posteriroi estimate of the class
assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and the
global model hyperparameters $\theta$ :

\begin{equation}
  \P(\vec{c}, \eta, \theta) \propto \prod_{i, j} P() \prod_{m, n} P(\eta_{mn} | \theta)  P(\theta) P(c | \alpha) p(\alpha) 
\end{equation}

% Methods section 3000 words
% methods summary does not appear in the online text
% methods section can have no figures or tables [grrr]
% todo describe multiple independent relations?

\subsection{Probabilistic Model}

Our model is a extension of  infinite stochastic block models \cite{Kemp, other guys} to incorporate spatial constraints.

Prior on cluster assignment. While dirichlet process is inconsistent, we deemed this to ultimately be inconsequential for this study.

Priors on gridded hyperparameter space.

\subsection{Inference} 
We perform posterior inference via Markov-chain Monte Carlo (MCMC),
annealing on the global likelihood during the traditional burn-in
phase. MCMC transition kernels for different parts of the state space
can be chained together to construct a kernel whose ergodic
distribution is the target ergodic distribution over the entire state space.

Our first transition kernel (``structural'') performs gibbs sampling 
of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
The lack of conjugacy in our likelihood model makes an explicit 
evaluation of the conditional assignment probabilities impossible, 
motivating us to use an auxiliary variable method \autocite{Neal}
in which a collection of ephemeral clusters are explicitly represented
for the duration of the Gibbs scan.

We then employ a transition kernel to update the per-component
parameter values $\eta_{mn}$. Conditioned on the assignment vector
$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
individual $\eta_{mn}$ are indepednent. We slice sample \autocite{neal}
each component's parameters, choosing the slice width as a function
of the global hyperparameter range.

The global hyper-parameters, both $\alpha$ and $\theta$, are allowed
to take on a discrete set of possible values. As $\theta$ is often a
tuple of possible values, we explore the cartesian product of all
possible values. We then Gibbs sample, which is always possible in a 
small, finite discrete state space.

We chain these three kernels together, and then globally
anneal on the likelihood from a temperature of $T=64$ down to 
$T=1$ over 300 iterations unless otherwise indicated, and
then run the chain for another $100$ iterations.

To pick the MAP, we run up to $100$ simultaneous chains in parallel, 
iniitalized from different random initial points in the statespace, 
and then pick the chain with the highest log likelihood. Early experiments
attempted to use the full ensemble of chains as an estimate
of the posterior but were abandoned due to the difficulty of
summarizing the posterior over such a complex state space, whcih is
an ongoing area of research in the probabilistic modeling community \autocite{}.

\subsection{Validation}
Posterior enumeration tests
extensive synthetic data
Show results in suplementary figures

\subsection {Mouse Retina}
Dense serial electron microscopy of a $VOLUMNE$ in the mouse
retina by \autocite{Helmstaedter2013} yielded a listing of places where
neurons come into contact. There were $number$ cells originally, whcih
were refined through various classification schemes. We selected the
$950$ for which the location of the soma could be reconstruted from
the provided cell plots (soma locations were not provided by the
study's authors in machine-readble form).

The authors were unable to disambiguate actual ``synapses'' in the
full data set, but were able to measure the points of contact between
cells. Thus we don't have ``real'' synapses. We also don't have directionality.

\subsection{C. elegans}
Where do we get the data
what cells

\subsection{Microprocessor}

\end{document}

### 2013-11-30 docs/paper.tex
updates to paper text

@@
 
-Our model is a extension of  infinite stochastic block models \cite{Kemp, other guys} to incorporate spatial constraints.
+Our model is a extension of infinite stochastic block models
+\cite{Kemp, other guys} to incorporate spatial constraints.
 
-Prior on cluster assignment. While dirichlet process is inconsistent, we deemed this to ultimately be inconsequential for this study.
+We describe three link functions, ``Logistic-distance'',
+
+Prior on cluster assignment. While dirichlet process is inconsistent,
+we deemed this to ultimately be inconsequential for this study.
 
@@
 iniitalized from different random initial points in the statespace,
-and then pick the chain with the highest log likelihood. Early experiments
-attempted to use the full ensemble of chains as an estimate
-of the posterior but were abandoned due to the difficulty of
-summarizing the posterior over such a complex state space, whcih is
-an ongoing area of research in the probabilistic modeling community \autocite{}.
+and then pick the chain with the highest log likelihood. Early
+experiments attempted to use the full ensemble of chains as an
+estimate of the posterior but were abandoned due to the difficulty of
+summarizing the posterior over such a complex state space, whcih is an
+ongoing area of research in the probabilistic modeling community
+\autocite{}.
 
 \subsection{Validation}
-Posterior enumeration tests
-extensive synthetic data
-Show results in suplementary figures
 
 \subsection {Mouse Retina}
-Dense serial electron microscopy of a $VOLUMNE$ in the mouse
-retina by \autocite{Helmstaedter2013} yielded a listing of places where
-neurons come into contact. There were $number$ cells originally, whcih
-were refined through various classification schemes. We selected the
-$950$ for which the location of the soma could be reconstruted from
-the provided cell plots (soma locations were not provided by the
-study's authors in machine-readble form).
-
-The authors were unable to disambiguate actual ``synapses'' in the
-full data set, but were able to measure the points of contact between
-cells. Thus we don't have ``real'' synapses. We also don't have directionality.
+Dense serial electron microscopy of a $VOLUMNE$ in the mouse retina by
+\autocite{Helmstaedter2013} yielded a listing of places where neurons
+come into contact. There were $number$ cells originally, and selected
+the $950$ for which the location of the soma could be reconstruted
+from the provided cell plots (soma locations were not provided by the
+study's authors in machine-readble form). Ultimately
+this left a matrix between the total synapse-like contact area between
+all pairs of 950 cells. Area was thresholded at $0.5\mum$, determined
+by hand, to yield a 950 $\times$ 950 entry matrix that served as input
+to our algorithm. We measured the distance between cells using
+the reconstructed soma centers, and used the Logistic-Distance
+spatial relation. Hyperprior distributions are shown in \ref{supplemental}.
 
 \subsection{C. elegans}
-Where do we get the data
-what cells
+
+We obtained the connectome of c. elegans from \autocite, and isolated
+the 279 nonpharyngeal neurons, with a total of 6393 chemical synapses
+and 890 gap junctions originally cleaned up in \autocite{Chen2006}. A
+cell's position was its distance along the anterior-posterior axis
+normalzed between zero and one. We used both networks, the chemical
+network as a directed graph and the electrical network as undirected
+graph. We use the synapse counts with the logistic-distance poisson
+likelihood, scaling the counts by a factor of 4 to compensate for 
+the Poisson's underdispersion.
 

### 2013-12-01 docs/paper.tex
mouse plot looks great, onto c elegans

@@
 \begin{abstract}
-New techniques produce massive data about neural connectivity, necessitating new analysis methods to discover underlying structure.
-There is strong evidence for microcircuitry, where the probability of connections depends on the type of pre- and post-synaptic neuron and their distance.
-Here we thus developed a nonparametric Bayesian technique that identifies type and circuitry motifs from the connection data.
-We show that the approach recovers known neuron types in the retina, reveals interesting structure in the nervous system of c.
-elegans, and automatically discovers the structure of microprocessors.
-Our approach is a first step towards condensing connectomics data into meaningful, human readable structure.
+  New techniques produce massive data about neural connectivity,
+  necessitating new analysis methods to discover underlying
+  structure. There is strong evidence for microcircuitry, where the
+  probability of connections depends on the type of pre- and
+  post-synaptic neuron and their distance. Here we thus developed a
+  nonparametric Bayesian technique that identifies type and circuitry
+  motifs from the connection data. We show that the approach recovers
+  known neuron types in the retina, reveals interesting structure in
+  the nervous system of c. elegans, and automatically discovers the
+  structure of microprocessors. Our approach is a first step towards
+  condensing connectomics data into meaningful, human readable
+  structure.
 \end{abstract}
@@
 \section{Introduction}
-Emerging techniques (cite Lichtman, Zador, Denk etc) promise to quantify the location of each neuron within a volume of interest, along with its connections to all other neurons.
-Far exceeding the capacity of neuroanatomists to trace small circuits, this leads to enormous datasets that quantify aspects of the nervous system.
-The rise of high-throughput sequencing techniques necessitated the development of novel computational methods to understand genomic structure, ushering in an era of bioinformatics as an independent discipline (cite).
+Emerging techniques (cite Lichtman, Zador, Denk etc) promise to
+quantify the location of each neuron within a volume of interest,
+along with its connections to all other neurons. Far exceeding the
+capacity of neuroanatomists to trace small circuits, this leads to
+enormous datasets that quantify aspects of the nervous system. The
+rise of high-throughput sequencing techniques necessitated the
+development of novel computational methods to understand genomic
+structure, ushering in an era of bioinformatics as an independent
+discipline (cite).
 
-The brain consists of multiple kinds of neurons, each of which is hypothesized to have a specific role in the overall computation.
-Neuron types differ in many ways, e.g.
-chemical or morphological, but they also differ in the way they connect to one another.
-In fact, the idea of well defined local connectivity motifs has been prominent in many areas, ranging from sensory (e.g.
-retina, cite) via processing (e.g.
-neocortex) to movement (e.g.
-spinal cord) (cites).
-Indeed, such motifs also exists in human-made computing circuits.
-It remains an important challenge to develop algorithms  to use anatomical data, e.g.
-connectomics, to back out underlying connectivity motifs.
+The brain consists of multiple kinds of neurons, each of which is
+hypothesized to have a specific role in the overall
+computation. Neuron types differ in many ways, e.g. chemical or
+morphological, but they also differ in the way they connect to one
+another. In fact, the idea of well defined local connectivity motifs
+has been prominent in many areas, ranging from sensory (e.g. retina,
+cite) via processing (e.g. neocortex) to movement (e.g. spinal cord)
+(cites). Indeed, such motifs also exists in human-made computing
+circuits. It remains an important challenge to develop algorithms to
+use anatomical data, e.g. connectomics, to back out underlying
+connectivity motifs.
 
-The discovery of structure is a crucial aspect of network science.
-Early approaches focused on global graph properties, such as the types of scaling present in the network (cite Strogatz scale-free stuff).
-While this approach leads to an understanding of global network, more recent work aims at identifying repeat patterns, or “motifs” in networks.
-A motif is a set of nodes with conserved relations that appears multiple times in a network, for example a set of three nodes that are each connected to one another.
-When it comes to neuroscience, we might expect that a motif consists of the kinds of cells that jointly make up a microcircuit along with their distance and cell-type dependent connectivity.
-The same motif should then appear in multiple locations in the nervous system.
-Combining motif-discovery ideas with specific knowledge about the kinds of motifs in the nervous system promises better structure discovery.
+The discovery of structure is a crucial aspect of network
+science. Early approaches focused on global graph properties, such as
+the types of scaling present in the network (cite Strogatz scale-free
+stuff).  While this approach leads to an understanding of global
+network, more recent work aims at identifying repeat patterns, or
+“motifs” in networks. A motif is a set of nodes with conserved
+relations that appears multiple times in a network, for example a set
+of three nodes that are each connected to one another. When it comes
+to neuroscience, we might expect that a motif consists of the kinds of
+cells that jointly make up a microcircuit along with their distance
+and cell-type dependent connectivity. The same motif should then
+appear in multiple locations in the nervous system. Combining
+motif-discovery ideas with specific knowledge about the kinds of
+motifs in the nervous system promises better structure discovery.
 
-The discovery of structure in probabilistic graphs is a crucial aspect in machine learning.
-Commonly used algorithms include spectral methods (cite), community-based-detection methods (cite), and stochastic block models (cite).
-While these approaches powerfully incorporate the probabilistic nature of neural connections (cite) they do not generally consider the known fact that the probability of connections between neurons depends on distance.
-Combining such probabilistic models with a spatial component to model distance promises to improve the detection of microcircuitry from neural data.
+The discovery of structure in probabilistic graphs is a crucial aspect
+in machine learning. Commonly used algorithms include spectral methods
+(cite), community-based-detection methods (cite), and stochastic block
+models (cite). While these approaches powerfully incorporate the
+probabilistic nature of neural connections (cite) they do not
+generally consider the known fact that the probability of connections
+between neurons depends on distance. Combining such probabilistic
+models with a spatial component to model distance promises to improve
+the detection of microcircuitry from neural data.
 
-Here we describe a Bayesian non-parametric model that can discover circuit structure automatically from connectomics data: both the cell types and their spatial patterns of interconnection.
-We apply it to three computing systems: the mouse retina connectome (cite), the c.
-elegans connectome (cite), and a ``connectome'' of a classical microprocessor.
-In all cases, we discover cell-type to cell-type connectivity that depends on spatial distance and these connectivity functions depend on the two involved cell types.
-Comparing the cell types discovered by the algorithms with those obtained by anatomists based on orthogonal information reveals a high degree of agreement.
-We present a scalable approach to infer microcircuitry from connectivity data.
+Here we describe a Bayesian non-parametric model that can discover
+circuit structure automatically from connectomics data: both the cell
+types and their spatial patterns of interconnection. We apply it to
+three computing systems: the mouse retina connectome (cite), the
+c. elegans connectome (cite), and a ``connectome'' of a classical
+microprocessor. In all cases, we discover cell-type to cell-type
+connectivity that depends on spatial distance and these connectivity
+functions depend on the two involved cell types. Comparing the cell
+types discovered by the algorithms with those obtained by anatomists
+based on orthogonal information reveals a high degree of agreement. We
+present a scalable approach to infer microcircuitry from connectivity
+data.
 
@@
 \section{Results}
-We use a Bayesian structure discovery approach that is based on assumptions about the way positions and cell types affect the probability of connections between cells,  a so-called generative model.
-The probability of connections between pairs of neurons decreases with distance, but how high the baseline probability of connection is and how quickly it decays with distance depends on the two cell-types.
-We combine annealing with Markov Chain Monte Carlo to perform simultaneously joint posterior inference over the structure, parameters, and hyperparameters of the model.
-This procedure yields a division of all neurons into cell-types along with the distance and type dependent connectivity probabilities.
-These results can then be compared with the results of previous anatomical studies that have used other aspects of neurons, such as chemical properties and cell-morphology to divide neurons into types.
+We use a Bayesian structure discovery approach that is based on
+assumptions about the way positions and cell types affect the
+probability of connections between cells, a so-called generative
+model. The probability of connections between pairs of neurons
+decreases with distance, but how high the baseline probability of
+connection is and how quickly it decays with distance depends on the
+two cell-types. We combine annealing with Markov Chain Monte Carlo to
+perform simultaneously joint posterior inference over the structure,
+parameters, and hyperparameters of the model. This procedure yields a
+division of all neurons into cell-types along with the distance and
+type dependent connectivity probabilities. These results can then be
+compared with the results of previous anatomical studies that have
+used other aspects of neurons, such as chemical properties and
+cell-morphology to divide neurons into types.
 
-Before we can use a probabilistic model to analyze real data we first need to assess if the model works properly.
-We thus started out by simulating data for which we know the correct structure and comparing the estimated structure based on the algorithm (see methods) with the one we used for simulation.
-We find that the model does a good job of recovering the structure we used to generate the data (Fig 1A).
-Indeed, it does far better than a simpler and well established model (stochastic block) that assumes that there is no distance dependence to the connectivity.
-The model converges relatively quickly, which is enabled by xxx cool tricks we used to speed it up.
-It appears that the model behaves the way it should and allows application to large datasets.
+Before we can use a probabilistic model to analyze real data we first
+need to assess if the model works properly. We thus started out by
+simulating data for which we know the correct structure and comparing
+the estimated structure based on the algorithm (see methods) with the
+one we used for simulation. We find that the model does a good job of
+recovering the structure we used to generate the data (Fig
+1A). Indeed, it does far better than a simpler and well established
+model (stochastic block) that assumes that there is no distance
+dependence to the connectivity. The model converges relatively
+quickly, which is enabled by xxx cool tricks we used to speed it
+up. It appears that the model behaves the way it should and allows
+application to large datasets.
 
@@
 \subsubsection{Mouse Retina}
-The retina of the mouse (cite) is an example where we should expect to have connectivity patterns that are well approximated by our generative model.
-It is known that there are multiple classes of cells that can be grouped into ganglion cells that transmit information to the rest of the brain, bipolar cells that connect between different cells, and amacrine cells that feed into the ganglion cells (cite some review).
-Recent research (cite) has produced a large dataset containing both the types of cells from orthogonal approaches, and also the connectivity matrix between all reconstructed cells.
-We want to know to which level we can classify neurons into types based on this connectivity only.
+The retina of the mouse (cite) is an example where we should expect to
+have connectivity patterns that are well approximated by our
+generative model. It is known that there are multiple classes of cells
+that can be grouped into ganglion cells that transmit information to
+the rest of the brain, bipolar cells that connect between different
+cells, and amacrine cells that feed into the ganglion cells (cite some
+review). Recent research (cite) has produced a large dataset
+containing both the types of cells from orthogonal approaches, and
+also the connectivity matrix between all reconstructed cells. We want
+to know to which level we can classify neurons into types based on
+this connectivity only.
 
-The algorithm took 8 hours to converge to a locally optimal solution, dividing neurons into a set of cell types (Fig 2a).
-For each pair of neurons there is a specific distance dependent connection probability (Fig 2b), which is well approximated by the model fit.
-Moreover, each type of cell is rather isotropically distributed across space (Fig 2c) as should be expected for true cell types.
-The algorithm is thus able to extract meaningful structure from the data.
+The algorithm took 8 hours to converge to a locally optimal solution,
+dividing neurons into a set of cell types (Fig 2a). For each pair of
+neurons there is a specific distance dependent connection probability
+(Fig 2b), which is well approximated by the model fit. Moreover, each
+type of cell is rather isotropically distributed across space (Fig 2c)
+as should be expected for true cell types. The algorithm is thus able
+to extract meaningful structure from the data.
 
-Comparing the results of the algorithm to other information sources allows evaluating the quality of the type determination.
-We find that the types tend to reflect the known laminar distribution in the retina (Fig 2a, outer ring), which is exciting as the algorithm did not actually use this type of information.
-Moreover, we can compare the results from our algorithm with cell-type determinations by professional anatomists, who distinguish between 71 types of neurons.
-The algorithm yields a separation of neurons into a smaller number of types that are, however, highly correlated with those of the anatomists.
+Comparing the results of the algorithm to other information sources
+allows evaluating the quality of the type determination. We find that
+the types tend to reflect the known laminar distribution in the retina
+(Fig 2a, outer ring), which is exciting as the algorithm did not
+actually use this type of information. Moreover, we can compare the
+results from our algorithm with cell-type determinations by
+professional anatomists, who distinguish between 71 types of
+neurons. The algorithm yields a separation of neurons into a smaller
+number of types that are, however, highly correlated with those of the
+anatomists.
 
-Our types closely reflect the (human-determined) segmentation of cells into retinal ganglion, narrow amacrine, medium/wide amacrine, and bipolar cells (Figure 2a, outermost ring).
-[TODO: Do we actually assign any of the unknown cells correctly?
-are types 72-79 “differet” or simply unknown?]
+Our types closely reflect the (human-determined) segmentation of cells
+into retinal ganglion, narrow amacrine, medium/wide amacrine, and
+bipolar cells (Figure 2a, outermost ring). [TODO: Do we actually
+assign any of the unknown cells correctly? are types 72-79 “differet”
+or simply unknown?]
 
@@
 
-The roundworm Caenorhabditis elegans is a model system in developmental neuroscience, with the location and connectivity of each of 302 neurons deterministically known.
-This incredible cross-organism consistency led to early mapping of the c.
-elegans chemical and electrical (gap junction) connectome.
-Unlike the retina, only the motor neurons in c.
-elegans exhibit regular distribution in space, as they follow the body axis.
-Most interneurons are concentrated in various ganglia that project throughout the entire animal, and the sensory neurons are primarily located in the animal head.
+The roundworm Caenorhabditis elegans is a model system in
+developmental neuroscience, with the location and connectivity of each
+of 302 neurons deterministically known. This incredible cross-organism
+consistency led to early mapping of the c. elegans chemical and
+electrical (gap junction) connectome. Unlike the retina, only the
+motor neurons in c. elegans exhibit regular distribution in space, as
+they follow the body axis. Most interneurons are concentrated in
+various ganglia that project throughout the entire animal, and the
+sensory neurons are primarily located in the animal head.
 
-Using both the chemical and electrical connectivity (see methods), we determined the underlying clusters explained by connectivity and distance (fig ba).
-A superficial inspection of the results shows clustering into groups consisting roughly homogeneously of motor neurons, sensory neurons, and interneurons.
-Closer examination reveals agreement with the classifications originally outlined by Brenner in 1986.
-[create structure that mirrors the mouse retina paragraphs] 
-[start paragraph with summary sentence] Motorneuron types AS, DA, and VA, all exclusively postsynaptic, are clustered together, as are motorneuron types VD and DD.
-VC, DB, and VB also mostly share a cluster.
-Various head motor neurons, including classes SMD, RMD, and RIM are clustered together.
-Interneurons with known anatomically-distinct connectivity patterns, such as RIA (2 cells) or AVA (2 cells) are clustered into pure groups.
-The algorithm even correctly places the single-cell class DVB in a distinct group.
-Note our clustering does not perfectly recover known groupings -- several combinations of head and sensory neurons are combined, and a difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are grouped together.
-[algorithm recovers surprising structure based merely on connectivity even though only 278 cells and not spatially homogeneous]
+Using both the chemical and electrical connectivity (see methods), we
+determined the underlying clusters explained by connectivity and
+distance (fig ba). A superficial inspection of the results shows
+clustering into groups consisting roughly homogeneously of motor
+neurons, sensory neurons, and interneurons. Closer examination reveals
+agreement with the classifications originally outlined by Brenner in
+1986.  [create structure that mirrors the mouse retina paragraphs]
+
+[start paragraph with summary sentence] Motorneuron types AS, DA, and
+VA, all exclusively postsynaptic, are clustered together, as are
+motorneuron types VD and DD. VC, DB, and VB also mostly share a
+cluster. Various head motor neurons, including classes SMD, RMD, and
+RIM are clustered together. Interneurons with known
+anatomically-distinct connectivity patterns, such as RIA (2 cells) or
+AVA (2 cells) are clustered into pure groups. The algorithm even
+correctly places the single-cell class DVB in a distinct group. Note
+our clustering does not perfectly recover known groupings -- several
+combinations of head and sensory neurons are combined, and a
+difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
+grouped together. [algorithm recovers surprising structure based
+merely on connectivity even though only 278 cells and not spatially
+homogeneous]
 
@@
 \subsection{Microprocessor}
-To show the applicability of our method to other connectome-style datasets, we obtained the spatial location and interconnectivity of a classic microprocessor, the MOS Technology 6502 used in the original Apple II, Atari, and Commodore 64 (cite datasource).
-We identified a region of the processor with complex but known structure containing the primary general-purpose registers X, Y, and S (figure).
+
+To show the applicability of our method to other connectome-style
+datasets, we obtained the spatial location and interconnectivity of a
+classic microprocessor, the MOS Technology 6502 used in the original
+Apple II, Atari, and Commodore 64 (cite datasource). We identified a
+region of the processor with complex but known structure containing
+the primary general-purpose registers X, Y, and S (figure).
 
@@
 this left a matrix between the total synapse-like contact area between
-all pairs of 950 cells. Area was thresholded at $0.5\mum$, determined
+all pairs of 950 cells. Area was thresholded at $0.5\mu m$, determined
 by hand, to yield a 950 $\times$ 950 entry matrix that served as input

### 2013-12-01 experiments/mouseretina/pipeline.py

@@ -656,8 +656,8 @@ def plot_params(exp_results, (plot_params_filename,)):
 CIRCOS_DIST_THRESHOLDS = [10, 20, 40, 60, 80]
 
 @transform(get_results, suffix(".samples"), 
-           [(".circos.%02d.png" % d, 
-             ".circos.%02d.small.png" % d, 
+           [(".circos.%02d.svg" % d, 
+             ".circos.%02d.small.svg" % d, 
              ".circos.%02d.colors.png" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
 def plot_circos_latent(exp_results, 
                        out_filenames):
@@ -740,7 +740,7 @@ def plot_circos_latent(exp_results,
     coarse_colors = {'other' : [210, 210, 210]}
     for n_i, n in enumerate(['gc', 'nac', 'mwac', 'bc']):
         coarse_colors[n] = colorbrewer.Set1[4][n_i]
-
+    print "THE COARSE COLORS ARE", coarse_colors
     
     for fi, (circos_filename_main, circos_filename_small, color_legend_filename) in enumerate(out_filenames):
         CLASS_N = len(np.unique(cell_assignment))
@@ -935,6 +935,8 @@ def plot_clustered_somapos(exp_results,
                c = cell_colors, s=60)
     ax.set_ylim(0, 85)
     ax.set_xlim(5, 115)
+    ax.set_aspect(1.0)
+    ax.plot([10, 20], [3, 3], linewidth=10, c='k')
     ax.set_xticks([])
     ax.set_yticks([])
 

### 2013-12-02 docs/paper.tex
substantial updates to including figures

@@
-\documentclass{article}
+\documentclass{nature}
 \usepackage[english,american]{babel}
-\usepackage[disable]{todonotes}
 \usepackage{subfiles}
@@
 \usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{fullpage}
 \usepackage{algorithm}
@@
 \usepackage[citestyle=authoryear,url=false,hyperref=auto,isbn=false]{biblatex}
+\addbibresource{connectomics motif.bib}
+\addbibresource{connectomics motif-relational models.bib}
+
+\DeclareGraphicsRule{.ai}{pdf}{.ai}{}
 
 \title{Discovering cell type motifs in connectomics data}
-\author{Eric Jonas \\ Konrad Kording}
+\author{Eric Jonas, Konrad Kording}
 
@@
 
-\listoftodos
+% \begin{affiliations}
+%   \item Something Something
+%   \item Something Something
+% \end{affiliations}
 
@@
 
-\section{Introduction}
+\section*{Introduction}
 Emerging techniques (cite Lichtman, Zador, Denk etc) promise to
@@
   \centering 
-  \caption{This is the pedagogy figure where we explain what is is that the model does }
+  \centerline{\includegraphics[width=8.5in]{overview.ai}}
+  \caption{This is the pedagogy figure where we explain what is is
+    that the model does. }
 \end{figure}
@@
   \centering 
+  \centerline{\includegraphics[width=8.5in]{synthetic.ai}}
   \caption{This is the synthetic data figure where we show that our model works with synthetic data across a variety of topologies and geometries}
@@
   \centering 
+  \centerline{\includegraphics[width=8.5in]{mouseretina.ai}}
   \caption{Mouse retina data}
@@
   \centering 
+  \centerline{\includegraphics[width=8.5in]{celegans.ai}}
   \caption{C elegans data}
@@
   \centering 
+  \centerline{\includegraphics[width=8.5in]{mos6502.ai}}
   \caption{Microprocessor data}
@@
 
+\begin{addendum}
+ \item Put acknowledgements here.
+ \item[Competing Interests] The authors declare that they have no
+competing financial interests.
+ \item[Correspondence] Correspondence and requests for materials
+should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
+\end{addendum}
+
 \end{document}

### 2013-12-02 experiments/celegans/pipeline.py

@@ -1081,7 +1081,7 @@ def plot_best_latent(exp_results,
         
 CIRCOS_DIST_THRESHOLDS = [0.01, 0.1, 0.25, 0.5, 0.9]
 @transform(get_results, suffix(".samples"), 
-           [(".circos.%d.png" % d, 
+           [(".circos.%d.svg" % d, 
              ) for d in range(len(CIRCOS_DIST_THRESHOLDS))])
 def plot_best_circos(exp_results, 
                      out_filenames):
@@ -1128,7 +1128,7 @@ def plot_best_circos(exp_results,
     for fi, circos_filenames in enumerate(out_filenames):
 
         circos_p = irm.plots.circos.CircosPlot(cell_assignment, 
-                                               ideogram_radius="0.7r")
+                                               ideogram_radius="0.65r")
 
         for relation, color in [('R1', 'red_a5'), 
                                 ('R2', 'blue_a5')]:
@@ -1165,7 +1165,7 @@ def plot_best_circos(exp_results,
                 for (src, dest), rate in v.iteritems():
                     scaled_rate = rate * rate_scale[relation]
                     if scaled_rate > thold:
-                        pix = int(10*scaled_rate)
+                        pix = int(5*scaled_rate)
                         print src, dest, rate, pix
 
                         ribbons.append((src, dest, pix))
@@ -1175,10 +1175,10 @@ def plot_best_circos(exp_results,
             role[neurons['role'] == 'S']= 2
             
             circos_p.add_plot('scatter', {'r0' : '1.00r', 
-                                          'r1' : '1.15r', 
+                                          'r1' : '1.25r', 
                                           'min' : 0, 
                                           'max' : 1.0, 
-                                          'glyph_size' : 20, 
+                                          'glyph_size' : 16, 
                                           'glyph' : 'circle', 
                                           'color' : 'black',
                                           'stroke_thickness' : 0}, 
@@ -1188,20 +1188,20 @@ def plot_best_circos(exp_results,
                                                                 'y1' : 1.0})],  
                                'axes': [('axis', {'color' : 'grey', 
                                                   'thickness' : 1, 
-                                                  'spacing' : '0.05r'})]})
+                                                  'spacing' : '0.25r'})]})
 
                             
-            circos_p.add_plot('heatmap', {'r0' : '1.15r', 
-                                          'r1' : '1.20r', 
+            circos_p.add_plot('heatmap', {'r0' : '1.27r', 
+                                          'r1' : '1.33r', 
[…]

### 2013-12-02 experiments/mos6502/sixrelation_circos.py

@@ -7,8 +7,8 @@ import irm
 CIRCOS_DIST_THRESHOLDS = [20, 50, 100, 200, 500]
 
 for dataset, model in [('xysregs', 'ldfl'), 
-                       ('xysregs', 'ld'), 
-                       ('xysregs', 'ndfw'), 
+                       #('xysregs', 'ld'), 
+                       #('xysregs', 'ndfw'), 
                        #('xysregs', 'bb'), 
                        #('decode', 'ldfl'), 
                        #('decode', 'ld'), 
@@ -40,7 +40,9 @@ for dataset, model in [('xysregs', 'ldfl'),
 
     df = x['subdf']
     print x['region']
-    df['cluster'] = sample['domains']['d1']['assignment']
+    assign = np.array(sample['domains']['d1']['assignment'])
+
+    df['cluster'] = assign
     for pin in ['gate', 'c1', 'c2']:
         df = df.join(wiredf['name'], on=[pin], rsuffix='.%s'% pin)
 
@@ -51,8 +53,35 @@ for dataset, model in [('xysregs', 'ldfl'),
 
     model_name = data['relations']['R1']['model']
 
+    class_ids = np.unique(assign)
+    CLASS_N = len(class_ids)
 
-    assign = np.array(sample['domains']['d1']['assignment'])
+    custom_color_map = {}
+    custom_color_map_raw = {}
+    for c_i, c_v in enumerate(class_ids):
+        color = pylab.cm.Set1(float(c_i) / (CLASS_N+1))
+        c = np.array(color)[:3]*255
+        custom_color_map['ccolor%d' % c_v]  = c.astype(int)
+        custom_color_map_raw[c_v] = color
+
+    colors = np.linspace(0, 360, CLASS_N)
+    color_str = ['ccolor%d' % int(d) for d in class_ids]
+
+    f = pylab.figure(figsize=(16, 16))
+    ax = f.add_subplot(1, 1, 1)
+    for c_n, c in df.groupby('cluster'):
+        ax.scatter(c['x'], c['y'], c = custom_color_map_raw[c_n],
+                   edgecolor='none', s=90)
+    ax.set_xlim(1200, 3200)
+    ax.set_ylim(1000, 4500)
+    ax.set_xticks([])
+    ax.set_yticks([])
+    ax.set_aspect(1.0)
+    ax.plot([1400, 1400], [1100, 1200], linewidth=4, c='k')
[…]

### 2013-12-02 experiments/mouseretina/pipeline.py

@@ -657,8 +657,7 @@ CIRCOS_DIST_THRESHOLDS = [10, 20, 40, 60, 80]
 
 @transform(get_results, suffix(".samples"), 
            [(".circos.%02d.svg" % d, 
-             ".circos.%02d.small.svg" % d, 
-             ".circos.%02d.colors.png" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
+             ".circos.%02d.small.svg" % d)  for d in range(len(CIRCOS_DIST_THRESHOLDS))])
 def plot_circos_latent(exp_results, 
                        out_filenames):
 
@@ -742,7 +741,7 @@ def plot_circos_latent(exp_results,
         coarse_colors[n] = colorbrewer.Set1[4][n_i]
     print "THE COARSE COLORS ARE", coarse_colors
     
-    for fi, (circos_filename_main, circos_filename_small, color_legend_filename) in enumerate(out_filenames):
+    for fi, (circos_filename_main, circos_filename_small) in enumerate(out_filenames):
         CLASS_N = len(np.unique(cell_assignment))
         
 
@@ -789,6 +788,7 @@ def plot_circos_latent(exp_results,
             pos_max = 120
             pos_r_min = 1.00
             pos_r_max = pos_r_min + 0.25
+            ten_um_frac = 10.0/(pos_max - pos_min)
 
             circos_p.add_plot('scatter', {'r0' : '%fr' % pos_r_min, 
                                           'r1' : '%fr' % pos_r_max, 
@@ -804,7 +804,7 @@ def plot_circos_latent(exp_results,
                                                                 'y1' : pos_max})],  
                                'axes': [('axis', {'color' : 'grey', 
                                                   'thickness' : 1, 
-                                                  'spacing' : '0.05r'})]})
+                                                  'spacing' : '%fr' % ten_um_frac})]})
             
             # circos_p.add_plot('heatmap', {'r0' : '1.34r', 
             #                                 'r1' : '1.37r', 
@@ -815,20 +815,20 @@ def plot_circos_latent(exp_results,
             #                   cell_types)
 
 
-            f_color_legend = pylab.figure()
-            ax_color_legend = f_color_legend.add_subplot(1, 1, 1)
+            # f_color_legend = pylab.figure()
[…]

### 2013-12-03 docs/paper.tex
so happy with this progress. Starting to feel like a real body of work.

@@
-\documentclass{nature}
+%\documentclass{nature}
+\documentclass{article}
 \usepackage[english,american]{babel}
+\usepackage[backend=biber,style=nature]{biblatex}
 \usepackage{subfiles}
@@
 \usepackage{multicol}
-\usepackage[citestyle=authoryear,url=false,hyperref=auto,isbn=false]{biblatex}
+
 \addbibresource{connectomics motif.bib}
@@
   \centering 
-  \centerline{\includegraphics[width=8.5in]{overview.ai}}
-  \caption{This is the pedagogy figure where we explain what is is
-    that the model does. }
+    \centerline{\includegraphics[width=183mm]{overview.ai}}
+  \caption{An overview of our method. a. Input data is
+an adjacency matrix indicating whether cell $i$ (row)
+synapses on cell $j$ (column). The location of each cell in space is known. 
+b. Our algorithm discovers hidden cell types in this connectivity data
+by assuming all cells of a type share a distance-dependent connectivity 
+profile, c.), with cells of other types. This results in a clustering 
+of the cells by those hidden types. b.) shows the cell adjacency
+matrix with cells of the same type grouped together. d.) shows
+the learned probability of connection between our different types
+at various distances, illustrating short-range and long-range
+connectivtiy patterns.}
+
 \end{figure}
@@
   \centering 
-  \centerline{\includegraphics[width=8.5in]{synthetic.ai}}
-  \caption{This is the synthetic data figure where we show that our model works with synthetic data across a variety of topologies and geometries}
+  \centerline{\includegraphics{synthetic.ai}}
+  \caption{Correct recovery of true numbers of hidden classes
+in synthetic data. a.) Other models overcluster the number
+of classes as they fail to take distance into account, whereas
+our model finds close to the true number of classes. b.) ARI, a measure
+of ``correct clustering'', for
+different true class counts, between our model and a traditional
+nonparametric stochastic block model. c.) Multiple Markov chains iterateively
+converge towards a highly-probable solution. }
+
 \end{figure}
@@
   \centering 
-  \centerline{\includegraphics[width=8.5in]{mouseretina.ai}}
-  \caption{Mouse retina data}
+  \centerline{\includegraphics{mouseretina.ai}}
+  \caption{Discovering cell classes in the mouse retina connectome. 
+a.) Input adajacency data for 950 cells for which soma positions were known. b. clsutered adjacency matrix. c.) connectivity diagram showing our clusters, as
+well as the cell depth and known cell type of each cell. Our model agrees
+with the cell types traditionally identified by anatomists, and identifies 
+laminar-specific connectivity patterns. d.) Connectivity between our
+clusters as a function of distance -- the cluster consisting primarially of
+retinal ganglion cells (lower left-center) exhibits the expected near and
+far connectivity.}
 \end{figure}
@@
   \centering 
-  \centerline{\includegraphics[width=8.5in]{celegans.ai}}
-  \caption{C elegans data}
+  \centerline{\includegraphics{celegans.ai}}
+  \caption{Discovering cell classes in the \textit{c. elegans} connectome. a.) Initial cell connectivity matrix, red are (directional) chemical synapses, blue are
+electrical synapses. Point size is proportaionl to synapse count between those cells. b.) Clustered adjaceny matrix showing identified clusters. c.) Inter-cluster
+connectivity, showing the position of each cell along the body axis as well
+as the ``true'' cell type -- both coarse (motor, sensory, and interneuron) as well
+as fine (small text labels). }
 \end{figure}
@@
   \centering 
-  \centerline{\includegraphics[width=8.5in]{mos6502.ai}}
+  \centerline{\includegraphics{mos6502.ai}}
   \caption{Microprocessor data}
@@
 \section{Discussion}
-We have presented a machine learning technique that allows cell types and connectivity structure to be discovered using only spatial and connectivity data.
-We have shown its applicability to a newly published connectomics dataset from the mouse retina, the classical connectivity of the worm c elegans, and the connectivity of a historical man-made microprocessor.
-We have found that spatial connectivity alone is sufficient to discover interesting motifs that were known to exist in the systems based on decades of previous research which had utilized extensive additional information.
-“
+We have presented a machine learning technique that allows cell types
+and connectivity structure to be discovered using only spatial and
+connectivity data. We have shown its applicability to a newly
+published connectomics dataset from the mouse retina, the classical
+connectivity of the worm c elegans, and the connectivity of a
+historical man-made microprocessor. We have found that spatial
+connectivity alone is sufficient to discover interesting motifs that
+were known to exist in the systems based on decades of previous
+research which had utilized extensive additional information. “
 
-For probabilistic models like ours, no known solution exists to exactly find the most probable parsing of the neurons into cell-types and connectivity patterns [homogenize these two wordings across the paper].
-We employ a collection of Markov-chain Monte carlo techniques [reference methods] but while different initializations converge to similar ultimate values, we can never realistically obtain the global optimum.
-There are a broad range of techniques that may offer better approximations to the global optimum (cite) and future work could adapt them to the problem we addressed here.
+For probabilistic models like ours, no known solution exists to
+exactly find the most probable parsing of the neurons into cell-types
+and connectivity patterns [homogenize these two wordings across the
+paper]. We employ a collection of Markov-chain Monte carlo techniques
+[reference methods] but while different initializations converge to
+similar ultimate values, we can never realistically obtain the global
+optimum. There are a broad range of techniques that may offer better
+approximations to the global optimum (cite) and future work could
+adapt them to the problem we addressed here.
 
-Our inference becomes slower as the amount of data increases.
-Our algorithm required 10xxx hours for 1000 neurons and would therefore take months for 10,000 neurons, which would not be acceptable.
-Various new techniques promise to allow similar calculations to take less time and scale to larger datasets (cites).
+Our inference becomes slower as the amount of data increases. Our
+algorithm required 10xxx hours for 1000 neurons and would therefore
+take months for 10,000 neurons, which would not be acceptable. Various
+new techniques promise to allow similar calculations to take less time
+and scale to larger datasets (cites).
 
@@
 
-[ how this changes things]
-There exist a range of previous approaches to the discovery of microcircuitry [coherence of language] (cites).
-These generally involve a great deal of manual labor and ad-hoc determination of what constitutes a “type” of cell -- to this day there are disagreements in the literature as to the “true” types in the mammalian retina.
-Much as phylogenomics has changed our understanding of animal ontologies, modern large scale data will allow the efficient unbiased discovery of structure.
-The sheer amount of available data demands the introduction of algorithmic approaches.
+[ how this changes things] There exist a range of previous approaches
+to the discovery of microcircuitry [coherence of language]
+(cites). These generally involve a great deal of manual labor and
+ad-hoc determination of what constitutes a “type” of cell -- to this
+day there are disagreements in the literature as to the “true” types
+in the mammalian retina. Much as phylogenomics has changed our
+understanding of animal ontologies, modern large scale data will allow
+the efficient unbiased discovery of structure. The sheer amount of
+available data demands the introduction of algorithmic approaches.
 
@@
 
-We then jointly compute the maximum a posteriroi estimate of the class
-assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and the
-global model hyperparameters $\theta$ :
+We then jointly compute the maximum a posteriroi (MAP) estimate of the
+class assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and
+the global model hyperparameters $\theta$ :
 
@@
 
-We describe three link functions, ``Logistic-distance'',
+We describe three link functions, ``Logistic-distance Bernoulli'',
+``logistic-distance fixed-lambda Bernoulli'', and ``logistid-Distance
+Poisson''.
 
-Prior on cluster assignment. While dirichlet process is inconsistent,
-we deemed this to ultimately be inconsequential for this study.
+The ``logistic-distance Bernoulli'' model associates with each latent class 
+two parameters, $\mu_{mn}$ and $\lambda_{mn}$, and the probability that two
+cells $e_i$ and $e_j$ are connected is given by
+\begin{equation}
+p^* = \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
+p = p^* \cdot (p_{max} - p_{min}) + p_{min}
+\end{equation}
+where $p_{max}$ and $p_{min}$ are per-graph hyperparameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})$. 
+Inference is performed on these hyperparameters as well, see TODO supplemental 
+for details.
 
-Priors on gridded hyperparameter space.
+We use a Dirichlet-process prior on cluster assignments, which allows
+the number of clusters to be determined automatically. In brief, the
+probability of a cell belonging to a cluster is proportional to the
+number of datapoints already in that cluster, $m_k$, such that $p(c_i
+= k) \propto \frac{m_k}{N + \alpha}$ and the probability of the cell
+belonging to a new cluster $k'$ is $p(c_i = k') \propto
+\frac{\alpha}{N + \alpha}$.
 
@@
 evaluation of the conditional assignment probabilities impossible, 
-motivating us to use an auxiliary variable method \autocite{Neal}
+motivating us to use an auxiliary variable method \autocite{Neal2000}
 in which a collection of ephemeral clusters are explicitly represented
@@
 $\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
-individual $\eta_{mn}$ are indepednent. We slice sample \autocite{neal}
+individual $\eta_{mn}$ are indepednent. We slice sample \autocite{NealSlice}
 each component's parameters, choosing the slice width as a function
@@
 
-\begin{addendum}
- \item Put acknowledgements here.
- \item[Competing Interests] The authors declare that they have no
-competing financial interests.
- \item[Correspondence] Correspondence and requests for materials
-should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
-\end{addendum}
+% \begin{addendum}
+%  \item Put acknowledgements here.
+%  \item[Competing Interests] The authors declare that they have no
+% competing financial interests.
+%  \item[Correspondence] Correspondence and requests for materials
+% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
+% \end{addendum}
+
+\printbibliography
 

### 2013-12-03 experiments/mouseretina/pipeline.py

@@ -795,6 +795,7 @@ def plot_circos_latent(exp_results,
                                           'min' : pos_min, 
                                           'max' : pos_max, 
                                           'glyph' : 'circle', 
+                                          'glyph_size' : 10, 
                                           'color' : 'black',
                                           'stroke_thickness' : 0
                                           }, 
@@ -802,7 +803,7 @@ def plot_circos_latent(exp_results,
                               {'backgrounds' : [('background', {'color': 'vvlgrey', 
                                                                 'y0' : pos_min, 
                                                                 'y1' : pos_max})],  
-                               'axes': [('axis', {'color' : 'grey', 
+                               'axes': [('axis', {'color' : 'vgrey', 
                                                   'thickness' : 1, 
                                                   'spacing' : '%fr' % ten_um_frac})]})
             

### 2013-12-04 docs/paper.tex
updates to writing; addition of todos

@@
 \usepackage[backend=biber,style=nature]{biblatex}
+\usepackage{todonotes}
 \usepackage{subfiles}
@@
 
+\todo{Citations}
+\todo{Log score plots for real data} 
+\todo{Fix vector notation}
+\todo{Get feedback on joint model score}
+\todo{Find other irm paper citation}
 % \begin{affiliations}
@@
 \section*{Introduction}
-Emerging techniques (cite Lichtman, Zador, Denk etc) promise to
+Emerging techniques \autocite{Lichtman, Zador, Denk} promise to
 quantify the location of each neuron within a volume of interest,
@@
 structure, ushering in an era of bioinformatics as an independent
-discipline (cite).
+discipline \autocite{}.
 
@@
 has been prominent in many areas, ranging from sensory (e.g. retina,
-cite) via processing (e.g. neocortex) to movement (e.g. spinal cord)
-(cites). Indeed, such motifs also exists in human-made computing
+\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal cord)
+\autocite{}. Indeed, such motifs also exists in human-made computing
 circuits. It remains an important challenge to develop algorithms to
@@
 science. Early approaches focused on global graph properties, such as
-the types of scaling present in the network (cite Strogatz scale-free
-stuff).  While this approach leads to an understanding of global
+the types of scaling present in the network \autocite {WattsStrogatz1998}
+.  While this approach leads to an understanding of global
 network, more recent work aims at identifying repeat patterns, or
-“motifs” in networks. A motif is a set of nodes with conserved
+“motifs” in networks. \todo{There's the Science2002 paper by Milo et. al. that explicitly looks at 'motifs'} A motif is a set of nodes with conserved
 relations that appears multiple times in a network, for example a set
@@
 in machine learning. Commonly used algorithms include spectral methods
-(cite), community-based-detection methods (cite), and stochastic block
-models (cite). While these approaches powerfully incorporate the
-probabilistic nature of neural connections (cite) they do not
+\autocite{}, community-based-detection methods \autocite{}, and stochastic block
+models \autocite{NowickiSnijdders2001}. While these approaches powerfully incorporate the
+probabilistic nature of neural connections \autocite{} they do not
 generally consider the known fact that the probability of connections
@@
 types and their spatial patterns of interconnection. We apply it to
-three computing systems: the mouse retina connectome (cite), the
-c. elegans connectome (cite), and a ``connectome'' of a classical
+three computing systems: the mouse retina connectome \autocite{}, the
+c. elegans connectome \autocite{}, and a ``connectome'' of a classical
 microprocessor. In all cases, we discover cell-type to cell-type
@@
 \subsubsection{Mouse Retina}
-The retina of the mouse (cite) is an example where we should expect to
-have connectivity patterns that are well approximated by our
-generative model. It is known that there are multiple classes of cells
-that can be grouped into ganglion cells that transmit information to
-the rest of the brain, bipolar cells that connect between different
-cells, and amacrine cells that feed into the ganglion cells (cite some
-review). Recent research (cite) has produced a large dataset
-containing both the types of cells from orthogonal approaches, and
-also the connectivity matrix between all reconstructed cells. We want
-to know to which level we can classify neurons into types based on
-this connectivity only.
+The retina of the mouse \autocite{Masland2001} is an example where we
+should expect to have connectivity patterns that are well approximated
+by our generative model. It is known that there are multiple classes
+of cells that can be grouped into ganglion cells that transmit
+information to the rest of the brain, bipolar cells that connect
+between different cells, and amacrine cells that feed into the
+ganglion cells (cite some review). Recent research \autocite{Helmstaedter2013} has
+produced a large dataset containing both the types of cells from
+orthogonal approaches, and also the connectivity matrix between all
+reconstructed cells. We want to know to which level we can classify
+neurons into types based on this connectivity only.
 
@@
 The roundworm Caenorhabditis elegans is a model system in
-developmental neuroscience, with the location and connectivity of each
-of 302 neurons deterministically known. This incredible cross-organism
-consistency led to early mapping of the c. elegans chemical and
-electrical (gap junction) connectome. Unlike the retina, only the
-motor neurons in c. elegans exhibit regular distribution in space, as
-they follow the body axis. Most interneurons are concentrated in
-various ganglia that project throughout the entire animal, and the
-sensory neurons are primarily located in the animal head.
+developmental neuroscience\autocite{White1986}, with the location and
+connectivity of each of 302 neurons deterministically known. This
+incredible cross-organism consistency led to early mapping of the
+c. elegans chemical and electrical (gap junction) connectome. Unlike
+the retina, only the motor neurons in c. elegans exhibit regular
+distribution in space, as they follow the body axis. Most interneurons
+are concentrated in various ganglia that project throughout the entire
+animal, and the sensory neurons are primarily located in the animal
+head.
 
@@
   \centerline{\includegraphics{celegans.ai}}
-  \caption{Discovering cell classes in the \textit{c. elegans} connectome. a.) Initial cell connectivity matrix, red are (directional) chemical synapses, blue are
-electrical synapses. Point size is proportaionl to synapse count between those cells. b.) Clustered adjaceny matrix showing identified clusters. c.) Inter-cluster
-connectivity, showing the position of each cell along the body axis as well
-as the ``true'' cell type -- both coarse (motor, sensory, and interneuron) as well
-as fine (small text labels). }
+  \caption{Discovering cell classes in the \textit{c. elegans}
+    connectome. a.) Initial cell connectivity matrix, red are
+    (directional) chemical synapses, blue are electrical
+    synapses. Point size is proportaionl to synapse count between
+    those cells. b.) Clustered adjaceny matrix showing identified
+    clusters. c.) Inter-cluster connectivity, showing the position of
+    each cell along the body axis as well as the ``true'' cell type --
+    both coarse (motor, sensory, and interneuron) as well as fine
+    (small text labels). }
 \end{figure}
@@
 classic microprocessor, the MOS Technology 6502 used in the original
-Apple II, Atari, and Commodore 64 (cite datasource). We identified a
+Apple II, Atari, and Commodore 64 \autocite{datasource}. We identified a
 region of the processor with complex but known structure containing
@@
 optimum. There are a broad range of techniques that may offer better
-approximations to the global optimum (cite) and future work could
+approximations to the global optimum \autocite{Wanga2012,FritzJonathan} and future work could
 adapt them to the problem we addressed here.
@@
 new techniques promise to allow similar calculations to take less time
-and scale to larger datasets (cites).
+and scale to larger datasets \autocite{}.
 
-*Fixed and hand selected likelihood link functions We picked the sigmoid. Should be estimated, Our small collection of parametric forms for spatial functions are obviously not exhaustive
+*Fixed and hand selected likelihood link functions We picked the
+sigmoid. Should be estimated, Our small collection of parametric forms
+for spatial functions are obviously not exhaustive
 
-*We do not use other sources of information that are totally there in the datasets. Cell morphology could be used. Laminar organization.
+*We do not use other sources of information that are totally there in
+the datasets. Cell morphology could be used. Laminar organization.
 
-*Larger datasets will generally allow algorithms to distinguish more distinct types (cite supplemental data) and more data should algorithms to get closer to the results from anatomy.
-Moreover, in general, for such problems precision increases with the size of the dataset and the 9xxx cells that we have are not sufficient to statistically distinguish all the cell types known in anatomy (71xxx).
-Still, using only connectivity it is possible to meaningfully divide neurons into classes.
+*Larger datasets will generally allow algorithms to distinguish more
+distinct types (cite supplemental data) and more data should
+algorithms to get closer to the results from anatomy. Moreover, in
+general, for such problems precision increases with the size of the
+dataset and the 9xxx cells that we have are not sufficient to
+statistically distinguish all the cell types known in anatomy
+(71xxx). Still, using only connectivity it is possible to meaningfully
+divide neurons into classes.
 
@@
 
-*Novel data analysis techniques are required to lead us into the big data age. Our approach is orthogonal to all others (because in high d everything is orthogonal).
+*Novel data analysis techniques are required to lead us into the big
+data age. Our approach is orthogonal to all others (because in high d
+everything is orthogonal).
 
-Ultimately, we will need to combine information across different sources into a joint understanding of neural function.
-Distinct cell types differ in morphology, connectivity, transcriptomics, relation to behavior or stimuli and many other ways.
-Algorithms based on each type information can then be combined to synthesize all the available information from one experiment or even across experiments into a joint model of brain function.
+Ultimately, we will need to combine information across different
+sources into a joint understanding of neural function. Distinct cell
+types differ in morphology, connectivity, transcriptomics, relation to
+behavior or stimuli and many other ways. Algorithms based on each type
+information can then be combined to synthesize all the available
+information from one experiment or even across experiments into a
+joint model of brain function.
 
@@
 
-We assume as input a collection of connectivity matrices between
-cells, as well as a distance function $d(e_i, e_j)$. We assume there
-exist an unknown number of latent (unobserved) cell types, $k$, $k \in
-\{1, 2, 3, \dots\}$, and that each cell $e_i$ belongs to a single cell
-type, $c_i = k$. The observed connectivity between two cells $R(e_i,
-e_j)$ then only depends on their latent type and their distance
-through a link function $f(c_i, c_j, d(e_i, e_j))$ If $c_i=m$ and
-$c_j=n$ then there exists a latent-class-specific set of parameters
-$\eta_{mn}$ which parametrizes $f$, as well as a set of global hyper parameters $\theta$.
+We assume as input a connectivity matrix $R$ defining the connections
+between cells, as well as a distance function $d(e_i, e_j)$
+representing a (physical) distance between adjacent cells. See the
+supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number of latent
+(unobserved) cell types, $k$, $k \in \{1, 2, 3, \dots\}$, and that
+each cell $e_i$ belongs to a single cell type, $c_i = k$. The observed
+connectivity between two cells $R(e_i, e_j)$ then only depends on
+their latent type and their distance through a link function $f(c_i,
+c_j, d(e_i, e_j))$ If $c_i=m$ and $c_j=n$ then there exists a
+latent-class-specific set of parameters $\eta_{mn}$ which parametrizes
+$f$, as well as a set of global hyper parameters $\theta$.
 
@@
 \begin{equation}
-  \P(\vec{c}, \eta, \theta) \propto \prod_{i, j} P() \prod_{m, n} P(\eta_{mn} | \theta)  P(\theta) P(c | \alpha) p(\alpha) 
+  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
 \end{equation}
@@
 Our model is a extension of infinite stochastic block models
-\cite{Kemp, other guys} to incorporate spatial constraints.
+\autocite{Kemp2006a, other guys} to incorporate spatial constraints.
+
+\todo{infinite subset relaitonal model has good reults}
 
@@
 cells $e_i$ and $e_j$ are connected is given by
-\begin{equation}
-p^* = \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
-p = p^* \cdot (p_{max} - p_{min}) + p_{min}
-\end{equation}
+\begin{eqnarray}
+p^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
+p &= & p^* \cdot (p_{max} - p_{min}) + p_{min}
+\end{eqnarray}
 where $p_{max}$ and $p_{min}$ are per-graph hyperparameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})$. 
@@
 $\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
-individual $\eta_{mn}$ are indepednent. We slice sample \autocite{NealSlice}
+individual $\eta_{mn}$ are indepednent. We slice sample \autocite{Neal2003}
 each component's parameters, choosing the slice width as a function
@@
 likelihood, scaling the counts by a factor of 4 to compensate for 
-the Poisson's underdispersion.
+the Poisson's overdispersion.
 
 \subsection{Microprocessor}
+We extracted the connection graph for all transistors in the MOS6502
+\autocite{visual6502}. Each transistor has three terminals: an input
+G, a source, and a drain, labeled C1 and C2. We identified the region
+consisting of the three general-purpose registers X, Y, and S via
+visual inspection and focused our efforts there. We created a total of
+six connectivity matrices by examining possible terminal parings. One
+graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and $e_j$ are
+connected via pins $g$ and $c_1$. 
+\todo{How do we handle symmetry here?}
 
@@
 % \end{addendum}
+\section{Supplemental}
+\subsection{Source code}
+\todo{Put up website}
+
+All source code and materials for running experiments can be
+obtained from the project website, at http://spatialrelation.github.io.
+
+\subsection{Extension to multiple graphs}
+\todo{derivation of extension to multiple graphs}
+
+\subsection{Hyperprior grids}
+\todo{Plot hyperparameter Graphs}
+
+\todo{Explain Adjusted Rand Index}
 

### 2013-12-05 docs/paper.tex
bib changes
some framing changes, still need to, you know, really work on text added hyperparams

@@
 
+\renewcommand{\vec}[1]{\mathbf{#1}}
+
 \addbibresource{connectomics motif.bib}
@@
 \todo{Log score plots for real data} 
-\todo{Fix vector notation}
-\todo{Get feedback on joint model score}
-\todo{Find other irm paper citation}
+\todo{Hierarhcical structure?}
 % \begin{affiliations}
@@
   structure. There is strong evidence for microcircuitry, where the
-  probability of connections depends on the type of pre- and
-  post-synaptic neuron and their distance. Here we thus developed a
-  nonparametric Bayesian technique that identifies type and circuitry
-  motifs from the connection data. We show that the approach recovers
+  probability of connections depends on the types of pre- and
+  post-synaptic neurons and their distance. Here we developed a
+  nonparametric Bayesian technique that identifies neuron type and circuitry
+  patterns from the connection data. We show that the approach recovers
   known neuron types in the retina, reveals interesting structure in
@@
 morphological, but they also differ in the way they connect to one
-another. In fact, the idea of well defined local connectivity motifs
+another. In fact, the idea of well defined local connectivity patterns
 has been prominent in many areas, ranging from sensory (e.g. retina,
-\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal cord)
-\autocite{}. Indeed, such motifs also exists in human-made computing
-circuits. It remains an important challenge to develop algorithms to
-use anatomical data, e.g. connectomics, to back out underlying
-connectivity motifs.
+\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal
+cord) \autocite{}. Indeed, such repeated patterns also exists in
+human-made computing circuits. It remains an important challenge to
+develop algorithms to use anatomical data, e.g. connectomics, to back
+out underlying connectivity motifs.
 
@@
 science. Early approaches focused on global graph properties, such as
-the types of scaling present in the network \autocite {WattsStrogatz1998}
-.  While this approach leads to an understanding of global
-network, more recent work aims at identifying repeat patterns, or
-“motifs” in networks. \todo{There's the Science2002 paper by Milo et. al. that explicitly looks at 'motifs'} A motif is a set of nodes with conserved
-relations that appears multiple times in a network, for example a set
-of three nodes that are each connected to one another. When it comes
-to neuroscience, we might expect that a motif consists of the kinds of
-cells that jointly make up a microcircuit along with their distance
-and cell-type dependent connectivity. The same motif should then
-appear in multiple locations in the nervous system. Combining
-motif-discovery ideas with specific knowledge about the kinds of
-motifs in the nervous system promises better structure discovery.
+the types of scaling present in the network \autocite
+{WattsStrogatz1998} .  While this approach leads to an understanding
+of global network, more recent work aims at identifying small-scale
+repeat patterns, or “motifs” in networks. \todo{There's the
+  Science2002 paper by Milo et. al. that explicitly looks at 'motifs'}
+A motif is a set of nodes with conserved relations that appears
+multiple times in a network, for example a set of three nodes that are
+each connected to one another.
+
+When it comes to neuroscience, we might
+expect that a motif consists of the kinds of cells that jointly make
+up a microcircuit along with their distance and cell-type dependent
+connectivity. The same motif should then appear in multiple locations
+in the nervous system. Combining motif-discovery ideas with specific
+knowledge about the kinds of motifs in the nervous system promises
+better structure discovery.
 
@@
 in machine learning. Commonly used algorithms include spectral methods
-\autocite{}, community-based-detection methods \autocite{}, and stochastic block
-models \autocite{NowickiSnijdders2001}. While these approaches powerfully incorporate the
-probabilistic nature of neural connections \autocite{} they do not
-generally consider the known fact that the probability of connections
-between neurons depends on distance. Combining such probabilistic
-models with a spatial component to model distance promises to improve
-the detection of microcircuitry from neural data.
+\autocite{},   community-based-detection   methods  \autocite{},   and
+stochastic block  models \autocite{NowickiSnijdders2001}.  While these
+approaches powerfully  incorporate the probabilistic nature of neural
+connections \autocite{} they do not generally consider the known fact
+that  the  probability  of  connections  between  neurons  depends  on
+distance. Combining such probabilistic models with a spatial component
+to model distance promises to  improve the detection of microcircuitry
+from neural data.
 
@@
 datasets, we obtained the spatial location and interconnectivity of a
-classic microprocessor, the MOS Technology 6502 used in the original
-Apple II, Atari, and Commodore 64 \autocite{datasource}. We identified a
-region of the processor with complex but known structure containing
-the primary general-purpose registers X, Y, and S (figure).
+transistors in a classic microprocessor, the MOS Technology 6502 (used
+in the Apple II) \autocite{datasource}. We identified a region of the
+processor with complex but known structure containing the primary
+general-purpose registers X, Y, and S (figure). Computer architects
+use common patterns of transistors when designing circuits, 
+with each transistor having a ``type'' in the circuit.
 
-Our algorithm identifies areas of spatial homogeneity 
-[mirror structure]
+Our algorithm identifies areas of spatial homogeneity that mirror the
+kown structure in the underlying architectural circuit, segementing
+transistor groups. One group contains the ``clocked'' transistors,
+one group with their source connected to ground, and one with their
+drain connected to ground. The repeat patterns of spatial connectivity
+are visible in figure~\ref{fig:microprocessor}a,
 
@@
 exactly find the most probable parsing of the neurons into cell-types
-and connectivity patterns [homogenize these two wordings across the
-paper]. We employ a collection of Markov-chain Monte carlo techniques
-[reference methods] but while different initializations converge to
-similar ultimate values, we can never realistically obtain the global
-optimum. There are a broad range of techniques that may offer better
-approximations to the global optimum \autocite{Wanga2012,FritzJonathan} and future work could
-adapt them to the problem we addressed here.
+and connectivity patterns . We employ a collection of Markov-chain
+Monte carlo techniques [reference methods] but while different
+initializations converge to similar ultimate values, we can never
+realistically obtain the global optimum. There are a broad range of
+techniques that may offer better approximations to the global optimum
+\autocite{Wanga2012,FritzJonathan} and future work could adapt them to
+the problem we addressed here.
 
@@
 Our model is a extension of infinite stochastic block models
-\autocite{Kemp2006a, other guys} to incorporate spatial constraints.
+\autocite{Kemp2006a,Xu2006} to incorporate spatial constraints.
 
-\todo{infinite subset relaitonal model has good reults}
+\todo{infinite subset relaitonal model has good summaries}
 
@@
 \subsection{Extension to multiple graphs}
-\todo{derivation of extension to multiple graphs}
+The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by simply considering the product of 
+\begin{equation}
+  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}), \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha)
+\end{equation}
+
+\subsection{Microprocessor Netlist annotation}
+\begin{figure}
+  \centering 
+    \centerline{\includegraphics[width=183mm]{mos6502/sixrelation.xysregs.ldfl.circos.4.R2.ai}}
+\end{figure}
 
 \subsection{Hyperprior grids}
-\todo{Plot hyperparameter Graphs}
+
+For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80.
+$p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99$.
+
+For the c.
+elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0.
+We globally set $rate_{min}=0.01$.
 

### 2013-12-05 experiments/mos6502/sixrelation.py

@@ -47,7 +47,7 @@ for dataset, model in [('xysregs', 'ldfl'),
     fid = open("sixrelation.%s.%s.output.html" % (dataset, model), 'w')
 
     for c_n, c in df.groupby('cluster'):
-        fid.write(c.sort('name.gate').to_html())
+        fid.write(c.sort(['name.c1']).to_html())
 
 
     f = pylab.figure(figsize=(16, 16))

### 2013-12-05 experiments/mos6502/sixrelation_circos.py

@@ -7,7 +7,7 @@ import irm
 CIRCOS_DIST_THRESHOLDS = [20, 50, 100, 200, 500]
 
 for dataset, model in [('xysregs', 'ldfl'), 
-                       #('xysregs', 'ld'), 
+                       ('xysregs', 'ld'), 
                        #('xysregs', 'ndfw'), 
                        #('xysregs', 'bb'), 
                        #('decode', 'ldfl'), 
@@ -96,6 +96,7 @@ for dataset, model in [('xysregs', 'ldfl'),
 
             circos_p = irm.plots.circos.CircosPlot(assign, 
                                                    karyotype_colors=color_str,
+                                                   ideogram_radius="0.65r",
                                                    custom_color_map=custom_color_map)
 
 
@@ -190,6 +191,7 @@ for dataset, model in [('xysregs', 'ldfl'),
                                        'label_font' : 'condensed', 
                                        'label_size' : '20p', 
                                        'label_rotate' : 'yes', 
+                                       'label_snuggle' : 'no'
                                    }, 
                               clean(df['name.c1']))
 
@@ -199,19 +201,22 @@ for dataset, model in [('xysregs', 'ldfl'),
                                        'label_font' : 'condensed', 
                                        'label_size' : '20p', 
                                        'label_rotate' : 'yes', 
+                                       'label_snuggle' : 'no'
+
                                    }, 
                               clean(df['name.c2']))
 
 
             circos_p.add_plot('text', {'r0' : '1.25r', 
-                                       'r1' : '1.35r', 
+                                       'r1' : '1.55r', 
                                        'label_font' : 'condensed', 
                                        'label_size' : '20p', 
                                        'label_rotate' : 'yes', 
+                                       'label_snuggle' : 'no'
                                    }, 
                               clean(df['name.gate']))
 
-            irm.plots.circos.write(circos_p, "sixrelation.%s.%s.circos.%d.%s.png" % (dataset, model, fi, relation))
+            irm.plots.circos.write(circos_p, "sixrelation.%s.%s.circos.%d.%s.svg" % (dataset, model, fi, relation))
 
 
             # now the tiny plots

### 2013-12-11 docs/paper.tex
Latest version of text

@@
   nonparametric Bayesian technique that identifies neuron type and circuitry
-  patterns from the connection data. We show that the approach recovers
+  patterns using only this connection data. We show that the approach recovers
   known neuron types in the retina, reveals interesting structure in
@@
 capacity of neuroanatomists to trace small circuits, this leads to
-enormous datasets that quantify aspects of the nervous system. The
-rise of high-throughput sequencing techniques necessitated the
+enormous datasets that quantify aspects of the nervous system. The impact of
+this kind of high-throughput transition has been seen before -- 
+rise of sequencing techniques necessitated the
 development of novel computational methods to understand genomic
@@
 {WattsStrogatz1998} .  While this approach leads to an understanding
-of global network, more recent work aims at identifying small-scale
-repeat patterns, or “motifs” in networks. \todo{There's the
-  Science2002 paper by Milo et. al. that explicitly looks at 'motifs'}
-A motif is a set of nodes with conserved relations that appears
-multiple times in a network, for example a set of three nodes that are
-each connected to one another.
-
-When it comes to neuroscience, we might
-expect that a motif consists of the kinds of cells that jointly make
-up a microcircuit along with their distance and cell-type dependent
-connectivity. The same motif should then appear in multiple locations
-in the nervous system. Combining motif-discovery ideas with specific
-knowledge about the kinds of motifs in the nervous system promises
-better structure discovery.
+of global network, more recent work aims at identifying very small-scale
+repeat patterns, or “motifs” in networks\autocite{Milo2002}
 
@@
 
+When it comes to neuroscience and other computing systems, we expect
+patterns of connectivity much more compelx than traditional motifs,
+exhibiting a strong spatial dependence arising from the complex
+genetic, chemical, and activity-based neural development process.
+
 Here we describe a Bayesian non-parametric model that can discover
@@
 c. elegans connectome \autocite{}, and a ``connectome'' of a classical
-microprocessor. In all cases, we discover cell-type to cell-type
+microprocessor. Note we do not explicitly consider cellular
+morphology, absolute position (cell layer), or any chemical or genetic
+information. Yet in all cases, we discover cell-type to cell-type
 connectivity that depends on spatial distance and these connectivity
@@
 perform simultaneously joint posterior inference over the structure,
-parameters, and hyperparameters of the model. This procedure yields a
-division of all neurons into cell-types along with the distance and
-type dependent connectivity probabilities. These results can then be
-compared with the results of previous anatomical studies that have
-used other aspects of neurons, such as chemical properties and
-cell-morphology to divide neurons into types.
+parameters, and hyperparameters of the model (see Methods). This
+procedure yields a division of all neurons into cell-types along with
+the distance and type dependent connectivity probabilities. These
+results can then be compared with the results of previous anatomical
+studies that have used other aspects of neurons, such as chemical
+properties and cell-morphology to divide neurons into types.
 
@@
 the retina, only the motor neurons in c. elegans exhibit regular
-distribution in space, as they follow the body axis. Most interneurons
-are concentrated in various ganglia that project throughout the entire
-animal, and the sensory neurons are primarily located in the animal
-head.
+distribution in space, being distributed along the body axis. Most
+interneurons are concentrated in various ganglia that project
+throughout the entire animal, and the sensory neurons are primarily
+located in a small number of very anteriior ganglia.
 
@@
 agreement with the classifications originally outlined by Brenner in
-1986.  [create structure that mirrors the mouse retina paragraphs]
+1986.
 
-[start paragraph with summary sentence] Motorneuron types AS, DA, and
-VA, all exclusively postsynaptic, are clustered together, as are
-motorneuron types VD and DD. VC, DB, and VB also mostly share a
-cluster. Various head motor neurons, including classes SMD, RMD, and
-RIM are clustered together. Interneurons with known
+We identify cell types that reflect the known motor/non-motor neuron
+classes, even though this system lacks the strong repeat
+microcircuitry our model was designed for.  Motorneuron types AS, DA,
+and VA, all exclusively postsynaptic, are clustered together, as are
+motorneuron types VD and DD. Traditional types VC, DB, and VB also
+mostly share a cluster. Various head motor neurons, including types
+SMD, RMD, and RIM are clustered together. Interneurons with known
 anatomically-distinct connectivity patterns, such as RIA (2 cells) or
-AVA (2 cells) are clustered into pure groups. The algorithm even
-correctly places the single-cell class DVB in a distinct group. Note
-our clustering does not perfectly recover known groupings -- several
+AVA (2 cells) are clustered into pure types. The algorithm even
+correctly places the single-cell type DVB by itself. Note our
+clustering does not perfectly reflect known divisions -- several
 combinations of head and sensory neurons are combined, and a
 difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
-grouped together. [algorithm recovers surprising structure based
-merely on connectivity even though only 278 cells and not spatially
-homogeneous]
+grouped together. Our identified cell types thus reflect a
+``coarsening'' of known types, based entirely on connectivity and
+distance information, evne when the organism exhibits substantially
+less spatial regularity than the retina.
 
@@
 To show the applicability of our method to other connectome-style
-datasets, we obtained the spatial location and interconnectivity of a
-transistors in a classic microprocessor, the MOS Technology 6502 (used
-in the Apple II) \autocite{datasource}. We identified a region of the
-processor with complex but known structure containing the primary
-general-purpose registers X, Y, and S (figure). Computer architects
-use common patterns of transistors when designing circuits, 
-with each transistor having a ``type'' in the circuit.
+datasets, we obtained the spatial location and interconnectivity of
+the transistors in a classic microprocessor, the MOS Technology 6502
+(used in the Apple II) \autocite{datasource}. Computer architects use
+common patterns of transistors when designing circuits, with each
+transistor having a ``type'' in the circuit. We identified a region of
+the processor with complex but known structure containing the primary
+8-bit registers X, Y, and S (figure).
 
 Our algorithm identifies areas of spatial homogeneity that mirror the
-kown structure in the underlying architectural circuit, segementing
-transistor groups. One group contains the ``clocked'' transistors,
-one group with their source connected to ground, and one with their
-drain connected to ground. The repeat patterns of spatial connectivity
-are visible in figure~\ref{fig:microprocessor}a,
+known structure in the underlying architectural circuit, segementing
+transistor types recognizable to computer architects. Using the
+original schematics, we see that one identified type contains the
+``clocked'' transistors, which retain digital state. Two other types
+contain transistors with pins C1 or C2 connected to ground, mostly
+serving as inverters.  An additional identified type controls the
+behavior of the three registers of interest (X, Y, and S) with respect
+to the SB data bus, either allowing them to latch or drive data from
+the bus. The repeat patterns of spatial connectivity are visible in
+figure~\ref{fig:microprocessor}a, showing the man-made horizontal and
+vertical layout of the same types of transistors.
 
@@
 We have presented a machine learning technique that allows cell types
-and connectivity structure to be discovered using only spatial and
+and microcircuitry to be discovered using only distance and
 connectivity data. We have shown its applicability to a newly
 published connectomics dataset from the mouse retina, the classical
-connectivity of the worm c elegans, and the connectivity of a
-historical man-made microprocessor. We have found that spatial
-connectivity alone is sufficient to discover interesting motifs that
-were known to exist in the systems based on decades of previous
-research which had utilized extensive additional information. “
+connectivity of the worm c. elegans, and the connectivity of a
+historical man-made microprocessor. We have found that distance and
+connectivity alone are sufficient to discover cell types and
+microcircuitry that were known to exist in the systems based on decades
+of previous research which had utilized extensive additional
+information.
 
@@
 Our inference becomes slower as the amount of data increases. Our
-algorithm required 10xxx hours for 1000 neurons and would therefore
-take months for 10,000 neurons, which would not be acceptable. Various
-new techniques promise to allow similar calculations to take less time
-and scale to larger datasets \autocite{}.
+algorithm required 30 hours for 1000 neurons and could take months for
+substantially longer datasets. Scaling this class of probabilistic
+model is an active area of research, and recent results in both
+variational methods \autocite{Hoffmann2013} and spectral learning
+  \autocite{Anandkumar2012} promise substantial speed increases.
 
-*Fixed and hand selected likelihood link functions We picked the
-sigmoid. Should be estimated, Our small collection of parametric forms
-for spatial functions are obviously not exhaustive
-
-*We do not use other sources of information that are totally there in
-the datasets. Cell morphology could be used. Laminar organization.
-
-*Larger datasets will generally allow algorithms to distinguish more
+  Larger datasets will generally allow algorithms to distinguish more
   distinct types (cite supplemental data) and more data should
@@
   general, for such problems precision increases with the size of the
-dataset and the 9xxx cells that we have are not sufficient to
+  dataset and the cells that we have are not sufficient to
   statistically distinguish all the cell types known in anatomy
-(71xxx). Still, using only connectivity it is possible to meaningfully
-divide neurons into classes.
+  (71xxx). Still, using only connectivity and distance it is possible
+  to meaningfully divide neurons into types.
 
-[ how this changes things] There exist a range of previous approaches
-to the discovery of microcircuitry [coherence of language]
-(cites). These generally involve a great deal of manual labor and
-ad-hoc determination of what constitutes a “type” of cell -- to this
-day there are disagreements in the literature as to the “true” types
-in the mammalian retina. Much as phylogenomics has changed our
-understanding of animal ontologies, modern large scale data will allow
-the efficient unbiased discovery of structure. The sheer amount of
-available data demands the introduction of algorithmic approaches.
+Our small collection of hand-selected distance-dependent likelihood
+functions are clearly non-exhaustive, and assume monotonicity
+of connectivity probability -- for a given class, closer cells
+are never less-likely to connect. This is known to be insufficient
+for various neural systems. Future models could incorporate
+a wider variety of likelihood functions, or even learn the global
+functional form from the data.
 
-*A new way of phenotyping, combining spatial motif ideas with clustering ideas
+There have existed a a range of previous approaches to the discovery
+of neural microcircuitry (cites). These generally involve a great deal
+of manual labor and ad-hoc determination of what constitutes a “type”
+of cell -- to this day there are disagreements in the literature as to
+the “true” types in the mammalian retina. Much as phylogenomics has
+changed our understanding of animal ontologies, modern large scale
+data will allow the efficient unbiased discovery of cell types and
+circuits. The sheer amount of available data demands the introduction
+of algorithmic approaches.
 
-*Novel data analysis techniques are required to lead us into the big
-data age. Our approach is orthogonal to all others (because in high d
-everything is orthogonal).
+The development of automatic identification and quantification of cell
+type may also provide new phenotypic tools for quantifying the effect
+of disease, genetic abnormalities, and genetic
+interventions. Particular pathologies and genetic manipulations may
+subtly change the number and connectivity of cell types in ways
+only visible with algorithmic analysis.
 
@@
 
-We assume as input a connectivity matrix $R$ defining the connections
-between cells, as well as a distance function $d(e_i, e_j)$
+We take as input a connectivity matrix $R$ defining the connections
+between cell $e_i$ and $e_j$, as well as a distance function $d(e_i, e_j)$
 representing a (physical) distance between adjacent cells. See the
 supplemental material for extension to multiple connectivity
-matrices. We assume there exist an unknown number of latent
-(unobserved) cell types, $k$, $k \in \{1, 2, 3, \dots\}$, and that
+matrices. We assume there exist an unknown number $k$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots\}$, and that
 each cell $e_i$ belongs to a single cell type, $c_i = k$. The observed
-connectivity between two cells $R(e_i, e_j)$ then only depends on
-their latent type and their distance through a link function $f(c_i,
-c_j, d(e_i, e_j))$ If $c_i=m$ and $c_j=n$ then there exists a
-latent-class-specific set of parameters $\eta_{mn}$ which parametrizes
-$f$, as well as a set of global hyper parameters $\theta$.
+connectivity between two cells $R(e_i, e_j)$ then depends only on
+their latent type and their distance through a link function $f(\cdot, d(e_i, e_j))$.
+We assume $f$ is parameterized based on the latent class, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well as a set of global hyper parameters $\theta$, such that the link function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
 
-We then jointly compute the maximum a posteriroi (MAP) estimate of the
+We then jointly infer the maximum a posteriroi (MAP) estimate of the
 class assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and
@@
 We describe three link functions, ``Logistic-distance Bernoulli'',
-``logistic-distance fixed-lambda Bernoulli'', and ``logistid-Distance
+``logistic-distance fixed-lambda Bernoulli'', and ``logistic-Distance
 Poisson''.
 
-The ``logistic-distance Bernoulli'' model associates with each latent class 
-two parameters, $\mu_{mn}$ and $\lambda_{mn}$, and the probability that two
-cells $e_i$ and $e_j$ are connected is given by
+% fixme class component, NOT class
+The ``logistic-distance Bernoulli'' link model assumes that, if cell
+$e_i$ is of class $m$ and cell $e_j$ is of class $n$, then $\eta_{mn}
+= (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
+and $e_j$ are connected is given by
 \begin{eqnarray}
@@
 where $p_{max}$ and $p_{min}$ are per-graph hyperparameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})$.
-Inference is performed on these hyperparameters as well, see TODO supplemental 
-for details.
 
@@
 \autocite{Helmstaedter2013} yielded a listing of places where neurons
-come into contact. There were $number$ cells originally, and selected
+come into contact. There were over 1000 cells originally, and selected
 the $950$ for which the location of the soma could be reconstruted
@@
 the reconstructed soma centers, and used the Logistic-Distance
-spatial relation. Hyperprior distributions are shown in \ref{supplemental}.
+spatial relation. Hyperprior griddings are shown in \ref{supplemental}.
 
@@
 graph. We use the synapse counts with the logistic-distance poisson
-likelihood, scaling the counts by a factor of 4 to compensate for 
+likelihood, scaling the counts by 4.0 to compensate for 
 the Poisson's overdispersion.
@@
 \subsection{Microprocessor}
-We extracted the connection graph for all transistors in the MOS6502
-\autocite{visual6502}. Each transistor has three terminals: an input
-G, a source, and a drain, labeled C1 and C2. We identified the region
-consisting of the three general-purpose registers X, Y, and S via
-visual inspection and focused our efforts there. We created a total of
-six connectivity matrices by examining possible terminal parings. One
-graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and $e_j$ are
-connected via pins $g$ and $c_1$. 
-\todo{How do we handle symmetry here?}
+We extracted the connection graph for the transistors in the MOS6502
+\autocite{visual6502}. Each transistor has three terminals (gate,
+source, drain), but the methods of the original dataset were unable to
+consistently resolve which of the C1 and C2 terminals were source and
+drain, leading to ambiguity in our encoding. We identified a region
+consisting of consisting of three registers X, Y, and S via visual
+inspection and focused our efforts there. We created a total of six
+connectivity matrices by examining possible terminal parings. One
+graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
+$e_j$ are connected via pins $g$ and $c_1$.
 
@@
 \begin{equation}
-  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}), \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha)
+  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}, \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha)
 \end{equation}
@@
 \subsection{Microprocessor Netlist annotation}
-\begin{figure}
-  \centering 
-    \centerline{\includegraphics[width=183mm]{mos6502/sixrelation.xysregs.ldfl.circos.4.R2.ai}}
-\end{figure}
+This is a matrix consisting of the different clusters of nets for 
+the registers.
 
@@
 For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80.
-$p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99$.
+$p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99]$.
 
 For the c.
-elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0.
+elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0.
 We globally set $rate_{min}=0.01$.
+
+For the microprocessor with the Logistic Distance fixed lambda Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced points between 10 and 500 and set $\lambda=\mu_{hp}/10$.
+$p_{min} \in \{0.001, 0.01, 0.02\}$  and both $p_\alpha$ and $p_\beta \in \{0.1, 1.0, 2.0\}$.
 

### 2013-12-11 experiments/mos6502/sixrelation_circos.py

@@ -7,7 +7,7 @@ import irm
 CIRCOS_DIST_THRESHOLDS = [20, 50, 100, 200, 500]
 
 for dataset, model in [('xysregs', 'ldfl'), 
-                       ('xysregs', 'ld'), 
+                       #('xysregs', 'ld'), 
                        #('xysregs', 'ndfw'), 
                        #('xysregs', 'bb'), 
                        #('decode', 'ldfl'), 
@@ -19,6 +19,16 @@ for dataset, model in [('xysregs', 'ldfl'),
     raw_data = pickle.load(open("data.pickle", 'r'))
     wiredf = raw_data['wiredf']
     tfdf = raw_data['tfdf']
+
+    null_names = wiredf['name'].isnull()
+    wiredf['name'][null_names] = ["n%d" % d for d in range(np.sum(null_names))]
+    # translations
+    wire_translations = {"dasb0": "sb0", 
+                         "dasb4" : "sb4"}
+    for w_old, w_new in wire_translations.iteritems():
+
+        wiredf['name'][wiredf['name'] == w_old] = w_new
+
 
     x = pickle.load(open("typed.%s.region.pickle" % dataset))
     adj_mat = x['adj_mat']
@@ -49,7 +59,7 @@ for dataset, model in [('xysregs', 'ldfl'),
     fid = open("sixrelation.%s.%s.output.html" % (dataset, model), 'w')
 
     for c_n, c in df.groupby('cluster'):
-        fid.write(c.sort('name.gate').to_html())
+        fid.write(c.sort(['name.gate', 'name.c1']).to_html())
 
     model_name = data['relations']['R1']['model']
 
@@ -189,7 +199,7 @@ for dataset, model in [('xysregs', 'ldfl'),
             circos_p.add_plot('text', {'r0' : '1.05r', 
                                        'r1' : '1.15r', 
                                        'label_font' : 'condensed', 
-                                       'label_size' : '20p', 
+                                       'label_size' : '10p', 
                                        'label_rotate' : 'yes', 
                                        'label_snuggle' : 'no'
                                    }, 
@@ -199,7 +209,7 @@ for dataset, model in [('xysregs', 'ldfl'),
             circos_p.add_plot('text', {'r0' : '1.15r', 
                                        'r1' : '1.25r', 
                                        'label_font' : 'condensed', 
-                                       'label_size' : '20p', 
+                                       'label_size' : '10p', 
                                        'label_rotate' : 'yes', 
                                        'label_snuggle' : 'no'
 
[…]

### 2013-12-11 experiments/mos6502/sixrelation_report.py

@@ -1,0 +1,106 @@
+import numpy as np
+import cPickle as pickle
+from matplotlib import pylab
+from matplotlib.backends.backend_pdf import PdfPages
+import irm    
+import pandas
+
+CIRCOS_DIST_THRESHOLDS = [20, 50, 100, 200, 500]
+
+for dataset, model in [('xysregs', 'ldfl'), 
+                       #('xysregs', 'ld'), 
+                       #('xysregs', 'ndfw'), 
+                       #('xysregs', 'bb'), 
+                       #('decode', 'ldfl'), 
+                       #('decode', 'ld'), 
+                       #('lower', 'ldfl'), 
+                       #('lower', 'ld')
+]:
+
+    raw_data = pickle.load(open("data.pickle", 'r'))
+    wiredf = raw_data['wiredf']
+    tfdf = raw_data['tfdf']
+
+    null_names = wiredf['name'].isnull()
+    wiredf['name'][null_names] = ["n%d" % d for d in range(np.sum(null_names))]
+    # translations
+    wire_translations = {"dasb0": "sb0", 
+                         "dasb4" : "sb4"}
+    for w_old, w_new in wire_translations.iteritems():
+
+        wiredf['name'][wiredf['name'] == w_old] = w_new
+
+
+    x = pickle.load(open("typed.%s.region.pickle" % dataset))
+    adj_mat = x['adj_mat']
+    #typed_adjmat = pickle.load(open("typed.adjmat.pickle"))
+    #pin_pairs = typed_adjmat['pin_pairs'] 
+
+    # Should really load these from above file
+
+    pin_pairs = [('gate', 'gate'), ('gate', 'c1'), ('gate', 'c2'), 
+                 ('c1', 'c1'), ('c1', 'c2'), ('c2', 'c2')]
+
+
+    RELATION_LIST = ["%s_%s" % a for a in pin_pairs]
+
+    sample = pickle.load(open("data/mos6502.typed.%s.%s.data-fixed_100_200-anneal_slow_400.0.latent.pickle" % (dataset, model),  'r'))
+    data = pickle.load(open("data/mos6502.typed.%s.%s.data" % (dataset, model),  'r'))
+
+
+
+
+    df = x['subdf']
+    print x['region']
+    assign = np.array(sample['domains']['d1']['assignment'])
+
+    class_ids = np.unique(assign)
+    CLASS_N = len(class_ids)
+
+
+    custom_color_map = {}
[…]

### 2013-12-11 docs/paper.tex
updates to text

@@
 \usepackage{multicol}
+\usepackage{csquotes}
+\usepackage{pdfpages}
 
@@
 \section*{Introduction}
-Emerging techniques \autocite{Lichtman, Zador, Denk} promise to
+Emerging techniques \autocite{Morgan2013,Zador2012} promise to
 quantify the location of each neuron within a volume of interest,
@@
 \autocite{},   community-based-detection   methods  \autocite{},   and
-stochastic block  models \autocite{NowickiSnijdders2001}.  While these
+stochastic block  models \autocite{Nowicki2001}.  While these
 approaches powerfully  incorporate the probabilistic nature of neural
@@
 c. elegans connectome \autocite{}, and a ``connectome'' of a classical
-microprocessor. Note we do not explicitly consider cellular
-morphology, absolute position (cell layer), or any chemical or genetic
-information. Yet in all cases, we discover cell-type to cell-type
-connectivity that depends on spatial distance and these connectivity
-functions depend on the two involved cell types. Comparing the cell
-types discovered by the algorithms with those obtained by anatomists
-based on orthogonal information reveals a high degree of agreement. We
-present a scalable approach to infer microcircuitry from connectivity
-data.
+microprocessor \autocite{James2010}. Note we do not explicitly
+consider cellular morphology, absolute position (cell layer), or any
+chemical or genetic information. Yet in all cases, we discover
+cell-type to cell-type connectivity that depends on spatial distance
+and these connectivity functions depend on the two involved cell
+types. Comparing the cell types discovered by the algorithms with
+those obtained by anatomists based on orthogonal information reveals a
+high degree of agreement. We present a scalable approach to infer
+microcircuitry from connectivity data.
 
@@
 connectivtiy patterns.}
-
+\label{fig:overview}
 \end{figure}
@@
 procedure yields a division of all neurons into cell-types along with
-the distance and type dependent connectivity probabilities. These
-results can then be compared with the results of previous anatomical
-studies that have used other aspects of neurons, such as chemical
-properties and cell-morphology to divide neurons into types.
+the distance and type dependent connectivity probabilities (see
+fig~\ref{fig:overview}). These results can then be compared with the
+results of previous anatomical studies that have used other aspects of
+neurons, such as chemical properties and cell-morphology to divide
+neurons into types.
 
@@
 one we used for simulation. We find that the model does a good job of
-recovering the structure we used to generate the data (Fig
-1A). Indeed, it does far better than a simpler and well established
+recovering the structure we used to generate the data (fig~\ref{fig:synthetic}a)
+Indeed, it does far better than a simpler and well established
 model (stochastic block) that assumes that there is no distance
@@
 converge towards a highly-probable solution. }
-
+\label{fig:synthetic}
 \end{figure}
@@
 The algorithm took 8 hours to converge to a locally optimal solution,
-dividing neurons into a set of cell types (Fig 2a). For each pair of
-neurons there is a specific distance dependent connection probability
-(Fig 2b), which is well approximated by the model fit. Moreover, each
-type of cell is rather isotropically distributed across space (Fig 2c)
-as should be expected for true cell types. The algorithm is thus able
-to extract meaningful structure from the data.
+dividing neurons into a set of cell types (fig~\ref{fig:retina}c, each
+wedge is a type). For each pair of neurons there is a specific
+distance dependent connection probability (fig~\ref{fig:retina}b,c,d),
+which is well approximated by the model fit. Moreover, each type of
+cell is rather isotropically distributed across space
+(fig~\ref{fig:retina}e) as should be expected for true cell types. The
+algorithm is thus able to extract meaningful structure from the data.
 
 Comparing the results of the algorithm to other information sources
-allows evaluating the quality of the type determination. We find that
-the types tend to reflect the known laminar distribution in the retina
-(Fig 2a, outer ring), which is exciting as the algorithm did not
-actually use this type of information. Moreover, we can compare the
-results from our algorithm with cell-type determinations by
-professional anatomists, who distinguish between 71 types of
-neurons. The algorithm yields a separation of neurons into a smaller
-number of types that are, however, highly correlated with those of the
-anatomists.
-
-Our types closely reflect the (human-determined) segmentation of cells
-into retinal ganglion, narrow amacrine, medium/wide amacrine, and
-bipolar cells (Figure 2a, outermost ring). [TODO: Do we actually
-assign any of the unknown cells correctly? are types 72-79 “differet”
-or simply unknown?]
+allows evaluating the quality of the type determination. Our types
+closely reflect the (anatomist-determined) segmentation of cells into
+retinal ganglion, narrow amacrine, medium/wide amacrine, and bipolar
+cells (Figure 2a, outermost ring). We find that the types we find tend
+to reflect the known laminar distribution in the retina
+(fig~\ref{fig:retina}c, middle ring), which is exciting as the
+algorithm did not actually use this type of information. The algorithm
+yields a separation of neurons into a smaller number of types than the fully
+granular listing of 71 types found by the original authors of the paper, 
+although is still highly correlated with those finer type distinctions (see supp~\ref{supp:mouseretina}. It is our expectation that, with larger datasets, 
+even closer agreement would be found.  [TODO:
+Do we assign any of the unknown cells correctly?  are types 72-79
+“different” or simply unknown?]
 
@@
 far connectivity.}
+\label{fig:mouseretina}
 \end{figure}
@@
 determined the underlying clusters explained by connectivity and
-distance (fig ba). A superficial inspection of the results shows
+distance (fig~ref{fig:celegans}a). A superficial inspection of the results shows
 clustering into groups consisting roughly homogeneously of motor
@@
   \centerline{\includegraphics{celegans.ai}}
-  \caption{Discovering cell classes in the \textit{c. elegans}
+  \caption{Discovering cell types in the \textit{c. elegans}
     connectome. a.) Initial cell connectivity matrix, red are
@@
     (small text labels). }
+  \label{fig:celegans}
 \end{figure}
@@
 the transistors in a classic microprocessor, the MOS Technology 6502
-(used in the Apple II) \autocite{datasource}. Computer architects use
+(used in the Apple II) \autocite{James2010}. Computer architects use
 common patterns of transistors when designing circuits, with each
@@
   \centerline{\includegraphics{mos6502.ai}}
-  \caption{Microprocessor data}
+  \caption{Discovering transistor types in the MOS 6502
+    microprocessor. a.) The clusters and connectivity versus distance
+    for connections between Gate and C1, Gate and C2, and C1 and C2
+    terminals on a transistor. Red and teal types have a terminal
+    pulled down to ground and mostly function as inverters. Purple
+    class are clocked, stateful transistors, orange control the ALU
+    and yellow control the special data bus. b.) The spatial
+    distribution of the transistors in each cluster. }
+\label{fig:mos6502}
 \end{figure}
@@
 model is an active area of research, and recent results in both
-variational methods \autocite{Hoffmann2013} and spectral learning
+variational methods \autocite{Hoffman2013} and spectral learning
   \autocite{Anandkumar2012} promise substantial speed increases.
@@
 
-\section{Methods}
-\section{Methods Summary}
-
-% this should clearly describe the inputs
-
-% the outputs
-
-% the model
+\newpage
+\section{Methods }
+\subsection{Methods Summary}
 
@@
 Our model is a extension of infinite stochastic block models
-\autocite{Kemp2006a,Xu2006} to incorporate spatial constraints.
+\autocite{Kemp2006a,Xu2006} to incorporate spatial constraints,
+inspired by attempts to extend these models with arbitrary
+discriminative functions\autocite{Murphy2012}.
 
-\todo{infinite subset relaitonal model has good summaries}
-
-We describe three link functions, ``Logistic-distance Bernoulli'',
-``logistic-distance fixed-lambda Bernoulli'', and ``logistic-Distance
-Poisson''.
+We describe the spatial ``Logistic-distance Bernoulli''  function here,
+and others in the supplemental material.
 
 % fixme class component, NOT class
-The ``logistic-distance Bernoulli'' link model assumes that, if cell
+The ``logistic-distance Bernoulli'' spatial model assumes that, if cell
 $e_i$ is of class $m$ and cell $e_j$ is of class $n$, then $\eta_{mn}
@@
 We extracted the connection graph for the transistors in the MOS6502
-\autocite{visual6502}. Each transistor has three terminals (gate,
+\autocite{visual6502source}. Each transistor has three terminals (gate,
 source, drain), but the methods of the original dataset were unable to
@@
 % \end{addendum}
+\newpage
 \section{Supplemental}
+
+\subsection{Other Likelihoods}
+\label{supp:otherlikelihoods}
+ - Reparameterized logistic-distance bernoulli with fixed lambda
+
+ - Logistic-distance poisson
+
 \subsection{Source code}
-\todo{Put up website}
 
@@
 \subsection{Extension to multiple graphs}
-The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by simply considering the product of 
+\label{supp:multigraph}
+The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by extending the likelihood to include the product of the likelihoods of the individual graphs.
+
 \begin{equation}
@@
 \subsection{Microprocessor Netlist annotation}
+\label{supp:mos6502netlist}
 This is a matrix consisting of the different clusters of nets for 
-the registers.
+the registers, with the colors consistent with figure~\ref{mos6502}
+
+\includepdf[pages=-]{mos6502.clusters.pdf}
 
 \subsection{Hyperprior grids}
+\label{supp:hyperpriors}
 

### 2013-12-11 experiments/mouseretina/pipeline.py

@@ -592,6 +592,8 @@ def plot_best_cluster_latent(exp_results,
         util.plot_latent(sample_latent, dist_matrix, latent_fname, 
                          model = data['relations']['R1']['model'],
                          PLOT_MAX_DIST=150.0, MAX_CLASSES=20)
+
+
         
 @transform(get_results, suffix(".samples"), [".hypers.pdf"])
 def plot_hypers(exp_results, (plot_hypers_filename,)):
@@ -943,6 +945,204 @@ def plot_clustered_somapos(exp_results,
 
     f.savefig(out_filename)
 
+@transform(get_results, suffix(".samples"), 
+           ".truth_latent.pdf" )
+def plot_truth_latent(exp_results, 
+                      out_filename):
+
+    sample_d = pickle.load(open(exp_results))
+    chains = sample_d['chains']
+    
+    exp = sample_d['exp']
+    data_filename = exp['data_filename']
+    data = pickle.load(open(data_filename))
+    data_basename, _ = os.path.splitext(data_filename)
+    meta = pickle.load(open(data_basename + ".meta"))
+
+    meta_infile = meta['infile']
+    print "meta_infile=", meta_infile
+
+    d = pickle.load(open(meta_infile, 'r'))
+    conn = d['dist_matrix']['link']
+    cell_id_permutation = d['cell_id_permutation']
+    
+    dist_matrix = d['dist_matrix']
+    orig_data = pickle.load(open(d['infile']))
+    cell_types = d['types'][:len(conn)]
+    
+    type_metadata_df = pickle.load(open("type_metadata.pickle", 'r'))['type_metadata']
+
+    chains = [c for c in chains if type(c['scores']) != int]
+    CHAINN = len(chains)
+
+    chains_sorted_order = np.argsort([d['scores'][-1] for d in chains])[::-1]
+    chain_pos = 0
+
+    best_chain_i = chains_sorted_order[chain_pos]
+    best_chain = chains[best_chain_i]
+    sample_latent = best_chain['state']
+    cell_assignment = np.array(sample_latent['domains']['d1']['assignment'])
+
+    soma_positions = pickle.load(open('soma.positions.pickle', 'r'))
+    pos_vec = soma_positions['pos_vec'][cell_id_permutation]
+    print "Pos_vec=", pos_vec
+    model_name = data['relations']['R1']['model']
+
+    # this is potentially fun: get the ranges for each type
[…]

### 2013-12-24 docs/paper.tex
making a finite version for comparison

@@
 \usepackage{pdfpages}
+\usepackage{placeins}
 
@@
 
-\todo{Citations}
-\todo{Log score plots for real data} 
-\todo{Hierarhcical structure?}
 % \begin{affiliations}
@@
   necessitating new analysis methods to discover underlying
-  structure. There is strong evidence for microcircuitry, where the
-  probability of connections depends on the types of pre- and
-  post-synaptic neurons and their distance. Here we developed a
-  nonparametric Bayesian technique that identifies neuron type and circuitry
-  patterns using only this connection data. We show that the approach recovers
+  structure. It has long been assumed that discovering the local patterns of microcircuitry is the key to understanding neural function. Here we developed a
+  nonparametric Bayesian technique that identifies neuron types and circuitry
+  patterns in connectomics data. We show that the approach recovers
   known neuron types in the retina, reveals interesting structure in
   the nervous system of c. elegans, and automatically discovers the
-  structure of microprocessors. Our approach is a first step towards
-  condensing connectomics data into meaningful, human readable
-  structure.
+  structure of microprocessors. Our approach extracts structural meaning from connectomics, enabling new approaches of deriving anatomical insights from these emerging datasets.
+
 \end{abstract}
@@
 When it comes to neuroscience and other computing systems, we expect
-patterns of connectivity much more compelx than traditional motifs,
+patterns of connectivity much more complex than traditional motifs,
 exhibiting a strong spatial dependence arising from the complex
@@
 at various distances, illustrating short-range and long-range
-connectivtiy patterns.}
+connectivity patterns.}
 \label{fig:overview}
@@
 parameters, and hyperparameters of the model (see Methods). This
-procedure yields a division of all neurons into cell-types along with
+procedure automatically learns the number of cell types, which cells 
+belong to which types, along with the 
 the distance and type dependent connectivity probabilities (see
@@
 one we used for simulation. We find that the model does a good job of
-recovering the structure we used to generate the data (fig~\ref{fig:synthetic}a)
-Indeed, it does far better than a simpler and well established
-model (stochastic block) that assumes that there is no distance
-dependence to the connectivity. The model converges relatively
-quickly, which is enabled by xxx cool tricks we used to speed it
-up. It appears that the model behaves the way it should and allows
-application to large datasets.
+recovering the structure we used to generate the data
+(fig~\ref{fig:synthetic}a) Indeed, it does far better than a simpler
+and well established model (the infinite stochastic block model, or
+ISBM) that assumes that there is no distance dependence to the
+connectivity. The model converges relatively quickly to an estimate of
+the most probable values for the cell types, which is enabled by using
+a combination of simulated annealing and parallelized Markov chain
+Monte Carlo (see methods for details). It appears that the model
+behaves the way it should and allows application to relatively large
+datasets.
 
@@
   \caption{Correct recovery of true numbers of hidden classes
-in synthetic data. a.) Other models overcluster the number
+in synthetic data. a.) Other models over-cluster the number
 of classes as they fail to take distance into account, whereas
@@
 different true class counts, between our model and a traditional
-nonparametric stochastic block model. c.) Multiple Markov chains iterateively
-converge towards a highly-probable solution. }
+ISBM. c.) For synthetic data, 
+the IRM over clusters, finding a large number of spatially-localized types. 
+In contrast, the spatial-relational model recovers the true spatial
+extent of the underlying types.}
 \label{fig:synthetic}
@@
 The algorithm took 8 hours to converge to a locally optimal solution,
-dividing neurons into a set of cell types (fig~\ref{fig:retina}c, each
+dividing neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
 wedge is a type). For each pair of neurons there is a specific
-distance dependent connection probability (fig~\ref{fig:retina}b,c,d),
+distance dependent connection probability (fig~\ref{fig:mouseretina}b,c,d),
 which is well approximated by the model fit. Moreover, each type of
 cell is rather isotropically distributed across space
-(fig~\ref{fig:retina}e) as should be expected for true cell types. The
+(fig~\ref{fig:mouseretina}e) as should be expected for true cell types. The
 algorithm is thus able to extract meaningful structure from the data.
@@
 retinal ganglion, narrow amacrine, medium/wide amacrine, and bipolar
-cells (Figure 2a, outermost ring). We find that the types we find tend
-to reflect the known laminar distribution in the retina
-(fig~\ref{fig:retina}c, middle ring), which is exciting as the
-algorithm did not actually use this type of information. The algorithm
-yields a separation of neurons into a smaller number of types than the fully
-granular listing of 71 types found by the original authors of the paper, 
-although is still highly correlated with those finer type distinctions (see supp~\ref{supp:mouseretina}. It is our expectation that, with larger datasets, 
-even closer agreement would be found.  [TODO:
-Do we assign any of the unknown cells correctly?  are types 72-79
-“different” or simply unknown?]
+cells (fig~\ref{fig:mouseretina}c, outermost ring). We find that the
+types we find tend to reflect the known laminar distribution in the
+retina (fig~\ref{fig:mouseretina}c, middle ring), which is exciting as
+the algorithm did not actually use this type of information.
+
+The algorithm yields a separation of neurons into a smaller number of
+types than the fully granular listing of 71 types found by the
+original authors of the paper, although is still highly correlated
+with those finer type distinctions (see
+supp~\ref{supp:mouseretina}. It is our expectation that, with larger
+datasets, even closer agreement would be found.
+
+\todo{Do we assign any of the unknown cells correctly?  are types 72-79
+“different” or simply unknown?}
+
+We compared the results of our model to the traditional ISBM to see if
+the addition of spatial information was necessary. We find that, as expected from our synthetic experiments, the
+IRM massively over-segmented the data, finding
+spatially-localized “communities” of cells which bore little relation
+to known types.  The average spatial extent of the types found by the ISBM
+was  $14.6\pm8.1 \mu m$. However, in the ground truth by
+anatomists the average distance was $34.3\pm7.6 \mu m$ and $35.3\pm 2.4 \mu m$, quite
+close to the $34.5\pm 1.0 \mu m$ discovered by our algorithm. Modeling the
+distance-dependence prevents over-segmentation and thus allows more
+meaningful estimates.
+
+% from experiments/mouseretina/spatial_var.txt
+
+% Relational Model: mean = 14.586508 std=8.136334
+% Spatial-Relational Model: mean = 34.540355 std=1.069483
+% Truth (fine): mean = 34.348443 std=7.567564
+% Truth (coarse): mean = 35.276636 std=2.372951
 
@@
   \caption{Discovering cell classes in the mouse retina connectome. 
-a.) Input adajacency data for 950 cells for which soma positions were known. b. clsutered adjacency matrix. c.) connectivity diagram showing our clusters, as
+a.) Input adjacency data for 950 cells for which soma positions were known. b. clustered adjacency matrix. c.) connectivity diagram showing our clusters, as
 well as the cell depth and known cell type of each cell. Our model agrees
@@
 laminar-specific connectivity patterns. d.) Connectivity between our
-clusters as a function of distance -- the cluster consisting primarially of
+clusters as a function of distance -- the cluster consisting primarily of
 retinal ganglion cells (lower left-center) exhibits the expected near and
@@
 throughout the entire animal, and the sensory neurons are primarily
-located in a small number of very anteriior ganglia.
+located in a small number of anterior ganglia.
 
@@
 classes, even though this system lacks the strong repeat
-microcircuitry our model was designed for.  Motorneuron types AS, DA,
+microcircuitry our model was designed for.  Motor-neuron types AS, DA,
 and VA, all exclusively postsynaptic, are clustered together, as are
-motorneuron types VD and DD. Traditional types VC, DB, and VB also
+motor-neuron types VD and DD. Traditional types VC, DB, and VB also
 mostly share a cluster. Various head motor neurons, including types
@@
 ``coarsening'' of known types, based entirely on connectivity and
-distance information, evne when the organism exhibits substantially
+distance information, even when the organism exhibits substantially
 less spatial regularity than the retina.
@@
     (directional) chemical synapses, blue are electrical
-    synapses. Point size is proportaionl to synapse count between
-    those cells. b.) Clustered adjaceny matrix showing identified
+    synapses. Point size is proportional to synapse count between
+    those cells. b.) Clustered adjacency matrix showing identified
     clusters. c.) Inter-cluster connectivity, showing the position of
@@
 Our algorithm identifies areas of spatial homogeneity that mirror the
-known structure in the underlying architectural circuit, segementing
+known structure in the underlying architectural circuit, segmenting
 transistor types recognizable to computer architects. Using the
@@
   statistically distinguish all the cell types known in anatomy
-  (71xxx). Still, using only connectivity and distance it is possible
+  ($\sim 70$). Still, using only connectivity and distance it is possible
   to meaningfully divide neurons into types.
@@
 
-There have existed a a range of previous approaches to the discovery
-of neural microcircuitry (cites). These generally involve a great deal
+There have existed a range of previous approaches to the discovery
+of neural microcircuitry \autocite{} These generally involve a great deal
 of manual labor and ad-hoc determination of what constitutes a “type”
@@
 The development of automatic identification and quantification of cell
-type may also provide new phenotypic tools for quantifying the effect
-of disease, genetic abnormalities, and genetic
-interventions. Particular pathologies and genetic manipulations may
-subtly change the number and connectivity of cell types in ways
-only visible with algorithmic analysis.
+type may also provide a new computational phenotype for quantifying the effect
+of disease, genetic interventions, and developmentally-experienced neural activity.
+Our method can in principle identify neuron-types across non-connected graphs, e.g.
+across animals.
+For example, the types of neurons in one animal can be associated with the types of neurons in another animal, in the same way as this is already possible through molecular markers \autocite{}.
+This could be particularly important if cell types appear that are due to properties of the stimuli and experience as opposed to the molecular properties of cells \autocite{}.
+This would allow comparative quantitative anatomy across animals, and the search for the causes of the ultimate connectivity.
 
@@
 joint model of brain function.
+
+\printbibliography
 
@@
 
-We then jointly infer the maximum a posteriroi (MAP) estimate of the
+We then jointly infer the maximum a posteriori (MAP) estimate of the
 class assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and
@@
 % methods section can have no figures or tables [grrr]
-% todo describe multiple independent relations?
 
@@
 
-Our model is a extension of infinite stochastic block models
-\autocite{Kemp2006a,Xu2006} to incorporate spatial constraints,
+Our model is a extension of the ISBM
+\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
 inspired by attempts to extend these models with arbitrary
@@
 $\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
-individual $\eta_{mn}$ are indepednent. We slice sample \autocite{Neal2003}
+individual $\eta_{mn}$ are independent. We slice sample \autocite{Neal2003}
 each component's parameters, choosing the slice width as a function
@@
 To pick the MAP, we run up to $100$ simultaneous chains in parallel,
-iniitalized from different random initial points in the statespace,
+initialized from different random initial points in the state space,
 and then pick the chain with the highest log likelihood. Early
@@
 estimate of the posterior but were abandoned due to the difficulty of
-summarizing the posterior over such a complex state space, whcih is an
+summarizing the posterior over such a complex state space, which is an
 ongoing area of research in the probabilistic modeling community
@@
 come into contact. There were over 1000 cells originally, and selected
-the $950$ for which the location of the soma could be reconstruted
+the $950$ for which the location of the soma could be reconstructed
 from the provided cell plots (soma locations were not provided by the
-study's authors in machine-readble form). Ultimately
+study's authors in machine-readable form). Ultimately
 this left a matrix between the total synapse-like contact area between
@@
 cell's position was its distance along the anterior-posterior axis
-normalzed between zero and one. We used both networks, the chemical
+normalized between zero and one. We used both networks, the chemical
 network as a directed graph and the electrical network as undirected
@@
 inspection and focused our efforts there. We created a total of six
-connectivity matrices by examining possible terminal parings. One
+connectivity matrices by examining possible terminal pairings. One
 graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
@@
 
+\FloatBarrier
+
 \subsection{Microprocessor Netlist annotation}
@@
 
+\FloatBarrier
+\subsection{Comparison to IRM for Mouse Retina}
+
+\begin{figure}[h!]
+  \centering 
+    \centerline{\includegraphics[width=5in]{mouseretina/spatial_var.pdf}}
+  \caption{}
+\label{fig:supp:irmspatialcomparemouse}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering 
+    \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
+  \caption{}
+\label{fig:supp:retinabbclusters}
+\end{figure}
+
+\begin{figure}[h!]
+
+  \centering 
+    \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
+  \caption{}
+\label{fig:supp:retinaldclusters}
+\end{figure}
+
+\newpage
+
+\FloatBarrier
 \subsection{Hyperprior grids}
@@
 
-\todo{Explain Adjusted Rand Index}
+\section{Measuring clustering similarity}
 
-\printbibliography
+The adjusted rand index (ARI) is a measure of the similarity of two different clusterings \autocite{Rand1971} -- two identical clusters have an ARI of 1.0 while progressively more dissimilar clusters have lower ARIs, becoming negative as the clustering gets anti-correlated.
 

### 2013-12-24 experiments/mouseretina/pipeline.py

@@ -1071,24 +1071,28 @@ def plot_truth_latent(exp_results,
     # compute the axes positions
     COL_NUMBER = 4
     COL_SPACE = 1.0 / COL_NUMBER
-    COL_WIDTH = COL_SPACE - 0.02
+    COL_WIDTH = COL_SPACE - 0.03
     COL_PRE = 0.02
     ROW_CONTRACT = 0.05
     ROW_PRE = 0.02
     ROW_SPACE = 0.01
 
+    VERT_SCALE = 1.0
     s = df['cluster'].value_counts()
 
     a = irm.util.multi_napsack(COL_NUMBER, np.array(s))
+    CLUST_MAX_PER_COL = len(a[0])
+    VERT_SCALE = 1.0 -  (CLUST_MAX_PER_COL+4) *ROW_SPACE
 
     MAX_LEN = np.sum([np.array(s)[ai] for ai in a[0]])
+
     for col_i, col in enumerate(a):
         pos = 0
         for row_pos in col:
             cluster_id = s.index.values[row_pos]
             sub_df = df[df['cluster'] == cluster_id]
             sub_df = sub_df.sort('cell_type')
-            height = len(sub_df) / float(MAX_LEN) * 0.90
+            height = len(sub_df) / float(MAX_LEN) * VERT_SCALE
 
             ax = f.add_axes([COL_PRE + col_i * COL_SPACE, 
                              1.0 - pos - height - ROW_PRE, 
@@ -1136,114 +1140,6 @@ def plot_truth_latent(exp_results,
     f.savefig(out_filename)
 
 
-    # for fi, (circos_filename_main, circos_filename_small) in enumerate(out_filenames):
-    #     CLASS_N = len(np.unique(cell_assignment))
-        
-
-    #     class_ids = sorted(np.unique(cell_assignment))
-
-    #     custom_color_map = {}
-    #     for c_i, c_v in enumerate(class_ids):
-    #         c = np.array(pylab.cm.Set1(float(c_i) / CLASS_N)[:3])*255
-    #         custom_color_map['ccolor%d' % c_v]  = c.astype(int)
-
-    #     colors = np.linspace(0, 360, CLASS_N)
-    #     color_str = ['ccolor%d' % int(d) for d in class_ids]
-        
-    #     for n, v in coarse_colors.iteritems():
-    #         custom_color_map['true_coarse_%s' % n] = v
-
-    #     circos_p = irm.plots.circos.CircosPlot(cell_assignment, 
-    #                                            ideogram_radius="0.7r",
[…]

### 2013-12-24 experiments/regirmfail/pipeline2.py

@@ -563,14 +563,16 @@ def merge_results(exp_results, merge_filename):
                       'chain_id' : di, 
                       'assign' : a, 
                       'score' : d['scores'][-1], 
-                      'true_assign' : true_assignvect}
+                      'true_assign' : true_assignvect, 
+                      'node_pos' : nodes_with_class['pos']}
             result.update(params)
             results.append(result)
     df = pandas.DataFrame(results)
     pickle.dump(df, open(merge_filename, 'w'))
 
 PLOT_DATASETS = ['class_compare', 'class_compare_big']
-@files(merge_results, [('%s.rand.pdf' % c, '%s.ari.pdf' % c) for c in PLOT_DATASETS])
+@files(merge_results, [('%s.rand.pdf' % c, '%s.ari.pdf' % c, 
+                        '%s.spatial_variance.pdf' % c) for c in PLOT_DATASETS])
 def plot_results(infile, outfiles):
     df = pickle.load(open(infile, 'r'))
     
@@ -627,6 +629,79 @@ def plot_results(infile, outfiles):
         ax.set_xticklabels([1, 2, 4, 8, 16])
         f.tight_layout()
         f.savefig(plot_files[1])
+
+
+        ## The future 
+
+        def cluster_var(row):
+            assign = row['assign']
+            true_assign = row['true_assign']
+            node_pos = row['node_pos']
+            def node_to_df(assign, nodes):
+                return pandas.DataFrame({'cluster' : assign, 'x' : node_pos[:, 0], 
+                                    'y' : node_pos[:, 1], 'z' : node_pos[:, 2]})
+
+            rdf1 = node_to_df(assign, node_pos)
+            rdf1['truth'] = False
+
+            rdf2 = node_to_df(true_assign, node_pos)
+            rdf2['truth'] = True
+
+            rdf = pandas.concat([rdf1, rdf2])
+            rdf[['x', 'y', 'z']] = rdf[['x', 'y', 'z']].astype(float)
+
+
+            return rdf.groupby(['truth', 'cluster']).var()
+
+        #a = df_cc.groupby(['dataset_name', 'jitter', 'model', 'nonzero_frac', 'class_n', 
+        #                    'side_n', 'seed', 'truth']).apply(lambda group: group.sort_index(by='score', ascending=False).head(1))
+        df_vars = []
+        for rid, r in a.iterrows():
+            cv = cluster_var(r)
+            cv['model'] = r['model']
+            cv['seed'] = r['seed']
[…]

### 2013-12-24 irm/gibbs.py
added fixed gibbs

@@ -81,3 +81,34 @@ def gibbs_sample_type_nonconj(domain_inf, M, rng, impotent=False):
 
         # for r in domain_inf.relations:
         #     r.assert_assigned()
+
+def gibbs_sample_fixed_k(domain_inf, rng, impotent=False):
+
+    T_N = domain_inf.entity_count()
+
+    if impotent:
+        print "gibbs_sample_type: IMPOTENT"
+
+    for entity_pos in np.random.permutation(T_N):
+        g = domain_inf.remove_entity_from_group(entity_pos)
+        # if domain_inf.group_size(g) == 0:
+        #     temp_group = g
+        # else:
+        #     temp_group = domain_inf.create_group(rng)
+
+
+        groups = domain_inf.get_groups()
+        scores = np.zeros(len(groups))
+        for gi, group_id in enumerate(groups):
+            scores[gi] = domain_inf.post_pred(group_id, entity_pos)
+        #print entity_pos, scores
+        sample_i = util.sample_from_scores(scores)
+        new_group = groups[sample_i]
+
+        if impotent:
+            new_group = g
+
+        domain_inf.add_entity_to_group(new_group, entity_pos)
+        # if new_group != temp_group:
+        #     assert domain_inf.group_size(temp_group) == 0
+        #     domain_inf.delete_group(temp_group)

### 2013-12-24 irm/irmio.py

@@ -13,7 +13,7 @@ def model_from_config_file(configfile):
 
 
 def create_model_from_data(data, relation_class=pyirmutil.Relation, 
-                      rng=None):
+                           rng=None, fixed_k=False):
 
     domains_config = data['domains']
     relations_config = data['relations']
@@ -42,7 +42,8 @@ def create_model_from_data(data, relation_class=pyirmutil.Relation,
     domain_interfaces = {}
     for d_name, d_config in domains_config.iteritems():
         D_N = d_config['N'] 
-        ti = model.DomainInterface(D_N, domains_to_relations[d_name])
+        ti = model.DomainInterface(D_N, domains_to_relations[d_name], 
+                                   fixed_k)
         domain_interfaces[d_name] = ti
 
         

### 2013-12-24 irm/model.py

@@ -25,7 +25,7 @@ class DomainInterface(object):
     Also computes the CRP
     """
 
-    def __init__(self, ENT_N, relations_dict):
+    def __init__(self, ENT_N, relations_dict, fixed_k = False):
         """
         relations : {'relationname' : ('DOMAINNAME', rel_obj)}
         where domainname is the name the relation knows this domain as
@@ -45,6 +45,9 @@ class DomainInterface(object):
         self.assignments[:] = NOT_ASSIGNED
         self.temp = 1.0
 
+        self.fixed_k = fixed_k
+        print "This model fixed_k= ", self.fixed_k
+
     def get_relation_pos(self, relation_name):
         return self.relation_pos[relation_name]
 
@@ -60,6 +63,9 @@ class DomainInterface(object):
     def get_groups(self):
         return self.gid_mapping.keys()
 
+    def group_count(self):
+        return len(self.gid_mapping)
+
     def set_temp(self, t):
         self.temp = t
 
@@ -72,6 +78,10 @@ class DomainInterface(object):
         new_gid = self.g_pos
         self.g_pos += 1
         self.gid_mapping[new_gid] = tuple(rel_groupid)
+
+        #if len(self.gid_mapping) > self.fixed_k:
+        #    print "WARNING: Creating too many groups in fixed K configuration" 
+
         return new_gid
     
     def group_size(self, gid):
@@ -100,7 +110,10 @@ class DomainInterface(object):
         [r.delete_group(t, g) for g, (t, r) in zip(rel_groupid, self.relations)]
         
         del self.gid_mapping[group_id]
-        
+
+        #if self.fixed_k:
+        #    print "WARNING: Deleting group in Fixed K configuration"
+
     def add_entity_to_group(self, group_id, entity_pos):
         assert self.assignments[entity_pos] == NOT_ASSIGNED
         rel_groupid = self.gid_mapping[group_id]
@@ -121,8 +134,11 @@ class DomainInterface(object):
 
     def get_prior_score(self):
         count_vect = util.assign_to_counts(self.assignments)
-        crp_score = util.crp_score(count_vect, self.alpha)
-        return crp_score
+        if not self.fixed_k :
+            score = util.crp_score(count_vect, self.alpha)
+        else:
+            score = util.dm_score(count_vect, self.alpha, self.group_count())
+        return score
         
     def post_pred(self, group_id, entity_pos):
         """
@@ -136,9 +152,13 @@ class DomainInterface(object):
[…]

### 2013-12-24 irm/runner.py

@@ -19,6 +19,9 @@ def default_kernel_nonconj_config():
     return [('nonconj_gibbs', {'M' : 10}), 
             ('slice_params', {'width' : 0.0})] # use default
 
+def default_kernel_fixed_config():
+    return [('fixed_gibbs', {}), 
+            ('slice_params', {'width' : 0.0})] # use default
     
 
 def add_domain_hp_grid_kernel(kernel_list, grid=None):
@@ -71,7 +74,12 @@ def do_inference(irm_model, rng, kernel_config, iteration,
         t1 = time.time()
         if kernel_name == 'conj_gibbs':
             for domain_name, domain_inf in irm_model.domains.iteritems():
-                gibbs.gibbs_sample_type(domain_inf, rng, params.get("impotent", False))
+                gibbs.gibbs_sample_type(domain_inf, rng, 
+                                        params.get("impotent", False))
+        elif kernel_name == 'fixed_gibbs':
+            for domain_name, domain_inf in irm_model.domains.iteritems():
+                gibbs.gibbs_sample_fixed_k(domain_inf, rng, 
+                                           params.get("impotent", False))
         elif kernel_name == 'nonconj_gibbs':
             for domain_name, domain_inf in irm_model.domains.iteritems():
                 gibbs.gibbs_sample_type_nonconj(domain_inf, 
@@ -131,7 +139,8 @@ def do_inference(irm_model, rng, kernel_config, iteration,
     return res
 
 class Runner(object):
-    def __init__(self, latent, data, kernel_config, seed=None):
+    def __init__(self, latent, data, kernel_config, seed=None, 
+                 fixed_k = False):
 
         # FIXME add seed
 
@@ -140,7 +149,8 @@ class Runner(object):
         if seed != None:
             pyirm.set_seed(self.rng, seed)
         
-        self.model = irmio.create_model_from_data(data, rng=self.rng)
+        self.model = irmio.create_model_from_data(data, rng=self.rng, 
+                                                  fixed_k = fixed_k)
         irmio.set_model_latent(self.model, latent, self.rng)
         self.iters = 0
         

### 2013-12-24 irm/util.py

@@ -301,3 +301,25 @@ def multi_napsack(K, vals):
     return out
 
 
+def dm_post_pred(mk, N, alpha, fixed_k):
+    """
+    dirichlet multinomial post pred
+    Taken from Rasmussen NIPS 99 eqn 13
+
+    Prob of adding a customer to the table which currently seats
+    mk people.
+    """
+    return np.log((mk + alpha/fixed_k) / (N - 1 + alpha))
+
+
+def dm_score(counts, alpha, fixed_k):
+    """
+    Dirichlet multinomial score of a count vector
+    """
+    score = 0
+    i = 0
+    for table in counts:
+        for customer in range(table):
+            score += dm_post_pred(customer, i + 1, alpha, fixed_k)
+            i += 1
+    return score

### 2013-12-24 tests/test_fixed_gibbs.py

@@ -1,0 +1,102 @@
+from nose.tools import *
+import irm
+import numpy as np
+from matplotlib import pylab
+from irm import runner, irmio
+import copy
+
+
+
+def test_fixed_k():
+    # create synthetic data with K groups
+    seed = 0
+
+    np.random.seed(seed)
+
+    GROUP_N = 30
+    ENTITIES_PER_GROUP = 10
+    N = GROUP_N * ENTITIES_PER_GROUP
+
+    model_name = "BetaBernoulli"
+    a = np.random.permutation(np.arange(N) % GROUP_N)
+    latent = {'domains' : 
+              {'d1' : 
+            {'assignment' : a, 
+         }}, 
+              'relations' : {'R1' : {'hps' : {'alpha' : 0.5, 
+                                              'beta' : 0.5}}}}
+
+    # ss = {}
+    # for g1 in range(GROUP_N):
+    #     for g2 in range(GROUP_N):
+    #         ss[(g1, g2)] = {'p' : np.random.beta(0.5, 0.5)}
+            
+    # latent['domains']['d1']['ss'] = ss
+
+    data = {'domains' : {'d1' : {'N' : N}}, 
+            'relations' : {'R1' : {'relation' : ('d1', 'd1'), 
+                                   'model' : model_name}}}
+    
+    new_latent, new_data = irm.data.synth.prior_generate(latent, data)
+    # print new_data
+    # m = new_data['relations']['R1']['data']
+    # f = pylab.figure()
+    # ax = f.add_subplot(1, 2, 1)
+    # ax.imshow(m, interpolation='nearest')
+    # ax2 = f.add_subplot(1, 2, 2)
+    # ai = np.argsort(a).flatten()
+    # m2 = m[ai, :]
+    # m2 = m2[:, ai]
+    # ax2.imshow(m2, interpolation='nearest')
+    # pylab.show()
+
+    # create model and initialize with that K
+    
+    # score
+
+    # do inference
+
+
+    # does score, assignment vector get better? 
+
+    kernel_config = irm.runner.default_kernel_fixed_config()
+
+    run_truth = runner.Runner(new_latent, new_data, kernel_config, seed=0, 
+                              fixed_k=True)
+
+    irmio.estimate_suffstats(run_truth.model, run_truth.rng)
+
+    # get ground truth
+    ground_truth_score = run_truth.get_score()
+
+    cleaned_up_latent = run_truth.get_state()
+
+    rand_init = copy.deepcopy(cleaned_up_latent)
[…]

### 2013-12-24 experiments/mouseretina/pipeline.py
teeing up for fixed-k experiment

@@ -42,6 +42,7 @@ EXPERIMENTS = [
     #('retina.1.0.ld.truth', 'truth_100', 'anneal_slow_400'), 
     ('retina.1.0.ld.0.0', 'fixed_10_20', 'debug'), 
     ('retina.count.edp', 'fixed_10_20', 'debug'), 
+    ('retina.1.0.bb.0.0', 'fixed_10_10', 'debug_fixed'), 
                
 ]
 
@@ -61,6 +62,9 @@ INIT_CONFIGS = {'fixed_10_200' : {'N' : 10,
                 'fixed_10_20' : {'N' : 10, 
                                   'config' : {'type' : 'fixed', 
                                               'group_num' : 20}}, 
+                'fixed_10_10' : {'N' : 10, 
+                                  'config' : {'type' : 'fixed', 
+                                              'group_num' : 10}}, 
                 'fixed_20_100' : {'N' : 20, 
                                   'config' : {'type' : 'fixed', 
                                               'group_num' : 100}}, 
@@ -96,6 +100,8 @@ KERNEL_CONFIGS = {
 
                   'debug' : {'ITERS' : 10, 
                              'kernels' : irm.runner.default_kernel_nonconj_config()}, 
+                  'debug_fixed' : {'ITERS' : 40, 
+                             'kernels' : irm.runner.default_kernel_fixed_config()}, 
 
                   }
 
@@ -451,6 +457,7 @@ def run_exp((data_filename, inits), wait_file, kernel_config_name):
     CHAINS_TO_RUN = len(inits)
     ITERS = kc['ITERS']
     kernel_config = kc['kernels']
+    fixed_k = kc.get('fixed_k', False)
     
     jids = cloud.map(irm.experiments.inference_run, inits, 
                      [data_filename]*CHAINS_TO_RUN, 
@@ -458,6 +465,8 @@ def run_exp((data_filename, inits), wait_file, kernel_config_name):
                      [ITERS] * CHAINS_TO_RUN, 
                      range(CHAINS_TO_RUN), 
                      [BUCKET_BASE]*CHAINS_TO_RUN, 
+                     [None] * CHAINS_TO_RUN, 
+                     [fixed_k] * CHAINS_TO_RUN, 
                      _label="%s-%s-%s" % (data_filename, inits[0], 
                                           kernel_config_name), 
                      _env='connectivitymotif', 

### 2013-12-24 irm/experiments.py

@@ -199,6 +199,7 @@ def from_bucket(filename, BUCKET_BASE):
 def inference_run(latent_filename, 
                   data_filename, 
                   kernel_config,  ITERS, seed, BUCKET_BASE, init_type=None, 
+                  fixed_k = False, 
                   latent_samp_freq=20):
     """
     For running on the cloud
@@ -207,7 +208,8 @@ def inference_run(latent_filename,
     latent = from_bucket(latent_filename, BUCKET_BASE)
     data = from_bucket(data_filename, BUCKET_BASE)
 
-    chain_runner = irm.runner.Runner(latent, data, kernel_config, seed)
+    chain_runner = irm.runner.Runner(latent, data, kernel_config, seed, 
+                                     fixed_k = fixed_k)
 
     if init_type != None:
         chain_runner.init(init_type)

### 2013-12-24 experiments/mouseretina/pipeline.py
fixed k update

@@ -101,6 +101,7 @@ KERNEL_CONFIGS = {
                   'debug' : {'ITERS' : 10, 
                              'kernels' : irm.runner.default_kernel_nonconj_config()}, 
                   'debug_fixed' : {'ITERS' : 40, 
+                                   'fixed_k' : True, 
                              'kernels' : irm.runner.default_kernel_fixed_config()}, 
 
                   }

### 2013-12-24 irm/runner.py

@@ -143,7 +143,7 @@ class Runner(object):
                  fixed_k = False):
 
         # FIXME add seed
-
+        print "FIXED_K=", fixed_k
         # create the model
         self.rng = pyirm.RNG()
         if seed != None:

### 2013-12-26 docs/paper.tex
added fixed gibbs
updates to mouse pipeline to plot

@@
 We compared the results of our model to the traditional ISBM to see if
-the addition of spatial information was necessary. We find that, as expected from our synthetic experiments, the
-IRM massively over-segmented the data, finding
-spatially-localized “communities” of cells which bore little relation
-to known types.  The average spatial extent of the types found by the ISBM
-was  $14.6\pm8.1 \mu m$. However, in the ground truth by
-anatomists the average distance was $34.3\pm7.6 \mu m$ and $35.3\pm 2.4 \mu m$, quite
-close to the $34.5\pm 1.0 \mu m$ discovered by our algorithm. Modeling the
-distance-dependence prevents over-segmentation and thus allows more
-meaningful estimates.
+the addition of spatial information was necessary. We find that, as
+expected from our synthetic experiments, the IRM massively
+over-segmented the data, finding spatially-localized “communities” of
+cells which bore little relation to known types.  The average spatial
+extent of the types found by the ISBM was $14.6\pm8.1 \mu m$. However,
+in the ground truth by anatomists the average distance was $34.3\pm7.6
+\mu m$ and $35.3\pm 2.4 \mu m$, quite close to the $34.5\pm 1.0 \mu m$
+discovered by our algorithm. Modeling the distance-dependence prevents
+over-segmentation and thus allows more meaningful estimates.
 
@@
 \label{supp:otherlikelihoods}
- - Reparameterized logistic-distance bernoulli with fixed lambda
 
- - Logistic-distance poisson
+We reparameterized the Logistic-Distance Bernoulli likelihood to
+better capture the microprocessor data structure. We explicitly
+setting the maximum probability $p$ of the logistic function on a
+per-component basis, drawing from a global $p \sim Beta(\alpha_{hp},
+\beta_{hp})$. Then $\lambda$ is set for each component as a global
+hyperparameter, $\lambda$.
+
+The ``logistic-distance Poisson'' spatial model assumes that, if cell
+$e_i$ is of class $m$ and cell $e_j$ is of class $n$, then $\eta_{mn}
+= (\mu_{mn}, r_{mn})$, and the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
+\begin{eqnarray}
+\lambda^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
+\lambda & = & \lambda^* \cdot (r_{mn} - r_{min}) + r_{min} \\
+R(e_i, e_j) \sim Poi(c | \lambda)
+\end{eqnarray}
+
+where $\lambda$ and $r_{min}$ are per-graph parameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $r_{mn} \sim \exp(r_{mn} | r_{scale}^{hp})$.
 
@@
 \FloatBarrier
-\subsection{Comparison to IRM for Mouse Retina}
+\subsection{Model Comparison for Mouse Retina}
+
+We compare the clusters found by our model, the traditional infinite
+stochastic block model, and a finite stochastic block model with the number
+of clusters set to the best-found number from our iSRM model ($k=12$).
+
+We see from fig~\ref{fig:supp:irmspatialcomparemouse} that the spatial
+variance of our clusters closely matches the ground truth, while the
+other two models find spatially-localized
+clusters. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
+clustering, the clustering found by the iSBM,
+as well as the result of parametrically varying the traditional stochastic
+block model for various values of $k$.
+
+The actual clusterings for iSRM, the iSBM, and $SBM_{k=12}$ are shown
+in figures \ref{fig:supp:retinabbclusters},
+\ref{fig:supp:retinaldclusters}, and
+\ref{fig:supp:retinabbfixedclusters} respectively. Each block is a
+cluster, with the true cell type indicated along the x-axis, and the
+cell ID written to the left or right.
 
@@
   \centering 
+    \centerline{\includegraphics[width=5in]{mouseretina/ari_vs_cluster.pdf}}
+  \caption{}
+\label{fig:supp:mousecompareari}
+\end{figure}
+
+\begin{figure}[h!]
+  \centering 
     \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
@@
 \label{fig:supp:retinabbclusters}
+\end{figure}
+
+\begin{figure}[h!]
+
+  \centering 
+    \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
+  \caption{}
+\label{fig:supp:retinabbfixedclusters}
 \end{figure}
@@
 
-The adjusted rand index (ARI) is a measure of the similarity of two different clusterings \autocite{Rand1971} -- two identical clusters have an ARI of 1.0 while progressively more dissimilar clusters have lower ARIs, becoming negative as the clustering gets anti-correlated.
+The adjusted rand index (ARI) is a measure of the similarity of two
+different clusterings \autocite{Rand1971} -- two identical clusters
+have an ARI of 1.0 while progressively more dissimilar clusters have
+lower ARIs, becoming negative as the clustering gets anti-correlated.
 

### 2013-12-26 experiments/mouseretina/pipeline.py

@@ -8,6 +8,7 @@ from matplotlib import pylab
 import matplotlib
 import pandas
 import colorbrewer
+from nose.tools import assert_equal 
 
 import matplotlib.gridspec as gridspec
 
@@ -38,13 +39,19 @@ EXPERIMENTS = [
     ('retina.1.1.bb.0.0', 'fixed_20_100', 'anneal_slow_400'), 
     ('retina.1.2.bb.0.0', 'fixed_20_100', 'anneal_slow_400'), 
     ('retina.1.3.bb.0.0', 'fixed_20_100', 'anneal_slow_400'), 
-    ('retina.count.edp', 'fixed_20_100', 'anneal_slow_400'), 
+    # ('retina.count.edp', 'fixed_20_100', 'anneal_slow_400'), 
     #('retina.1.0.ld.truth', 'truth_100', 'anneal_slow_400'), 
-    ('retina.1.0.ld.0.0', 'fixed_10_20', 'debug'), 
-    ('retina.count.edp', 'fixed_10_20', 'debug'), 
+    #('retina.1.0.ld.0.0', 'fixed_10_20', 'debug'), 
+    #('retina.count.edp', 'fixed_10_20', 'debug'), 
     ('retina.1.0.bb.0.0', 'fixed_10_10', 'debug_fixed'), 
                
 ]
+MANUAL_K = [5, 10, 12, 15, 20, 30, 50, 70]
+for k in MANUAL_K:
+    EXPERIMENTS += [('retina.1.1.bb.0.0', 'fixed_20_%d' % k, 'anneal_slow_fixed_400'), 
+                    ('retina.1.2.bb.0.0', 'fixed_20_%d' % k, 'anneal_slow_fixed_400')]
+
+
 
 THOLDS = [0.01, 0.1, 0.5, 1.0]
     
@@ -74,6 +81,10 @@ INIT_CONFIGS = {'fixed_10_200' : {'N' : 10,
                               'config' : {'type' : 'truth'}}, 
 }
 
+for k in MANUAL_K:
+    INIT_CONFIGS['fixed_20_%d' % k] = {'N' : 20, 
+                                       'config' : {'type' : 'fixed', 
+                                                   'group_num' : k}}
 
 slow_anneal = irm.runner.default_kernel_anneal()
 slow_anneal[0][1]['anneal_sched']['start_temp'] = 64.0
[…]

### 2013-12-26 irm/experiments.py

@@ -20,7 +20,7 @@ It's only here because it's easy to import.
 
 
 def create_init(latent_filename, data_filename, out_filenames, 
-                init= None):
+                init= None, keep_ground_truth=True):
     """ 
     CONVENTION: when we create N inits, the first is actually 
     initialized from the "ground truth" of the intial init (whatever
@@ -58,7 +58,7 @@ def create_init(latent_filename, data_filename, out_filenames,
         else:
             raise NotImplementedError("Unknown init type")
             
-        if c > 0: # first one stays the same
+        if (not keep_ground_truth) or (c > 0) : # first one stays the same
             latent['domains']['d1']['assignment'] = a
 
         # generate new suffstats, recompute suffstats in light of new assignment

### 2013-12-26 docs/paper.tex
updates to citations

@@
 structure, ushering in an era of bioinformatics as an independent
-discipline \autocite{}.
+discipline \autocite{Koboldt2013}.
 
@@
 has been prominent in many areas, ranging from sensory (e.g. retina,
-\autocite{} via processing (e.g. neocortex) to movement (e.g. spinal
-cord) \autocite{}. Indeed, such repeated patterns also exists in
-human-made computing circuits. It remains an important challenge to
-develop algorithms to use anatomical data, e.g. connectomics, to back
-out underlying connectivity motifs.
+\autocite{Masland2001} via processing (e.g. neocortex \autocite{Mountcastle1997})
+to movement (e.g. spinal cord) \autocite{Grillner2005}. Indeed, such
+repeated patterns also exists in human-made computing circuits. It
+remains an important challenge to develop algorithms to use anatomical
+data, e.g. connectomics, to back out underlying connectivity motifs.
 
@@
 The discovery of structure in probabilistic graphs is a crucial aspect
-in machine learning. Commonly used algorithms include spectral methods
-\autocite{},   community-based-detection   methods  \autocite{},   and
+in machine learning. Commonly used algorithms include
+community-based-detection methods \autocite{Girvan2002}, and
 stochastic block models \autocite{Nowicki2001}.  While these
-approaches powerfully  incorporate the probabilistic nature of neural
-connections \autocite{} they do not generally consider the known fact
-that  the  probability  of  connections  between  neurons  depends  on
-distance. Combining such probabilistic models with a spatial component
-to model distance promises to  improve the detection of microcircuitry
-from neural data.
+approaches can incorporate the probabilistic nature of neural
+connections \autocite{Hill2012} they do not generally consider the
+known fact that the probability of connections between neurons depends
+on distance. Combining such probabilistic models with a spatial
+component to model distance promises to improve the detection of
+microcircuitry from neural data.
 
@@
 types and their spatial patterns of interconnection. We apply it to
-three computing systems: the mouse retina connectome \autocite{}, the
-c. elegans connectome \autocite{}, and a ``connectome'' of a classical
+three computing systems: the mouse retina connectome \autocite{Helmstaedter2013}, the
+c. elegans connectome \autocite{White1986}, and a ``connectome'' of a classical
 microprocessor \autocite{James2010}. Note we do not explicitly
@@
 There have existed a range of previous approaches to the discovery
-of neural microcircuitry \autocite{} These generally involve a great deal
+of neural microcircuitry \autocite{missing} These generally involve a great deal
 of manual labor and ad-hoc determination of what constitutes a “type”
@@
 across animals.
-For example, the types of neurons in one animal can be associated with the types of neurons in another animal, in the same way as this is already possible through molecular markers \autocite{}.
-This could be particularly important if cell types appear that are due to properties of the stimuli and experience as opposed to the molecular properties of cells \autocite{}.
+For example, the types of neurons in one animal can be associated with the types of neurons in another animal, in the same way as this is already possible through molecular markers \autocite{Brown2009}.
+This could be particularly important if cell types appear that are due to properties of the stimuli and experience as opposed to the molecular properties of cells \autocite{missing}.
 This would allow comparative quantitative anatomy across animals, and the search for the causes of the ultimate connectivity.
@@
 ongoing area of research in the probabilistic modeling community
-\autocite{}.
+\autocite{missing}.
 
@@
 The adjusted rand index (ARI) is a measure of the similarity of two
-different clusterings \autocite{Rand1971} -- two identical clusters
+different clusterings \autocite{Hubert1985} -- two identical clusters
 have an ARI of 1.0 while progressively more dissimilar clusters have

### 2013-12-26 experiments/mouseretina/pipeline.py

@@ -9,6 +9,7 @@ import matplotlib
 import pandas
 import colorbrewer
 from nose.tools import assert_equal 
+from sklearn import metrics
 
 import matplotlib.gridspec as gridspec
 
@@ -1246,8 +1247,11 @@ def compute_cluster_metrics(exp_results,
     ca = irm.util.canonicalize_assignment(df['cluster'])
 
 
-    ari = rand.compute_adj_rand_index(canon_true_fine, ca)
-    ari_coarse = rand.compute_adj_rand_index(canon_true_coarse, ca)
+    ari = metrics.adjusted_rand_score(canon_true_fine, ca)
+    ari_coarse = metrics.adjusted_rand_score(canon_true_coarse, ca)
+
+    ami = metrics.adjusted_mutual_info_score(canon_true_fine, ca)
+    ami_coarse = metrics.adjusted_mutual_info_score(canon_true_coarse, ca)
 
                                              
     jaccard = rand.compute_jaccard(canon_true_fine, ca)
@@ -1266,6 +1270,8 @@ def compute_cluster_metrics(exp_results,
     
     pickle.dump({'ari' : ari, 
                  'ari_coarse' : ari_coarse, 
+                 'ami' : ami, 
+                 'ami_coarse' : ami_coarse, 
                  'jaccard' : jaccard, 
                  'jaccard_coarse' : jaccard_coarse,
                  'n11' : ss['n11'], 
@@ -1285,6 +1291,8 @@ def merge_cluster_metrics(infiles, outfile):
         res.append({'filename' : infile, 
                     'ari' : d['ari'], 
                     'ari_coarse' : d['ari_coarse'], 
+                    'ami' : d['ami'], 
+                    'ami_coarse' : d['ami_coarse'], 
                     'jaccard_coarse' : d['jaccard_coarse'], 
                     'jaccard' : d['jaccard'], 
                     'cluster_n' : d['cluster_n'],
@@ -1398,11 +1406,12 @@ def plot_cluster_aris(infile, (outfile_plot, outfile_rpt)):
     f = pylab.figure()
     ax = f.add_subplot(1, 1, 1)
 
-    ax.scatter(finite_df['cluster_n'], finite_df['ari'], s=50, c='b', label="SBM")
+    index = 'ari'
+    ax.scatter(finite_df['cluster_n'], finite_df[index], s=50, c='b', label="SBM")
 
 
[…]

### 2013-12-26 experiments/regirmfail/pipeline2.py

@@ -7,6 +7,7 @@ import time
 from matplotlib import pylab
 import pandas
 import rand
+from sklearn import metrics
 
 import irm
 import irm.data

### 2013-12-27 docs/paper.tex
about to add in all the erica comments

@@
 \usepackage[backend=biber,style=nature]{biblatex}
-\usepackage{todonotes}
+\usepackage[disable]{todonotes}
 \usepackage{subfiles}
@@
   necessitating new analysis methods to discover underlying
-  structure. It has long been assumed that discovering the local patterns of microcircuitry is the key to understanding neural function. Here we developed a
-  nonparametric Bayesian technique that identifies neuron types and circuitry
-  patterns in connectomics data. We show that the approach recovers
-  known neuron types in the retina, reveals interesting structure in
-  the nervous system of c. elegans, and automatically discovers the
-  structure of microprocessors. Our approach extracts structural meaning from connectomics, enabling new approaches of deriving anatomical insights from these emerging datasets.
+  structure. It has long been assumed that discovering the local
+  patterns of microcircuitry is the key to understanding neural
+  function. Here we developed a nonparametric Bayesian technique that
+  identifies neuron types and circuitry patterns in connectomics
+  data. We show that the approach recovers known neuron types in the
+  retina, reveals interesting structure in the nervous system of
+  c. elegans, and automatically discovers the structure of
+  microprocessors. Our approach extracts structural meaning from
+  connectomics, enabling new approaches of deriving anatomical
+  insights from these emerging datasets.
 
@@
 types and their spatial patterns of interconnection. We apply it to
-three computing systems: the mouse retina connectome \autocite{Helmstaedter2013}, the
-c. elegans connectome \autocite{White1986}, and a ``connectome'' of a classical
+three computing systems: the mouse retina connectome
+\autocite{Helmstaedter2013}, the c. elegans connectome
+\autocite{White1986}, and a ``connectome'' of a classical
 microprocessor \autocite{James2010}. Note we do not explicitly
@@
 We use a Bayesian structure discovery approach that is based on
-assumptions about the way positions and cell types affect the
+prior knowledge about the way positions and cell types affect the
 probability of connections between cells, a so-called generative
-model. The probability of connections between pairs of neurons
+model. We term our model the ``infinite spatial-relational model''. 
+The probability of connections between pairs of neurons
 decreases with distance, but how high the baseline probability of
 connection is and how quickly it decays with distance depends on the
-two cell-types. We combine annealing with Markov Chain Monte Carlo to
+two cell-types. Note that this simplistic prior information is all that 
+we tell our model; no additional physiological information is included.
+
+We combine annealing with Markov Chain Monte Carlo to
 perform simultaneously joint posterior inference over the structure,
@@
 and well established model (the infinite stochastic block model, or
-ISBM) that assumes that there is no distance dependence to the
+iSBM) that assumes that there is no distance dependence to the
 connectivity. The model converges relatively quickly to an estimate of
@@
 different true class counts, between our model and a traditional
-ISBM. c.) For synthetic data, 
-the IRM over clusters, finding a large number of spatially-localized types. 
-In contrast, the spatial-relational model recovers the true spatial
+iSBM. c.) For synthetic data, 
+the iSBM overclusters, finding a large number of spatially-localized types. 
+In contrast, the infinite spatial-relational model recovers the true spatial
 extent of the underlying types.}
@@
 \subsubsection{Mouse Retina}
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics{mouseretina.ai}}
+  \caption{Discovering cell classes in the mouse retina connectome. 
+a.) Input adjacency data for 950 cells for which soma positions were known. b. clustered adjacency matrix. c.) connectivity diagram showing our clusters, as
+well as the cell depth and known cell type of each cell. Our model agrees
+with the cell types traditionally identified by anatomists, and identifies 
+laminar-specific connectivity patterns. d.) Connectivity between our
+clusters as a function of distance -- the cluster consisting primarily of
+retinal ganglion cells (lower left-center) exhibits the expected near and
+far connectivity. e. The spatial distribution of our cell types -- each cell type tesselates space.}
+\label{fig:mouseretina}
+\end{figure}
+
 The retina of the mouse \autocite{Masland2001} is an example where we
@@
 between different cells, and amacrine cells that feed into the
-ganglion cells (cite some review). Recent research \autocite{Helmstaedter2013} has
-produced a large dataset containing both the types of cells from
-orthogonal approaches, and also the connectivity matrix between all
-reconstructed cells. We want to know to which level we can classify
-neurons into types based on this connectivity only.
+ganglion cells \autocite{cite some review}. Recent research
+\autocite{Helmstaedter2013} has produced a large dataset containing
+both the types of cells from orthogonal approaches, and also the
+connectivity matrix between all reconstructed cells. We want to know
+to which level we can classify neurons into types based on this
+connectivity only.
 
@@
 
-We compared the results of our model to the traditional ISBM to see if
+We compared the results of our model to the traditional iSBM to see if
 the addition of spatial information was necessary. We find that, as
-expected from our synthetic experiments, the IRM massively
+expected from our synthetic experiments, the iSBM massively
 over-segmented the data, finding spatially-localized “communities” of
 cells which bore little relation to known types.  The average spatial
-extent of the types found by the ISBM was $14.6\pm8.1 \mu m$. However,
+extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. However,
 in the ground truth by anatomists the average distance was $34.3\pm7.6
@@
 
-\begin{figure}
+\subsection{C. elegans}
+
+\begin{figure}[ht!]
   \centering 
-  \centerline{\includegraphics{mouseretina.ai}}
-  \caption{Discovering cell classes in the mouse retina connectome. 
-a.) Input adjacency data for 950 cells for which soma positions were known. b. clustered adjacency matrix. c.) connectivity diagram showing our clusters, as
-well as the cell depth and known cell type of each cell. Our model agrees
-with the cell types traditionally identified by anatomists, and identifies 
-laminar-specific connectivity patterns. d.) Connectivity between our
-clusters as a function of distance -- the cluster consisting primarily of
-retinal ganglion cells (lower left-center) exhibits the expected near and
-far connectivity.}
-\label{fig:mouseretina}
+  \centerline{\includegraphics{celegans.ai}}
+  \caption{Discovering cell types in the \textit{c. elegans}
+    connectome. a.) Initial cell connectivity matrix, red are
+    (directional) chemical synapses, blue are electrical
+    synapses. Point size is proportional to synapse count between
+    those cells. b.) Clustered adjacency matrix showing identified
+    clusters. c.) Inter-cluster connectivity, showing the position of
+    each cell along the body axis as well as the ``true'' cell type --
+    both coarse (motor, sensory, and interneuron) as well as fine
+    (small text labels). }
+  \label{fig:celegans}
 \end{figure}
-
-\subsection{C. elegans}
 
@@
 determined the underlying clusters explained by connectivity and
-distance (fig~ref{fig:celegans}a). A superficial inspection of the results shows
+distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
 clustering into groups consisting roughly homogeneously of motor
@@
 
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics{celegans.ai}}
-  \caption{Discovering cell types in the \textit{c. elegans}
-    connectome. a.) Initial cell connectivity matrix, red are
-    (directional) chemical synapses, blue are electrical
-    synapses. Point size is proportional to synapse count between
-    those cells. b.) Clustered adjacency matrix showing identified
-    clusters. c.) Inter-cluster connectivity, showing the position of
-    each cell along the body axis as well as the ``true'' cell type --
-    both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). }
-  \label{fig:celegans}
-\end{figure}
-
 \subsection{Microprocessor}
-
-To show the applicability of our method to other connectome-style
-datasets, we obtained the spatial location and interconnectivity of
-the transistors in a classic microprocessor, the MOS Technology 6502
-(used in the Apple II) \autocite{James2010}. Computer architects use
-common patterns of transistors when designing circuits, with each
-transistor having a ``type'' in the circuit. We identified a region of
-the processor with complex but known structure containing the primary
-8-bit registers X, Y, and S (figure).
-
-Our algorithm identifies areas of spatial homogeneity that mirror the
-known structure in the underlying architectural circuit, segmenting
-transistor types recognizable to computer architects. Using the
-original schematics, we see that one identified type contains the
-``clocked'' transistors, which retain digital state. Two other types
-contain transistors with pins C1 or C2 connected to ground, mostly
-serving as inverters.  An additional identified type controls the
-behavior of the three registers of interest (X, Y, and S) with respect
-to the SB data bus, either allowing them to latch or drive data from
-the bus. The repeat patterns of spatial connectivity are visible in
-figure~\ref{fig:microprocessor}a, showing the man-made horizontal and
-vertical layout of the same types of transistors.
-
 \begin{figure}
@@
 \end{figure}
+
+To show the applicability of our method to other connectome-style
+datasets, we obtained the spatial location and interconnectivity of
+the transistors in a classic microprocessor, the MOS Technology 6502
+(used in the Apple II) \autocite{James2010}. Computer architects use
+common patterns of transistors when designing circuits, with each
+transistor having a ``type'' in the circuit. We identified a region of
+the processor with complex but known structure containing the primary
+8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).
+
+Our algorithm identifies areas of spatial homogeneity that mirror the
+known structure in the underlying architectural circuit, segmenting
+transistor types recognizable to computer architects. Using the
+original schematics, we see that one identified type contains the
+``clocked'' transistors, which retain digital state. Two other types
+contain transistors with pins C1 or C2 connected to ground, mostly
+serving as inverters.  An additional identified type controls the
+behavior of the three registers of interest (X, Y, and S) with respect
+to the SB data bus, either allowing them to latch or drive data from
+the bus. The repeat patterns of spatial connectivity are visible in
+figure~\ref{fig:mos6502}a, showing the man-made horizontal and
+vertical layout of the same types of transistors.
 
@@
 and connectivity patterns . We employ a collection of Markov-chain
-Monte carlo techniques [reference methods] but while different
+Monte carlo techniques (see Methods) but while different
 initializations converge to similar ultimate values, we can never
@@
 techniques that may offer better approximations to the global optimum
-\autocite{Wanga2012,FritzJonathan} and future work could adapt them to
+\autocite{Wanga2012} and future work could adapt them to
 the problem we addressed here.
@@
 
-There have existed a range of previous approaches to the discovery
-of neural microcircuitry \autocite{missing} These generally involve a great deal
-of manual labor and ad-hoc determination of what constitutes a “type”
-of cell -- to this day there are disagreements in the literature as to
-the “true” types in the mammalian retina. Much as phylogenomics has
-changed our understanding of animal ontologies, modern large scale
-data will allow the efficient unbiased discovery of cell types and
-circuits. The sheer amount of available data demands the introduction
-of algorithmic approaches.
+There have existed a range of previous approaches to the discovery of
+neural microcircuitry \todo{Do we want a citation here?}. These
+generally involve a great deal of manual labor and ad-hoc
+determination of what constitutes a “type” of cell -- to this day
+there are disagreements in the literature as to the “true” types in
+the mammalian retina. Much as phylogenomics has changed our
+understanding of animal ontologies, modern large scale data will allow
+the efficient unbiased discovery of cell types and circuits. The sheer
+amount of available data demands the introduction of algorithmic
+approaches.
 
 The development of automatic identification and quantification of cell
-type may also provide a new computational phenotype for quantifying the effect
-of disease, genetic interventions, and developmentally-experienced neural activity.
-Our method can in principle identify neuron-types across non-connected graphs, e.g.
-across animals.
-For example, the types of neurons in one animal can be associated with the types of neurons in another animal, in the same way as this is already possible through molecular markers \autocite{Brown2009}.
-This could be particularly important if cell types appear that are due to properties of the stimuli and experience as opposed to the molecular properties of cells \autocite{missing}.
-This would allow comparative quantitative anatomy across animals, and the search for the causes of the ultimate connectivity.
+type may also provide a new computational phenotype for quantifying
+the effect of disease, genetic interventions, and
+developmentally-experienced neural activity. Our method can in
+principle identify neuron-types across non-connected graphs,
+e.g. across animals. For example, the types of neurons in one animal
+can be associated with the types of neurons in another animal, in the
+same way as this is already possible through molecular markers
+\autocite{Brown2009}. This could be particularly important if cell
+types appear that are due to properties of the stimuli and experience
+as opposed to just the molecular properties of cells, such as color
+and orientation selective types in primary visual cortex
+\autocite{Sincich2005,Lennie2005}. This would allow comparative
+quantitative anatomy across animals, and the search for the causes of
+the ultimate connectivity.
 
@@
 We take as input a connectivity matrix $R$ defining the connections
-between cell $e_i$ and $e_j$, as well as a distance function $d(e_i, e_j)$
-representing a (physical) distance between adjacent cells. See the
-supplemental material for extension to multiple connectivity
-matrices. We assume there exist an unknown number $k$ of latent
-(unobserved) cell types, $k \in \{1, 2, 3, \dots\}$, and that
-each cell $e_i$ belongs to a single cell type, $c_i = k$. The observed
-connectivity between two cells $R(e_i, e_j)$ then depends only on
-their latent type and their distance through a link function $f(\cdot, d(e_i, e_j))$.
-We assume $f$ is parameterized based on the latent class, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well as a set of global hyper parameters $\theta$, such that the link function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
+between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
+e_j)$ representing a (physical) distance between adjacent cells. See
+the supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number $K$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots\, K}$, and that each
+cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
+of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
+observed connectivity between two cells $R(e_i, e_j)$ then depends
+only on their latent type and their distance through a link function
+$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
+latent class, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+as a set of global hyper parameters $\theta$, such that the link
+function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
 
 We then jointly infer the maximum a posteriori (MAP) estimate of the
-class assignment vector ${c_i}$, the parameter matrix $\eta_{mn}$, and
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
 the global model hyperparameters $\theta$ :
@@
 
-Our model is a extension of the ISBM
+Our model is a extension of the iSBM
 \autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
@@
 
-We use a Dirichlet-process prior on cluster assignments, which allows
-the number of clusters to be determined automatically. In brief, the
-probability of a cell belonging to a cluster is proportional to the
-number of datapoints already in that cluster, $m_k$, such that $p(c_i
-= k) \propto \frac{m_k}{N + \alpha}$ and the probability of the cell
-belonging to a new cluster $k'$ is $p(c_i = k') \propto
-\frac{\alpha}{N + \alpha}$.
+We use a Dirichlet-process prior on class assignments, which allows
+the number of classs to be determined automatically. In brief, for $N$
+total cells, the probability of a cell belonging to a class is
+proportional to the number of datapoints already in that class, $N_k$,
+such that $p(c_i = k) \propto \frac{m_k}{N + \alpha}$ and the
+probability of the cell belonging to a new class $k'$ is $p(c_i = k')
+\propto \frac{\alpha}{N + \alpha}$. $\alpha$ is the global
+concentration parameter -- larger values of $\alpha$ make the model
+more likely to propose new classes. We grid the parameter $\alpha$ and
+allow the best value to be learned from the data.
 
@@
 motivating us to use an auxiliary variable method \autocite{Neal2000}
-in which a collection of ephemeral clusters are explicitly represented
+in which a collection of ephemeral classs are explicitly represented
 for the duration of the Gibbs scan.
@@
 tuple of possible values, we explore the cartesian product of all
-possible values. We then Gibbs sample, which is always possible in a 
-small, finite discrete state space.
+possible values. We then Gibbs sample (our final transition kernel),
+which is always possible in a small, finite, discrete state space.
 
@@
 
-\subsection{Validation}
+\todo{Validation: do we want to write extensively about posterior
+  enumeration tests, etc. ? }
 
 \subsection {Mouse Retina}
-Dense serial electron microscopy of a $VOLUMNE$ in the mouse retina by
-\autocite{Helmstaedter2013} yielded a listing of places where neurons
-come into contact. There were over 1000 cells originally, and selected
-the $950$ for which the location of the soma could be reconstructed
-from the provided cell plots (soma locations were not provided by the
-study's authors in machine-readable form). Ultimately
-this left a matrix between the total synapse-like contact area between
-all pairs of 950 cells. Area was thresholded at $0.5\mu m$, determined
-by hand, to yield a 950 $\times$ 950 entry matrix that served as input
-to our algorithm. We measured the distance between cells using
-the reconstructed soma centers, and used the Logistic-Distance
-spatial relation. Hyperprior griddings are shown in \ref{supplemental}.
+\label{supp:mouseretina}
+Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
+in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
+of places where neurons come into contact. There were over 1000 cells
+originally, and selected the $950$ for which the location of the soma
+could be reconstructed from the provided cell plots (soma locations
+were not provided by the study's authors in machine-readable
+form). Ultimately this left a matrix between the total synapse-like
+contact area between all pairs of 950 cells. Area was thresholded at
+$0.5\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
+matrix that served as input to our algorithm. We measured the distance
+between cells using the reconstructed soma centers, and used the
+Logistic-Distance spatial relation. Hyperprior griddings are shown in
+supplemental section \ref{supp:hyperpriors}.
 
@@
 
-We obtained the connectome of c. elegans from \autocite, and isolated
+We obtained the connectome of c. elegans from \autocite{}, and isolated
 the 279 nonpharyngeal neurons, with a total of 6393 chemical synapses
@@
 setting the maximum probability $p$ of the logistic function on a
-per-component basis, drawing from a global $p \sim Beta(\alpha_{hp},
+per-component basis, drawing from a global $p \sim \Beta(\alpha_{hp},
 \beta_{hp})$. Then $\lambda$ is set for each component as a global
@@
 
-The ``logistic-distance Poisson'' spatial model assumes that, if cell
-$e_i$ is of class $m$ and cell $e_j$ is of class $n$, then $\eta_{mn}
-= (\mu_{mn}, r_{mn})$, and the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
+The ``logstic-distance Poisson'' spatial model is used to explicitly mode the count of synapses, $c$, between two neurons.
+The probability of c synapses between two neurons is distributed $c \sim \textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a scaled logistic function (the logistic function has range $[0, 1]$.
+For each component \eta_{mn} we learn both the threshold $\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$ and $n$ are likely to have on average $20$ synases if they are closer than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$  "
+
+Thus the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
 \begin{eqnarray}
-\lambda^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
-\lambda & = & \lambda^* \cdot (r_{mn} - r_{min}) + r_{min} \\
-R(e_i, e_j) \sim Poi(c | \lambda)
+r^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
+r & = & r^* \cdot (r_{mn} - r_{min}) + r_{min} \\
+R(e_i, e_j) \sim \textrm{Poisson}(c | r)
 \end{eqnarray}
@@
 All source code and materials for running experiments can be
-obtained from the project website, at http://spatialrelation.github.io.
+obtained from the project website, at \\
+
+\href{http://spatialrelation.github.io }{http://spatialrelation.github.io}
 
@@
 This is a matrix consisting of the different clusters of nets for 
-the registers, with the colors consistent with figure~\ref{mos6502}
+the registers, with the colors consistent with figure~\ref{fig:mos6502}.
 
@@
 
-We compare the clusters found by our model, the traditional infinite
+We compare the classs found by our model, the traditional infinite
 stochastic block model, and a finite stochastic block model with the number
-of clusters set to the best-found number from our iSRM model ($k=12$).
+of classes set to the best-found number from our iSRM model ($k=12$).
 
 We see from fig~\ref{fig:supp:irmspatialcomparemouse} that the spatial
-variance of our clusters closely matches the ground truth, while the
+variance of our classes closely matches the ground truth, while the
 other two models find spatially-localized
-clusters. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
+classs. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
 clustering, the clustering found by the iSBM,

### 2013-12-27 docs/paper.tex
updates to figs fixing last-min bugs

@@
   \caption{An overview of our method. a. Input data is
-an adjacency matrix indicating whether cell $i$ (row)
+a connectivity matrix indicating whether cell $i$ (row)
 synapses on cell $j$ (column). The location of each cell in space is known. 
@@
 profile, c.), with cells of other types. This results in a clustering 
-of the cells by those hidden types. b.) shows the cell adjacency
+of the cells by those hidden types. b.) shows the cell connectivity
 matrix with cells of the same type grouped together. d.) shows
@@
 probability of connections between cells, a so-called generative
-model. We term our model the ``infinite spatial-relational model''. 
+model. We term our model the ``infinite spatial-relational model'' (iSRM), 
+in keeping with the existing nonparametric literature.  
 The probability of connections between pairs of neurons
@@
   \caption{Discovering cell classes in the mouse retina connectome. 
-a.) Input adjacency data for 950 cells for which soma positions were known. b. clustered adjacency matrix. c.) connectivity diagram showing our clusters, as
+a.) Input connectivity data for 950 cells for which soma positions were known. b. clustered connectivity matrix. c.) connectivity diagram showing our clusters, as
 well as the cell depth and known cell type of each cell. Our model agrees
@@
   \centering 
-  \centerline{\includegraphics{celegans.ai}}
+  \centerline{\includegraphics[width=5in]{celegans.ai}}
   \caption{Discovering cell types in the \textit{c. elegans}
@@
 
-  Larger datasets will generally allow algorithms to distinguish more
-  distinct types (cite supplemental data) and more data should
-  algorithms to get closer to the results from anatomy. Moreover, in
-  general, for such problems precision increases with the size of the
-  dataset and the cells that we have are not sufficient to
-  statistically distinguish all the cell types known in anatomy
-  ($\sim 70$). Still, using only connectivity and distance it is possible
-  to meaningfully divide neurons into types.
+  Larger datasets will allow algorithms to distinguish more distinct
+  types \todo{Add supplemental data} and we expect closer agreement
+  with existing anotomical knowledge as more data become available.
+  Moreover, in general, for such problems precision increases with the
+  size of the dataset and the cells that we have are not sufficient to
+  statistically distinguish all the cell types known in anatomy (such
+  as the $\sim 70$ in the retina). Still, using only connectivity and
+  distance it is possible to meaningfully divide neurons into types.
 
@@
 
-There have existed a range of previous approaches to the discovery of
+There exist a range of previous approaches to the discovery of
 neural microcircuitry \todo{Do we want a citation here?}. These
@@
 \autocite{Sincich2005,Lennie2005}. This would allow comparative
-quantitative anatomy across animals, and the search for the causes of
-the ultimate connectivity.
+quantitative anatomy across animals, and aid the search for the
+ultimate causes of connectivity.
 
@@
 matrices. We assume there exist an unknown number $K$ of latent
-(unobserved) cell types, $k \in \{1, 2, 3, \dots\, K}$, and that each
+(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
 cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
@@
 $f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
-latent class, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
 as a set of global hyper parameters $\theta$, such that the link
@@
 The ``logistic-distance Bernoulli'' spatial model assumes that, if cell
-$e_i$ is of class $m$ and cell $e_j$ is of class $n$, then $\eta_{mn}
+$e_i$ is of type $m$ and cell $e_j$ is of type $n$, then $\eta_{mn}
 = (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
@@
 \end{eqnarray}
-where $p_{max}$ and $p_{min}$ are per-graph hyperparameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})$.
+where $p_{max}$ and $p_{min}$ are global per-graph parameters.
+
+We place an exponential priors on the latent parameters:
+\begin{eqnarray}
+ \mu_{mn} \sim \exp(\mu | \mu^{hp}) \\
+\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})
+\end{eqnarray}
+
+using  $\lambda^{hp}$ and $\mu^{hp}$ as global per-graph hyperparameters.
 
@@
 ongoing area of research in the probabilistic modeling community
-\autocite{missing}.
+\todo{cite more? missing}.
 
@@
 
-We obtained the connectome of c. elegans from \autocite{}, and isolated
-the 279 nonpharyngeal neurons, with a total of 6393 chemical synapses
-and 890 gap junctions originally cleaned up in \autocite{Chen2006}. A
-cell's position was its distance along the anterior-posterior axis
-normalized between zero and one. We used both networks, the chemical
-network as a directed graph and the electrical network as undirected
-graph. We use the synapse counts with the logistic-distance poisson
-likelihood, scaling the counts by 4.0 to compensate for 
-the Poisson's overdispersion.
+We obtained the connectome of c. elegans from data published
+previously \autocite{Varshney2011}, and isolated the 279 nonpharyngeal
+neurons, with a total of 6393 chemical synapses and 890 gap junctions
+originally cleaned up in \autocite{Chen2006}. A cell's position was
+its distance along the anterior-posterior axis normalized between zero
+and one. We used both networks, the chemical network as a directed
+graph and the electrical network as undirected graph. We use the
+synapse counts with the logistic-distance poisson likelihood, scaling
+the counts by 4.0 to compensate for the Poisson's overdispersion.
 
@@
 drain, leading to ambiguity in our encoding. We identified a region
-consisting of consisting of three registers X, Y, and S via visual
+consisting of three registers X, Y, and S via visual
 inspection and focused our efforts there. We created a total of six
@@
 We reparameterized the Logistic-Distance Bernoulli likelihood to
-better capture the microprocessor data structure. We explicitly
+better capture the microprocessor data structure. We are explicitly
 setting the maximum probability $p$ of the logistic function on a
-per-component basis, drawing from a global $p \sim \Beta(\alpha_{hp},
+per-component basis, drawing from a global $p \sim \operatorname{Beta}(\alpha_{hp},
 \beta_{hp})$. Then $\lambda$ is set for each component as a global
@@
 The probability of c synapses between two neurons is distributed $c \sim \textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a scaled logistic function (the logistic function has range $[0, 1]$.
-For each component \eta_{mn} we learn both the threshold $\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$ and $n$ are likely to have on average $20$ synases if they are closer than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$  "
+For each component $\eta_{mn}$ we learn both the threshold $\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$ and $n$ are likely to have on average $20$ synases if they are closer than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$  "
 
@@
     \centerline{\includegraphics[width=5in]{mouseretina/spatial_var.pdf}}
-  \caption{}
+  \caption{The distribution of the spatial extent of cell types found by each of
+the above methods, and ground truth clusters}
 \label{fig:supp:irmspatialcomparemouse}
@@
     \centerline{\includegraphics[width=5in]{mouseretina/ari_vs_cluster.pdf}}
-  \caption{}
+  \caption{ARI vs cluster number for the iSRM, iSBM, and the SBM with varying K.}
 \label{fig:supp:mousecompareari}
@@
   \centering 
-    \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{}
+    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
+  \caption{The cell IDs and found types (grouped) vs known types (y-axis)  found by the infinite stochastic block model, dramatically overclustering relative to ground truth}
 \label{fig:supp:retinabbclusters}
@@
   \centering 
-    \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
-  \caption{}
-\label{fig:supp:retinabbfixedclusters}
+    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
+  \caption{The cell IDs and found types (grouped) vs known types (y-axis) by the infinite spatial-relational model.}
+\label{fig:supp:retinaldclusters}
 \end{figure}
@@
   \centering 
-    \centerline{\includegraphics[width=6.5in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{}
-\label{fig:supp:retinaldclusters}
+    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
+  \caption{The cell IDs and found types (grouped) vs known types (y-axis) via the stochastic block model with $k=12$}
+\label{fig:supp:retinabbfixedclusters}
 \end{figure}

### 2013-12-27 experiments/mouseretina/pipeline.py

@@ -1109,14 +1109,13 @@ def plot_truth_latent(exp_results,
     COL_PRE = 0.02
     ROW_CONTRACT = 0.05
     ROW_PRE = 0.02
-    ROW_SPACE = 0.01
-
-    VERT_SCALE = 1.0
+    ROW_SPACE = 0.005
+    ROW_HEIGHT_MIN = 0.015
     s = df['cluster'].value_counts()
 
     a = irm.util.multi_napsack(COL_NUMBER, np.array(s))
     CLUST_MAX_PER_COL = len(a[0])
-    VERT_SCALE = 1.0 -  (CLUST_MAX_PER_COL+4) *ROW_SPACE
+    VERT_SCALE = 0.95 -  (CLUST_MAX_PER_COL+4) *ROW_SPACE
 
     MAX_LEN = np.sum([np.array(s)[ai] for ai in a[0]])
 
@@ -1127,7 +1126,7 @@ def plot_truth_latent(exp_results,
             sub_df = df[df['cluster'] == cluster_id]
             sub_df = sub_df.sort('cell_type')
             height = len(sub_df) / float(MAX_LEN) * VERT_SCALE
-
+            height = np.max([height, ROW_HEIGHT_MIN])
             ax = f.add_axes([COL_PRE + col_i * COL_SPACE, 
                              1.0 - pos - height - ROW_PRE, 
                              COL_WIDTH, height])
@@ -1156,7 +1155,11 @@ def plot_truth_latent(exp_results,
             ax.set_yticks([])
             ax.grid(1)
             ax.set_xlim(0, TYPE_N)
-            ax.set_ylim(-0.5, CN+0.5)
+            if CN > 3:
+                ax.set_ylim(-1, CN+0.5)
+            else:
+                ax.set_ylim(-1, 3)
+
             ax.set_xticks([10, 20, 30, 40, 50, 60, 70])
 
 

### 2013-12-27 experiments/regirmfail/pipeline2.py

@@ -591,15 +591,17 @@ def plot_results(infile, outfiles):
                    'ld' : 'r'}
         f = pylab.figure(figsize=(4, 3))
         ax = f.add_subplot(1, 1, 1)
-
+        labels = {'bb' : "iSBM", 
+                  'ld' : "iSRM"}
         for g_idx, g in a.groupby(['model']):
-            print g_idx
             ax.scatter(g.index.get_level_values('class_n'), g['empirical_class_n'], c=colors[g_idx],
-                          edgecolor='none')
-        ax.plot([1, 16], [1, 16], c='k')
-        ax.set_xlabel("true class number")
-        ax.set_ylabel("estimated class number")
+                          edgecolor='none', label= labels[g_idx])
+        ax.plot([1, 16], [1, 16], c='k', label="ground truth")
+        ax.set_xlabel("true type number")
+        ax.set_ylabel("estimated type number")
         ax.set_xticks([1, 2, 4, 8, 16])
+        ax.legend(loc="upper left", fontsize=10)
+                  
         f.tight_layout()
         f.savefig(plot_files[0])
 
@@ -621,11 +623,12 @@ def plot_results(infile, outfiles):
             ax.bar(np.arange(N)*CLASS_SPACE + offsets[g_idx], h['ari'], width=WIDTH, 
                     color=colors[g_idx])
             ax.errorbar(np.arange(N)*CLASS_SPACE + offsets[g_idx] + WIDTH/2, 
-                        h['ari'], yerr= herr['ari'], capsize=0,elinewidth=4,ecolor='k', linewidth=0)
+                        h['ari'], yerr= herr['ari'], capsize=0,elinewidth=3, linewidth=0, ecolor='grey')
         #ax.plot([1, 16], [1, 1], c='k')
-        ax.set_xlabel("true class number")
-        ax.set_ylabel("adjusted rand index")
+        ax.set_xlabel("true type number")
+        ax.set_ylabel("adjusted Rand index")
         ax.set_ylim(0, 1.0)
+        ax.set_yticks([0.0, 1.0])
         ax.set_xticks(np.arange(N)*CLASS_SPACE + 1)
         ax.set_xticklabels([1, 2, 4, 8, 16])
         f.tight_layout()
@@ -692,8 +695,8 @@ def plot_results(infile, outfiles):
             if i == 0:
                 handles, labels = ax.get_legend_handles_labels()
                 ax.legend(handles, [    'Ground Truth', 
-                                        'Relational Model', 
-                                        'Spatial Relational Model', 
[…]

### 2013-12-31 docs/paper.tex
reasonably good version before we go hacking it up for submission
updates

@@
 \usepackage[disable]{todonotes}
-\usepackage{subfiles}
 \usepackage{grffile}
@@
 
+%\renewcommand{\includegraphics}[2][]{\fbox{}}
+
 \addbibresource{connectomics motif.bib}
@@
 
-\title{Discovering cell type motifs in connectomics data}
-\author{Eric Jonas, Konrad Kording}
+\title{Automatic discovery of circuit structure from neural connectomics}
+\author{Eric Jonas$^1$, Konrad Kording$^{2, 3,4}$}
 
@@
 
-% \begin{affiliations}
-%   \item Something Something
-%   \item Something Something
-% \end{affiliations}
+\begin{small}
+\begin{enumerate}
+  \item Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Cambridge, Massachusetts
+  \item Department of Physical Medicine and Rehabilitation, Northwestern University and Rehabilitation Institute of Chicago, Chicago, Illinois
+  \item Department of Physiology, Northwestern University, Chicago, Illinois
+  \item Department of Applied Mathematics, Northwestern University, Chicago, Illinois
+\end{enumerate}
+\end{small}
 
@@
 
-\begin{figure}[ht!]
+The roundworm Caenorhabditis elegans is a model system in
+developmental neuroscience\autocite{White1986}, with the location and
+connectivity of each of 302 neurons deterministically known. This
+incredible cross-organism consistency led to early mapping of the
+c. elegans chemical and electrical (gap junction) connectome. Unlike
+the retina, only the motor neurons in c. elegans exhibit regular
+distribution in space, being distributed along the body axis. Most
+interneurons are concentrated in various ganglia that project
+throughout the entire animal, and the sensory neurons are primarily
+located in a small number of anterior ganglia.
+
+\begin{figure}
   \centering 
@@
 \end{figure}
-
-The roundworm Caenorhabditis elegans is a model system in
-developmental neuroscience\autocite{White1986}, with the location and
-connectivity of each of 302 neurons deterministically known. This
-incredible cross-organism consistency led to early mapping of the
-c. elegans chemical and electrical (gap junction) connectome. Unlike
-the retina, only the motor neurons in c. elegans exhibit regular
-distribution in space, being distributed along the body axis. Most
-interneurons are concentrated in various ganglia that project
-throughout the entire animal, and the sensory neurons are primarily
-located in a small number of anterior ganglia.
 
@@
 
+\subsection{Methods Summary}
+
+We take as input a connectivity matrix $R$ defining the connections
+between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
+e_j)$ representing a (physical) distance between adjacent cells. See
+the supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number $K$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
+cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
+of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
+observed connectivity between two cells $R(e_i, e_j)$ then depends
+only on their latent type and their distance through a link function
+$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
+latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+as a set of global hyper parameters $\theta$, such that the link
+function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
+
+We then jointly infer the maximum a posteriori (MAP) estimate of the
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
+the global model hyperparameters $\theta$ :
+
+\begin{equation}
+  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
+\end{equation}
+
+Inference is performed in three steps via composible transition 
+kernels -- one for structural, one for per-type parameters, and
+one kernel for global parameters and hyperparameters. Details
+of data preprocessing, inference parameters, and runtime can
+be found in the Methods section.
+
 \printbibliography
+
+\begin{itemize}
+
+ \item \textbf{Acknowledgments} We thank Josh Vogelstein for discussions and reading of the manuscript, Finale Doshi-Velez for early discussions on the model, and Erica Peterson, and Jonathan Glidden for extensive manuscript review.
+\item \textbf{Author Contributions} KK and EJ developed model. EJ derived inference, implemented code, tested, and ran experiments. KK and EJ wrote manuscript text and solicited feedback. 
+ \item \textbf{Competing Interests} The authors declare that they have no
+competing financial interests.
+ \item \textbf{Correspondence} Correspondence and requests for materials
+should be addressed to E.J.~(email: jonas@mit.edu).
+\end{itemize}
 
@@
 \section{Methods }
-\subsection{Methods Summary}
+
+\subsection{Probabilistic Model}
+
+Our model is a extension of the iSBM
+\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
+inspired by attempts to extend these models with arbitrary
+discriminative functions\autocite{Murphy2012}.
 
@@
 % methods section can have no figures or tables [grrr]
-
-\subsection{Probabilistic Model}
-
-Our model is a extension of the iSBM
-\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
-inspired by attempts to extend these models with arbitrary
-discriminative functions\autocite{Murphy2012}.
 
@@
 \label{supp:mos6502netlist}
-This is a matrix consisting of the different clusters of nets for 
+This is a matrix consisting of the different clusters of electrical nets for 
 the registers, with the colors consistent with figure~\ref{fig:mos6502}.

### 2014-01-02 experiments/mouseretina/diagnostic_plots.py
updates to paper proper

@@ -246,7 +246,7 @@ def plot_adjmat_byclass((type_file, pos_file, xlsxdata_file), (adj_mat_plot,)):
 
 
 @files(["type_metadata.pickle", "soma.positions.pickle", "xlsxdata.pickle"], 
-       ['adjmat.byz.png'])
+       ['adjmat.byz.jpg'])
 def plot_adjmat_byz((type_file, pos_file, xlsxdata_file), (adj_mat_plot,)):
     soma_pos = pickle.load(open(pos_file, 'r'))
     type_metadata = pickle.load(open(type_file, 'r'))['type_metadata']
@@ -279,7 +279,7 @@ def plot_adjmat_byz((type_file, pos_file, xlsxdata_file), (adj_mat_plot,)):
     ax.set_ylim(CELL_N, 0)
     ax.set_title("mouse retina connectivity matrix")
     f.tight_layout()
-    f.savefig(adj_mat_plot, dpi=200)
+    f.savefig(adj_mat_plot, dpi=70)
 
 
 

### 2014-01-03 docs/getgraphics.py
updates to build system so that we can build for the arxiv

@@ -1,0 +1,11 @@
+import numpy as np
+import re
+import sys
+fname = sys.argv[1]
+
+texfile = open(fname, 'r').read()
+
+r = re.compile("\\includegraphics(\[.+\])*\{([\w0-9\./\-]+)\}")
+g = [a[1] for a in r.findall(texfile)]
+for f in g:
+    print f, 

### 2014-01-06 docs/paper.tex
updates to paper proper
longer final c elegans run

@@
 \usepackage{fullpage}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+%\usepackage{algorithm}
+%\usepackage{algorithmic}
 \usepackage{amsmath}
-\usepackage{subcaption}
-\usepackage{multicol}
+%\usepackage{subcaption}
+%\usepackage{multicol}
 \usepackage{csquotes}
@@
   New techniques produce massive data about neural connectivity,
-  necessitating new analysis methods to discover underlying
-  structure. It has long been assumed that discovering the local
-  patterns of microcircuitry is the key to understanding neural
-  function. Here we developed a nonparametric Bayesian technique that
-  identifies neuron types and circuitry patterns in connectomics
-  data. We show that the approach recovers known neuron types in the
-  retina, reveals interesting structure in the nervous system of
-  c. elegans, and automatically discovers the structure of
-  microprocessors. Our approach extracts structural meaning from
-  connectomics, enabling new approaches of deriving anatomical
-  insights from these emerging datasets.
+  necessitating new analysis methods to discover the biological and
+  computational basis of this connectivity. It has long been assumed
+  that discovering the local patterns of microcircuitry is the key to
+  understanding neural function. Here we developed a nonparametric
+  Bayesian technique that identifies neuron types and microcircuitry
+  patterns in connectomics data. We show that the approach recovers
+  known neuron types in the retina, reveals interesting structure in
+  the nervous system of c. elegans, and automatically discovers the
+  structure of microprocessors. Our approach extracts structural
+  meaning from connectomics, enabling new approaches of deriving
+  anatomical insights from these emerging datasets.
 
@@
 another. In fact, the idea of well defined local connectivity patterns
-has been prominent in many areas, ranging from sensory (e.g. retina,
-\autocite{Masland2001} via processing (e.g. neocortex \autocite{Mountcastle1997})
-to movement (e.g. spinal cord) \autocite{Grillner2005}. Indeed, such
-repeated patterns also exists in human-made computing circuits. It
-remains an important challenge to develop algorithms to use anatomical
-data, e.g. connectomics, to back out underlying connectivity motifs.
+(microcircuits) has been prominent in many areas, ranging from sensory
+(e.g. retina, \autocite{Masland2001} via processing (e.g. neocortex
+\autocite{Mountcastle1997}) to movement (e.g. spinal cord)
+\autocite{Grillner2005}. Indeed, such repeated patterns also exists in
+human-made computing circuits. It remains an important challenge to
+develop algorithms to use anatomical data, e.g. connectomics, to back
+out underlying microcircuitry.
 
@@
 \section{Results}
-We use a Bayesian structure discovery approach that is based on
-prior knowledge about the way positions and cell types affect the
+We use a Bayesian structure discovery approach that is based on prior
+knowledge about the way positions and cell types affect the
 probability of connections between cells, a so-called generative
-model. We term our model the ``infinite spatial-relational model'' (iSRM), 
-in keeping with the existing nonparametric literature.  
-The probability of connections between pairs of neurons
-decreases with distance, but how high the baseline probability of
-connection is and how quickly it decays with distance depends on the
-two cell-types. Note that this simplistic prior information is all that 
-we tell our model; no additional physiological information is included.
+model. Basing our model on previous infinite stochastic block models
+(iSBM)\autocite{Kemp2006a,Xu2006}, we assume that each neuron is a
+member of a single ``type'', and that the probability of two neurons
+being connected is a function of their types. Crucially, we
+incorporate the prior knowledge that spatial proximity matters, adding
+distance-dependence to the iSBM.  For example, retinal ganglion cells
+may synapse on nearby, but not far away, amacrine cells.  Note that
+this simplistic prior information is all that we tell our model; no
+additional physiological information is included.  We term our model
+the ``infinite spatial-relational model'' (iSRM), in keeping with the
+existing nonparametric literature, as it discovers a potentially
+unbounded number of unobserved, or ``latent'', cell types.
 
@@
 one we used for simulation. We find that the model does a good job of
-recovering the structure we used to generate the data
-(fig~\ref{fig:synthetic}a) Indeed, it does far better than a simpler
-and well established model (the infinite stochastic block model, or
-iSBM) that assumes that there is no distance dependence to the
-connectivity. The model converges relatively quickly to an estimate of
-the most probable values for the cell types, which is enabled by using
-a combination of simulated annealing and parallelized Markov chain
-Monte Carlo (see methods for details). It appears that the model
-behaves the way it should and allows application to relatively large
-datasets.
+recovering the correct number of cell types,
+(fig~\ref{fig:synthetic}a), the the cell identities
+(fig~\ref{fig:synthetic}b), and the spatial extent of each type
+(fig~\ref{fig:synthetic}c).  Indeed, it does far better than the
+existing infinite stochastic block model that assumes that there is no
+distance dependence to the connectivity. The model converges
+relatively quickly to an estimate of the most probable values for the
+cell types, which is enabled by using a combination of simulated
+annealing and parallelized Markov chain Monte Carlo (see methods for
+details). It appears that the model behaves the way it should and
+allows application to relatively large datasets.
 
@@
   \caption{Correct recovery of true numbers of hidden classes
-in synthetic data. a.) Other models over-cluster the number
-of classes as they fail to take distance into account, whereas
+in synthetic data. a.) The iSBM over-segments the number
+of classes as it fails to take distance into account, whereas
 our model finds close to the true number of classes. b.) ARI, a measure
@@
 cells which bore little relation to known types.  The average spatial
-extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. However,
-in the ground truth by anatomists the average distance was $34.3\pm7.6
-\mu m$ and $35.3\pm 2.4 \mu m$, quite close to the $34.5\pm 1.0 \mu m$
-discovered by our algorithm. Modeling the distance-dependence prevents
+extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. Using
+the anatomists' ground truth, the mean spatial extent of types was
+$34.3\pm7.6 \mu m$, quite close to the $34.5\pm 1.0 \mu m$ discovered
+by our algorithm. Modeling the distance-dependence prevents
 over-segmentation and thus allows more meaningful estimates.
@@
 microcircuitry our model was designed for.  Motor-neuron types AS, DA,
-and VA, all exclusively postsynaptic, are clustered together, as are
-motor-neuron types VD and DD. Traditional types VC, DB, and VB also
-mostly share a cluster. Various head motor neurons, including types
+VA, and VB are clustered together,
+
+as are
+
+``The processes of VDn and DDn do not have extended, apparently undifferentiated distal regions, as do the other motoneuron classes; instead, they end abruptly in close proximity to the end of a neighbouring process of the same class, often with a gap junction between them''
+
+motor-neuron types VD and DD, and head motor neuron types RMD and SMD.
+
+Various head motor neurons, including types
 SMD, RMD, and RIM are clustered together. Interneurons with known
 anatomically-distinct connectivity patterns, such as RIA (2 cells) or
-AVA (2 cells) are clustered into pure types. The algorithm even
-correctly places the single-cell type DVB by itself. Note our
+AVA/AVB (4 cells) are clustered into pure types.
+
+Note our
 clustering does not perfectly reflect known divisions -- several
@@
  \item \textbf{Acknowledgments} We thank Josh Vogelstein for discussions and reading of the manuscript, Finale Doshi-Velez for early discussions on the model, and Erica Peterson, and Jonathan Glidden for extensive manuscript review.
+Funding for compute time was provided by Amazon Web Services ``AWS in Education'' grants.
 \item \textbf{Author Contributions} KK and EJ developed model. EJ derived inference, implemented code, tested, and ran experiments. KK and EJ wrote manuscript text and solicited feedback. 

### 2014-01-06 experiments/celegans/pipeline.py

@@ -43,6 +43,16 @@ EXPERIMENTS = [
     ('celegans.2r.ldp.01', 'fixed_100_200', 'anneal_slow_400'),  
     ('celegans.2r.ldp.02', 'fixed_100_200', 'anneal_slow_400'),  
     ('celegans.2r.ldp.03', 'fixed_100_200', 'anneal_slow_400'),  
+    ('celegans.2r.edp.00', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.edp.01', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.edp.02', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.edp.03', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.edp.04', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.ldp.00', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.ldp.01', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.ldp.02', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.ldp.03', 'fixed_100_200', 'anneal_slow_800'),  
+    ('celegans.2r.ldp.04', 'fixed_100_200', 'anneal_slow_800'),  
     #('celegans.2r.ldp.00', 'fixed_10_100', 'anneal_slow_20'),  
     #('celegans.2r.edp.00', 'fixed_10_100', 'anneal_slow_20'),  
 
@@ -69,12 +79,14 @@ LD_HPS = [(0.1, 0.1, 0.9, 0.01)] # , (0.1, 0.1, 0.5, 0.001),
 LDFL_HPS = [(0.2, 0.4, 0.01, 1.0, 1.0)] # , (0.1, 0.1, 0.5, 0.001),
 NDFW_HPS = [(0.1, 0.001, 0.1, 1.0, 1.0)]
 SDB_HPS = [(1.0, 1.0, 0.1, 0.001, 0.5, 4.0)]
-EDP_HPS = [(0.1, 1.0, 1.0), (0.1, 1.0, 2.0), (0.1, 1.0, 4.0), (0.1, 1.0, 8.0)]
[…]

### 2014-01-06 docs/paper.tex
konrad updates

@@
   computational basis of this connectivity. It has long been assumed
-  that discovering the local patterns of microcircuitry is the key to
+  that discovering the local patterns of microcircuitry is crucial to
   understanding neural function. Here we developed a nonparametric
@@
 human-made computing circuits. It remains an important challenge to
-develop algorithms to use anatomical data, e.g. connectomics, to back
+develop algorithms to use anatomical data, e.g. connectomics, to automatically back
 out underlying microcircuitry.
@@
 \section{Results}
-We use a Bayesian structure discovery approach that is based on prior
-knowledge about the way positions and cell types affect the
-probability of connections between cells, a so-called generative
-model. Basing our model on previous infinite stochastic block models
-(iSBM)\autocite{Kemp2006a,Xu2006}, we assume that each neuron is a
-member of a single ``type'', and that the probability of two neurons
-being connected is a function of their types. Crucially, we
-incorporate the prior knowledge that spatial proximity matters, adding
-distance-dependence to the iSBM.  For example, retinal ganglion cells
-may synapse on nearby, but not far away, amacrine cells.  Note that
-this simplistic prior information is all that we tell our model; no
-additional physiological information is included.  We term our model
-the ``infinite spatial-relational model'' (iSRM), in keeping with the
-existing nonparametric literature, as it discovers a potentially
-unbounded number of unobserved, or ``latent'', cell types.
-
-We combine annealing with Markov Chain Monte Carlo to
-perform simultaneously joint posterior inference over the structure,
-parameters, and hyperparameters of the model (see Methods). This
-procedure automatically learns the number of cell types, which cells 
-belong to which types, along with the 
-the distance and type dependent connectivity probabilities (see
-fig~\ref{fig:overview}). These results can then be compared with the
-results of previous anatomical studies that have used other aspects of
-neurons, such as chemical properties and cell-morphology to divide
-neurons into types.
+From the generic assumption that cell types exist and that connection probability depends on the involved cell types and distance, we develop a Bayesian structure discovery algorithm.
+We start with existing infinite stochastic block models (iSBM)\autocite{Kemp2006a,Xu2006}, which have been shown to meaningfully cluster connection graphs.
+We extend this approach by adding distance dependence to model salient aspects of microcircuitry.
+For example, retinal ganglion cells may synapse on nearby, but not far away, amacrine cells.
+We term our model the ``infinite spatial-relational model'' (iSRM), in keeping with the existing nonparametric literature.
+This procedure automatically learns the number of cell types, which cells belong to which types, along with the the distance and type dependent connectivity probabilities (see fig~\ref{fig:overview}).
+We apply our algorithm to datasets from mouse retina, c-elegans and a historical microprocessor.
+Anatomists classify cells based on many features, giving us a meaningful baseline to compare against.
 
@@
 \subsection{C. elegans}
-
+[TODO Say this is harder than retina because of these facts]
 The roundworm Caenorhabditis elegans is a model system in

### 2014-01-06 docs/paper.tex
updated c elegans assessment

@@
 \subsection{C. elegans}
-[TODO Say this is harder than retina because of these facts]
+
 The roundworm Caenorhabditis elegans is a model system in
 developmental neuroscience\autocite{White1986}, with the location and
-connectivity of each of 302 neurons deterministically known. This
-incredible cross-organism consistency led to early mapping of the
-c. elegans chemical and electrical (gap junction) connectome. Unlike
-the retina, only the motor neurons in c. elegans exhibit regular
-distribution in space, being distributed along the body axis. Most
-interneurons are concentrated in various ganglia that project
-throughout the entire animal, and the sensory neurons are primarily
-located in a small number of anterior ganglia.
+connectivity of each of 302 neurons developmentally determined,
+leading to early measurement of the connectome. Unlike the retina,
+only the motor neurons in c. elegans exhibit regular distribution in
+space, being distributed along the body axis. Most interneurons are
+concentrated in various ganglia that project throughout the entire
+animal, and the sensory neurons are primarily located in a small
+number of anterior ganglia.
 
@@
 microcircuitry our model was designed for.  Motor-neuron types AS, DA,
-VA, and VB are clustered together,
+and VA, all exclusively postsynaptic, are identified as a common type,
+as are motor-neuron types VD and DD. Traditional types VC, DB, and VB
+also mostly share a cluster. Various head motor neurons, including
+types SMD and RMD, are clustered together. Interneurons with known
+anatomically-distinct connectivity patterns, such as AVA (2 cells),
+are clustered into pure types.The algorithm even correctly places the
+single-cell types DVB and DVC by themselves.
 
-as are
-
-``The processes of VDn and DDn do not have extended, apparently undifferentiated distal regions, as do the other motoneuron classes; instead, they end abruptly in close proximity to the end of a neighbouring process of the same class, often with a gap junction between them''
-
-motor-neuron types VD and DD, and head motor neuron types RMD and SMD.
-
-Various head motor neurons, including types
-SMD, RMD, and RIM are clustered together. Interneurons with known
-anatomically-distinct connectivity patterns, such as RIA (2 cells) or
-AVA/AVB (4 cells) are clustered into pure types.
-
-Note our
-clustering does not perfectly reflect known divisions -- several
-combinations of head and sensory neurons are combined, and a
-difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
-grouped together. Our identified cell types thus reflect a
-``coarsening'' of known types, based entirely on connectivity and
-distance information, even when the organism exhibits substantially
-less spatial regularity than the retina.
+Note our clustering does not perfectly reflect known divisions --
+several combinations of head and sensory neurons are combined, and a
+difficult-to-explain group of mostly VB and DB motor neuron types,
+with VC split between various groups. Our identified cell types thus
+reflect a ``coarsening'' of known types, based entirely on
+connectivity and distance information, even when the organism exhibits
+substantially less spatial regularity than the retina.
 

### 2014-01-06 experiments/celegans/pipeline.py

@@ -1175,14 +1175,14 @@ def plot_best_circos(exp_results,
                                                    sample_latent['relations'][relation]['hps'], 
                                                    model_name)
                 rate_scale = {'R1' : 1.0, 
-                              'R2' : 2.0}
-                thold = 0.5
+                              'R2' : 1.0}
+                thold = 2.0
                 ribbons = []
                 links = []
                 for (src, dest), rate in v.iteritems():
                     scaled_rate = rate * rate_scale[relation]
                     if scaled_rate > thold:
-                        pix = int(5*scaled_rate)
+                        pix = int(3*scaled_rate)
                         print src, dest, rate, pix
 
                         ribbons.append((src, dest, pix))

### 2014-01-07 docs/paper.tex
updates to paper; added extra figs

@@
 
-\todo{Do we assign any of the unknown cells correctly?  are types 72-79
-“different” or simply unknown?}
-
 We compared the results of our model to the traditional iSBM to see if
@@
 only the motor neurons in c. elegans exhibit regular distribution in
-space, being distributed along the body axis. Most interneurons are
+space -- along the body axis. Most interneurons are
 concentrated in various ganglia that project throughout the entire
@@
 
+\newpage
+
+\FloatBarrier
+\subsection{Detailed learned connectivity for mouse retina}
+
+Here we show the recovered distance-dependent connectivty profile
+versus the empirical distance-dependent connectivty for the
+retina dataset. The Logistic-distance Poisson likelihood closely
+matches the empirical likelihood for most classes.
+
+\begin{figure}[h!]
+  \centering 
+    \centerline{\includegraphics[width=7in,page=2]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}}
+  \caption{Learned connectivity vs distance between each of the discovered cell types in the mouse retina.
+The rows are the source type, and the columns are the destination type.
+Blue is the empirical plot of the probability of connection vs distance for those two types; red is the learned function}.
+
+\label{fig:supp:mousecompareari}
+\end{figure}
+
 \FloatBarrier

### 2014-01-07 irm/plot.py

@@ -212,7 +212,7 @@ def plot_t1t1_params(fig, conn_and_dist, assign_vect, ss, hps, MAX_DIST=10,
 
                 p = htrue.astype(float) / (hfalse + htrue)
 
-                ax.plot(bins[:-1], p)
+                ax.plot(bins[:-1], p, c='b', linewidth=3)
 
 
             if model == "LogisticDistance":
@@ -220,7 +220,7 @@ def plot_t1t1_params(fig, conn_and_dist, assign_vect, ss, hps, MAX_DIST=10,
                 print "MAX_DISTANCE=", MAX_DIST, np.max(fine_bins), np.max(bins), c
                 y = util.logistic(fine_bins, c['mu'], c['lambda']) 
                 y = y * (hps['p_max'] - hps['p_min']) + hps['p_min']
-                ax.plot(fine_bins, y, c='r') 
+                ax.plot(fine_bins, y, c='r', linewidth=2) 
                 ax.text(0, 0.2, r"mu: %3.2f" % c['mu'], fontsize=4)
                 ax.text(0, 0.6, r"lamb: %3.2f" % c['lambda'], fontsize=4)
                 ax.axvline(c['mu'], c='k')

### 2014-01-07 docs/paper.tex
updated paper text

@@
 
-\title{Automatic discovery of circuit structure from neural connectomics}
+\title{Automatic discovery of cell types and microcircuitry from neural connectomics}
 \author{Eric Jonas$^1$, Konrad Kording$^{2, 3,4}$}
@@
 \newpage
-\section{Supplemental}
+\section{Supplemental Material}
 
@@
 
-\href{http://spatialrelation.github.io }{http://spatialrelation.github.io}
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/netmotifs/}
+
+Please contact the author for pre-publication access.
 

### 2014-01-08 docs/paper.tex
oops

@@
 between different cells, and amacrine cells that feed into the
-ganglion cells \autocite{cite some review}. Recent research
+ganglion cells. Recent research
 \autocite{Helmstaedter2013} has produced a large dataset containing
@@
 summarizing the posterior over such a complex state space, which is an
-ongoing area of research in the probabilistic modeling community
-\todo{cite more? missing}.
+ongoing area of research in the probabilistic modeling community.
 

### 2014-01-08 irm/componentmodels.h
added NICH

@@ -1938,6 +1938,137 @@ struct LogisticDistancePoisson {
 
 }; 
 
+struct NormalInverseChiSq { 
+    typedef float value_t; 
+    
+    struct suffstats_t { 
+        uint32_t count; 
+        float mean; 
+        float var; 
+    }; 
+
+    struct hypers_t { 
+        float mu; 
+        float kappa; 
+        float sigmasq; 
+        float nu; 
+    }; 
+    
+    static void ss_sample_new(suffstats_t * ss, hypers_t * hps, 
+                              rng_t & rng) { 
+        ss->count = 0; 
+        ss->mean = 0; 
+        ss->var = 0; 
+    }
+    template<typename RandomAccessIterator>
+    static void ss_add(suffstats_t * ss, hypers_t * hps, value_t val, 
+                       dppos_t dp_pos, RandomAccessIterator data) {
+        ss->count++; 
+        float delta = val - ss->mean; 
+        ss->mean += delta/ss->count; 
+        ss->var += delta * (val - ss->mean); 
+
+    }
+    template<typename RandomAccessIterator>
+    static void ss_rem(suffstats_t * ss, hypers_t * hps, value_t val, 
+                       dppos_t dp_pos, RandomAccessIterator data) {
+        float total = ss->mean * ss->count; 
+        float delta = val - ss->mean; 
+        ss->count--; 
+        if (ss->count == 0) { 
+            ss->mean = 0; 
+        } else { 
+            ss->mean = (total - val) / ss->count; 
+        }
+
+        if (ss->count <= 1) {
+            ss->var = 0; 
+        } else {
+            ss->var -= delta * (val - ss->mean); 
+        }
+
+        
+    }
+
+    static std::tuple<float, float, float, float> 
+    intermediates(hypers_t * hps, 
+                  suffstats_t * ss) { 
+        float total = ss->mean * ss->count; 
+        float mu_1 = hps->mu - ss->mean; 
+        float kappa_n = hps->kappa + ss->count; 
+        float mu_n = (hps->kappa * hps->mu + total) / kappa_n; 
+        float nu_n = hps->nu + ss->count; 
+        float sigmasq_n = 1. / nu_n * (
+                                       hps->nu * hps->sigmasq 
+                                       + ss->var
[…]

### 2014-01-08 irm/componentslice.cc

@@ -234,6 +234,18 @@ void slice_sample_exec<BetaBernoulli>
 }
 
 template<>
+void slice_sample_exec<NormalInverseChiSq>
+(rng_t & rng, float width, 
+ NormalInverseChiSq::suffstats_t * ss, 
+ NormalInverseChiSq::hypers_t * hps, 
+ std::vector<NormalInverseChiSq::value_t>::iterator data,
+ const std::vector<dppos_t> & dppos, 
+ float temp){
+    // Doesn't do anything because the model is conjugate
+
+}
+
+template<>
 void slice_sample_exec<GammaPoisson>
 (rng_t & rng, float width, 
  GammaPoisson::suffstats_t * ss, 

### 2014-01-08 irm/componentslice.h

@@ -142,6 +142,15 @@ void slice_sample_exec<LogisticDistancePoisson>
  const std::vector<dppos_t> & dppos,
  float temp);
 
+template<>
+void slice_sample_exec<NormalInverseChiSq>
+(rng_t & rng, float width, 
+ NormalInverseChiSq::suffstats_t * ss, 
+ NormalInverseChiSq::hypers_t * hps, 
+ std::vector<NormalInverseChiSq::value_t>::iterator data,
+ const std::vector<dppos_t> & dppos, 
+ float temp);
+
 }
 
 #endif

### 2014-01-08 irm/irmio.py

@@ -294,6 +294,8 @@ def estimate_suffstats(irm_model, rng, ITERS=10):
 
             elif relation.modeltypestr == "BetaBernoulli":
                 pass
+            elif relation.modeltypestr == "NormalInverseChiSq":
+                pass
             elif relation.modeltypestr == "GammaPoisson":
                 pass
 

### 2014-01-08 irm/models.py

@@ -1,6 +1,7 @@
 import numpy as np
 from scipy.special import betaln
 import util
+from scipy.stats import chi2, t, norm
 
 class BetaBernoulli(object):
     def create_hps(self):
@@ -788,6 +789,65 @@ class LogisticDistancePoisson(object):
         
 
 
+
+class NormalInverseChiSq(object):
+    """
+    """
+    def data_dtype(self):
+        """
+        """
+        return np.float32
+
+    
+    def sample_hps(self):
+        """
+        draw a sample of the HPs from some prior
+        """
+        return {'mu' :  np.random.normal(0, 5), 
+                'kappa' : np.random.gamma(1.2, 4), 
+                'sigmasq' : np.random.gamma(1.2, 4), 
+                'nu' : np.random.uniform(1.0, 5.0)}
+
+    def sample_param(self, hps):
+        """
+        draw a sample 
+
+        """
+        sigmasq = hps['nu'] * hps['sigmasq'] / chi2.rvs(hps['nu'])
+        
+        std = np.sqrt(sigmasq / hps['kappa'])
+        mu = np.random.normal(hps['mu'], std)
+
+        
+        return {'sigmasq' : sigmasq, 
+                'mu' : mu}
+
+    def sample_data(self, ss, hps):
+        """
+        NOTE THIS ONLY SAMPLES FROM THE PARAM P and not from 
+        suffstats. 
+        """
+        
+        mu = ss['mu']
+        sigmasq = ss['sigmasq']
+
+
+        return np.random.normal(mu, np.sqrt(sigmasq))
+
+    def est_parameters(self, data, hps):
+        """
+        A vector of data for this component, and the hypers
+        
+        """
+
+    def param_eval(self, d, ss, hps):
+        """
+        At distance dist, evaluate the prob of connection
+        """
+        raise NotImplemented("No distance dependentce") 
+        
+
+
 NAMES = {'BetaBernoulli' : BetaBernoulli, 
          'BetaBernoulliNonConj' : BetaBernoulliNonConj, 
          'SigmoidDistance' : SigmoidDistance, 
@@ -798,6 +858,7 @@ NAMES = {'BetaBernoulli' : BetaBernoulli,
          'NormalDistanceFixedWidth': NormalDistanceFixedWidth, 
          'SquareDistanceBump' : SquareDistanceBump, 
          'ExponentialDistancePoisson' : ExponentialDistancePoisson, 
[…]

### 2014-01-08 irm/pyirm.cc

@@ -28,9 +28,13 @@ IComponentContainer * create_component_container(std::string data, bp::tuple dat
     if(modeltype == "BetaBernoulli") { 
         IComponentContainer * cc = new ComponentContainer<BetaBernoulli>(data, data_dims_v); 
         return cc; 
-    } else    if(modeltype == "GammaPoisson") { 
+     } else    if(modeltype == "GammaPoisson") { 
 
         IComponentContainer * cc = new ComponentContainer<GammaPoisson>(data, data_dims_v); 
+        return cc; 
+     } else    if(modeltype == "NormalInverseChiSq") { 
+
+        IComponentContainer * cc = new ComponentContainer<NormalInverseChiSq>(data, data_dims_v); 
         return cc; 
     }     else if(modeltype == "BetaBernoulliNonConj") { 
         IComponentContainer * cc = new ComponentContainer<BetaBernoulliNonConj>(data, data_dims_v); 

### 2014-01-08 irm/pyirmutil.py

@@ -12,6 +12,9 @@ class Relation(object):
         elif isinstance(modeltype, models.GammaPoisson):
             modeltypestr = "GammaPoisson"
             self.conjugate = True
+        elif isinstance(modeltype, models.NormalInverseChiSq):
+            modeltypestr = "NormalInverseChiSq"
+            self.conjugate = True
         elif isinstance(modeltype, models.BetaBernoulliNonConj):
             modeltypestr = "BetaBernoulliNonConj"
         elif isinstance(modeltype, models.AccumModel):

### 2014-01-08 irm/util.h

@@ -201,10 +201,22 @@ inline float log_sum_exp(float x, float y) {
         b = x; 
     }
     return a + log(1.0 + exp(b - a)); 
-    
-
 }
 
+inline float log_t_pdf(float x, float nu, float mu, float sigmasq)
+{
+    /*
+    Murphy, Eq. 304
+
+    */
+    float c = lgamma(.5 * (nu + 1.))
+        - (lgamma(.5 * nu) + .5 * (logf(nu * 3.1415926535f* sigmasq))); 
+    float xt = (x - mu); 
+    float s = xt * xt / sigmasq; 
+    float d = -(.5 * (nu + 1.)) * logf(1. + s / nu); 
+    return c + d; 
+    
+}
 
 #ifdef USE_LOGEXP_APPROX
 #define MYLOG fasterlog

### 2014-01-08 tests/test_inference.py

@@ -19,10 +19,14 @@ MODELS =  ["BetaBernoulliNonConj",
            'GammaPoisson',
            'BetaBernoulli', 
            'NormalDistanceFixedWidth', 
-    #'SquareDistanceBump',  # NOTE THIS IS SO NOT READY FOR PRODUCTION the delta prob is broken
+
            'ExponentialDistancePoisson',
            'LogisticDistancePoisson'
+
+    #'SquareDistanceBump',  # NOTE THIS IS SO NOT READY FOR PRODUCTION the delta prob is broken
 ]
+MODELS = ['NormalInverseChiSq']
+
 
 KERNELS = {'default_nonconj' : irm.runner.default_kernel_nonconj_config(), 
            #'default_anneal':  irm.runner.default_kernel_anneal()
